### 2017-02-14 paper155.tex
First commit of MICCAI 2017 paper

\documentclass[runningheads,a4paper]{llncs}

%
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{url}
\usepackage{graphicx}
\usepackage{multirow}
%
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\operatorname{min}}\;}
\mathchardef\mhyphen="2D
\raggedbottom 
%
% Allow easy processing of labeled images in figures
\newcounter{lfigcounter}
\newenvironment{IonTab}{\begin{table}[htb]}{\end{table}}
\newenvironment{IonFig}{\setcounter{lfigcounter}{1}\begin{figure}} {\end{figure}}
\newenvironment{IonFigH}{\setcounter{lfigcounter}{1}\begin{figure}[h]}{\end{figure}}
\newenvironment{IonFigT}{\setcounter{lfigcounter}{1}\begin{figure}[!t]}{\end{figure}}
\newenvironment{IonFigB}{\setcounter{lfigcounter}{1}\begin{figure}[b]}{\end{figure}}
\def\ionbox#1{\makebox[#1]{(\alph{lfigcounter})}\stepcounter{lfigcounter}}
%
\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{An Image-based instantaneous phase estimation method for gating and temporal super-resolution of cardiac ultrasound}

% a short form should be given in case it is too long for the running head
\titlerunning{Image-based Phase Estimation Method for Cardiac Ultrasound}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%

% anonymous stuff
\author{*}
\authorrunning{*}   
\tocauthor{*}
\institute{*}

\maketitle

\begin{abstract}
Point-of-care low-cost hand-held ultrasound devices are emerging as an increasingly effective tool for rapid non-invasive visual assessment of cardiac structure and function on the bed-side. Their diagnostic value can be enriched further by coupling them with automated image analysis algorithms to accurately detect and quantify any cardiac abnormalities at the point-of-care potentially even in asymptomatic patients. However, a major challenge to this is the poor spatio-temporal resolution and high noise level of images obtained using these devices. In an effort to address this problem, we present a novel method for de-noising, gating, and temporal super-resolution of noisy low frame-rate periodic image sequences such as the ones obtained using low-cost hand-held cardiac ultrasound devices. We first leverage the inherent image-level redundancy in a periodic image sequence to decouple signal from noise using a low-rank + sparse decomposition algorithm. Next, we transform the complex high-dimensional periodic image sequence to a narrow-band 1D time series with the same periodicity characteristics using a novel strategy based on inter-frame similarity. Next, we decompose this time series use a season-trend decomposition method to obtain the heart-beat (season) and respiratory (trend) signals. Next, we use the Hilbert transform to estimate the instantaneous cardiac and respiratory phases of each frame to facilitate gating and super-resolution. Lastly, we use Nadarya-Watson kernel regression to construct an image manifold parameterized by intra-period phase which can then be sampled to reconstruct a de-noised single-period image sequence at a higher temporal-resolution. We demonstrate the utility and efficacy of our method both on synthetic data and on real image sequences of the beating heart acquired using a low-cost hand-held ultrasound device.
%\keywords{Denoising, Gating, Super-resolution, hand-held, Cardiac, Ultrasound}
\end{abstract}
\vspace{-0.5cm}
\section{Introduction}
\label{sec:intro}
Low-cost hand-held ultrasound devices are emerging as an increasingly effective tool for rapid non-invasive visual assessment of cardiac structure and function on the bed-side\cite{Wiley2014}. Their diagnostic value can be enriched further by coupling them with automated image analysis algorithms to accurately detect and quantify any cardiac abnormalities at the point-of-care potentially even in asymptomatic patients. However, a major challenge to this endeavour is the poor spatio-temporal resolution and high noise level of images obtained using these devices. Fortunately, in a periodic image sequence, such as the ones obtained in cardiac ultrasound, the expected temporal image-level redundancy/repetition can be exploited both for de-noising and super-resolution. These are precisely the problems that we intend to address here. Specifically, in this paper, we present a novel method for de-noising, gating and temporal super-resolution of noisy low frame-rate periodic image sequences such as the ones obtained using low-cost hand-held cardiac ultrasound devices. To our knowledge, there only one paper \cite{Makihara2011} that tries to address a similar problem on human gait sequences but the challenges posed by cardiac ultrasound videos that we intend to address here are different.
\vspace{-0.5cm}
\section{Method}
\label{sec:method}
%
In this section, we present the theory underlying the proposed method for de-noising, gating and temporal super-resolution of a noisy periodic image sequence. Specifically, we begin by describing how we de-noise a periodic image sequence using a low-rank $+$ sparse decomposition algorithm in Section \ref{sec:method:de-noising}. We then present a novel method for estimating the intra-period phase of each frame using the de-noised image sequence in Section \ref{sec:method:phase_estimation}. Finally, in Section \ref{sec:method:super-resolution}, we present a method that uses the phase-tagged image sequence to reconstruct a single-period video at higher-temporal resolution. 
%
\begin{IonFigT}
\centering
%
\makebox[1.5in]{Input}\makebox[1.5in]{LowRank}\makebox[1.5in]{Sparse}\\
\includegraphics[width=1.5in]{figures/VibratingCircle/imInput.png}
\includegraphics[width=1.5in]{figures/VibratingCircle/imLowRank.png}
\includegraphics[width=1.5in]{figures/VibratingCircle/imSparse.png}\\
\vspace{0.1cm}
\includegraphics[width=1.5in]{figures/VibratingCircle/imInput_MidX.png}
\includegraphics[width=1.5in]{figures/VibratingCircle/imLowRank_MidX.png}
\includegraphics[width=1.5in]{figures/VibratingCircle/imSparse_MidX.png}
\ionbox{4.5in}\\
\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imInput.png}
\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imLowRank.png}
\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imSparse.png}\\
\vspace{0.1cm}
\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imInput_MidX.png}
\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imLowRank_MidX.png}
\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imSparse_MidX.png}
\ionbox{4.5in}\\
%
\caption{Noise suppression using low-rank + sparse decomposition in (a) synthetic video of an expanding/contracting circle, and (b) real ultrasound video of beating heart. Images show temporal maximum intensity projections (top-row) and mid Y-T slices (bottom row) of the input (left), low-rank (middle), and sparse(right) videos.}
\label{fig:de-noising}
\end{IonFigT}
\vspace{-0.5cm}
\subsection{Noise suppression using low-rank $+$ sparse decomposition}
\label{sec:method:de-noising}
One of the main challenges for quantitative analysis of ultrasound image sequences in general, and the ones obtained using low-cost hand-held devices of interest to us in particular, is their high noise level. Fortunately, in a periodic image sequence, such as the ones seen in cardiac ultrasound, the expected image-level redundancy/repetition in time can be exploited to decouple the underlying signal from any distortions such as noise/occlusion if they do not affect the same set of pixels consistently over time. Inspired by their successful application to the problem of background-subtraction in surveillance videos\cite{Candes2009} by exploiting a similar kind of temporal redundancy, we use a low-rank $+$ sparse decomposition algorithm to suppress the noise in a periodic image sequence. Specifically, given a periodic image sequence $P = \{I_1, I_2, ..., I_n\}$ containing $n$ images with $m$ pixels each, we create an $m \times n$ matrix $D$ by stacking the images in the sequence as columns such that $i^{th}$ column of $D$ is a vectorized form of  image $I_i$ in the sequence $P$. We then decompose $D$ into the sum of a low-rank matrix $L$ and a sparse matrix $S$ by solving the following convex optimization problem using an augmented lagrangian approach:
\begin{equation}
\argmin{L,S}\left(\|L\|_{*} + \lambda \|S\|_1\right) 
\; s.t. \; D = L + S
\end{equation}
where $\| \cdot \|_*$ and $\| \cdot \|_1$ are the nuclear-norm and $\ell \mhyphen 1$ norm, respectively, and $\lambda > 0$ is a balancing parameter that we set to $1/{\sqrt{max\left(m,n\right)}}$. Due to the expected repetition of images across periods in a periodic image sequence and assuming that the noise/occlusions corrupting the images are sparse in time, the low-rank component $L$ approximates the noise-free image sequence and the sparse component $S$ captures the residual noise. Prior to low-rank + sparse decomposition, we apply a median filter to suppress any impulse noise and downsize the images if necessary to reduce computational time. For further analysis, we use the de-noised sequence of images $\{L_1, ..., L_n\}$ in the low-rank matrix $L$.
%
%
\begin{IonFigT}
\centering
%
\makebox[1.5in]{Similarity}\makebox[1.5in]{Similarity - Trend}\makebox[1.5in]{Similarity - Detrended}\\
\begin{tabular}{cc}
	\includegraphics[width=1.5in]{figures/VibratingCircle/simMat.png}
	\includegraphics[width=1.5in]{figures/VibratingCircle/simMat_Trend.png}
	\includegraphics[width=1.5in]{figures/VibratingCircle/simMat_Seasonal.png}
\end{tabular}\\
\ionbox{4.5in}\\
%
\begin{tabular}{cc}
	\begin{tabular}{c}
		\includegraphics[width=3.5in]{figures/VibratingCircle/ts_seasonal.png}\\
		\includegraphics[width=3.5in]{figures/VibratingCircle/ts_instaphase_comparison.png}
	\end{tabular}&	
	\begin{tabular}{c}

		\includegraphics[width=1.0in]		
			{figures/VibratingCircle/Input_est_phaseord_MidX.png}\\
		\makebox[1.0in]{Input}\\				
		\includegraphics[width=1.0in]		
			{figures/VibratingCircle/Lowrank_est_phaseord_MidX.png}\\		
		\makebox[1.0in]{Low-rank}				
	\end{tabular}			
\end{tabular}\\
\ionbox{3.5in}\ionbox{1.0in}\\
\caption{Intra-period phase estimation on synthetic expanding-contracting circle video: (a-left) Inter-frame similarity matrix, (a-mid) Trend computed using local regression, (a-right) Detrended similarity matrix, (b) Similarity profile of the row in the detrended similarity matrix with minimum spectral entropy (top) and the instantaneous intra-period phase (bottom) computed using the hilbert transform along with the groundtruth phaseto generate the input video, (c) Mid Y-T slice of input and low-rank videos with the frames ordered in ascending order of the estimated intra-period phase. Mean error in estimated phase against ground truth = 5.4 degrees.}
\label{fig:phase_estimation_synthetic}
\end{IonFigT}
%
%
\begin{IonFigT}
\centering
%
\makebox[1.5in]{Similarity}\makebox[1.5in]{Similarity - Trend}\makebox[1.5in]{Similarity - Detrended}\\
\begin{tabular}{cc}
	\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/simMat.png}
	\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/simMat_Trend.png}
	\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/simMat_Seasonal.png}
\end{tabular}\\
\ionbox{4.5in}\\
%
\begin{tabular}{cc}
	\begin{tabular}{c}
		\includegraphics[width=2.75in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/ts_seasonal.png}\\
		\includegraphics[width=2.75in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/ts_instaphase.png}
	\end{tabular}&	
	\begin{tabular}{c}
		\includegraphics[width=1.70in]		
			{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/Input_est_phaseord_MidX.png}\\
		\makebox[1.7in]{Input}\\				
	\end{tabular}			
\end{tabular}\\
\ionbox{2.75in}\ionbox{1.70in}\\
\caption{Intra-period phase estimation on cardiac ultrasound video: (a-left) Inter-frame similarity matrix, (a-mid) Trend computed using local regression, (a-right) Detrended similarity matrix, (b) Similarity profile of the row in the detrended similarity matrix with minimum spectral entropy (top) and the instantaneous intra-period phase (bottom) computed using the hilbert transform, (c) Mid Y-T slice of input and low-rank videos with the frames ordered in ascending order of the estimated intra-period phase.}
%
\label{fig:phase_estimation_heart}
\end{IonFigT}
%
\vspace{-0.5cm}
%
\subsection{Intra-period phase estimation of each frame}
\label{sec:method:phase_estimation}
%
To facilitate gating and super-resolution, we need to estimate the intra-period phase/position of each frame in the periodic image sequence. When compared to the widely studied problem of super-resolution using multiple low-resolution images of the same scene\cite{Nasrollahi2014}, the process of intra-period phase estimation in a periodic image sequence is analogous to alignment of the low-resolution images at a sub-pixel precision using a registration algorithm. While there have been numerous efforts for the estimation of instantaneous intra-period phase\cite{Lu2013} in periodic univariate time series data (wherein a single-variable is measured at each time point), we are not aware of any methods that tackle this problem in a multivariate setting which is the case with periodic image sequences wherein upto a million variables (intensities of pixels) could be measured at each time point. Our strategy was to find a way to simplify this seemingly complex multi-variate problem by transforming it into a univariate one and take advantage of existing methods to solve the problem at hand. To accomplish this, we compute the similarity/affinity between all pairs of frames in the given periodic image sequence containing $n$ images to create a matrix $A \in R^{n \times n}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames of the given sequence. In principle, any of the similarity metrics used in image registration algorithms can be chosen to quantify the similarity between a given pair of frames. Here in, we treat the images in the given sequence as points in high-dimensional space, apply principal component analysis (PCA) to reduce their dimensionality\cite{Bishop2006} such that 99\% of the total variance is covered, and use the $-ve$ distance in the PCA reduced subspace to quantify the similarity between any given pair of frames. Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series that measures the similarity of the corresponding frame in the image sequence with all other frames. Additionally, for a periodic image sequence, we can expect them to be periodic provided that the corresponding frame is not significantly corrupted by distortions such as noise and changes in global illumination over time among others. To mitigate the potential problem of global illumintation changes over time, we detrend and extract the seasonal part of the time series corresponding to each row of $A$ by dividing it with the trend signal estimated using a locally-weighted regression algorithm called loess\cite{Cleveland1990}. Let $A_s$ be the matrix whose rows contain the seasonal part obtained by detrending the corresponding row in $A$. Next, we minimize the risk from noise, by finding the row $u(t)$ of $A_s$ whose power spectrum or periodogram has minimum entropy. In our experiments, the time series $u(t)$ was nearly sinusoidal in morphology with most of its energy in its power spectrum concentrated within a very narrow-band with the same periodicity characteristics as the image sequence it was derived from. Based on this observation, we decided to use its Hilbert transform \cite{Lu2013} to estimate the instantaneous intra-period phase  of each frame. Specifically, we compute the instantaneous phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the time series $u(t)$ as follows:
\begin{equation}
\phi(t) = arctan \left( \frac{H_u(t)}{u(t)}\right)
\end{equation} 
where $H_u(t)$ is the Hilbert transform of $u(t)$. Lastly, we map the instantaneous phase $\phi(t)$ to the range $\left [  0, 1\right )$ which can then be used to obtain the intra-period phase of each frame.
%
%
\begin{IonFigT}
\centering
%
\includegraphics[height=0.9in]{figures/VibratingCircle/phase_derivative_error.png}
%
\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_2_linear_interpolation.png}
\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_4_linear_interpolation.png}
\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_8_linear_interpolation.png}
\hspace{.2in}
%
\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_2_kernel_regression.png}
\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_4_kernel_regression.png}
\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_8_kernel_regression.png}
\hspace{.2in}
%
\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_2_bspline_registration.png}
\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_4_bspline_registration.png}
\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_8_bspline_registration.png}
\hspace{.2in}
\\
\ionbox{1.0in}\ionbox{1.0in}\ionbox{1.0in}\ionbox{1.0in}
%
\\
%
%%\makebox[0.1in]{}
%\includegraphics[height=0.8in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imHeartSuperRes_ValveMMode_4.png}
%\includegraphics[height=0.8in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imHeartSuperRes_ValveMMode_8.png}
%\includegraphics[height=0.8in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imHeartSuperRes_ValveMMode_16.png}\\
%\ionbox{3in}\makebox[0.1in]{}\ionbox{1in}
%
\caption{(a) Error with groundtruth of the phase difference between consecutive frames of phase-ordered synthetic video generated with varying levels of noise, (c-d) Mid YT slices of reconstructed one-period video of expanding-contracting circle at magnifications (left-right) of 2, 4, and 8 respectively}
\label{fig:super_resolution}
\end{IonFigT}
%
\vspace{-0.5cm}
%
\subsection{Reconstruction of one-period video at high temporal resolution}
\label{sec:method:super-resolution}
%
Once the intra-period phase of each frame has been estimated, it can be used (i) to facilitate the extraction of quantitative measurements from a desired part of the period, a process commonly referred to as gating, and (ii) to enable spatio-temporal super-resolution by considering the set of images within each period of a periodic image sequence as distinct observations of the same data. In this section, we present a method that uses a phase-tagged periodic image sequence to reconstruct a single-period video at a higher temporal resolution.  Given a de-noised periodic sequence $\{L_1, ..., L_n\}$ of $n$ images (with $m$ pixels each) along with their estimated intra-period phases $\{\phi_1, ..., \phi_n\}$, we first use Nadarya-Watson kernel regression\cite{Bishop2006} to construct a phase-parameterized image manifold in the form of a function $M(\phi): [0, 1) \to R^m $ that maps any intra-period phase $\phi$ to an image as follows:
\begin{equation}
M(\phi) = \frac{\sum_{i = 1}^{n} K(\phi, \phi_i) L_i}{\sum_{i = 1}^{n} K(\phi, \phi_i)} 
\end{equation}
where in we define the kernel $K$ using a gausssian with standard-deviation $\sigma$ equal to a multiple of the average distance between consecutive phases after sorting them. Note that, we account for the periodic nature of the intra-period phase whenever we measure distance between two phase values. A single-period video can now be reconstructed by sampling the manifold $M(\phi)$ at any desired resolution from the range $[0, 1)$ of intra-period phase. Note that the aformentioned process will facilitate temporal super-resolution only if the set of images observed in each distinct period of the given image sequence differ in intra-period phase by some non-zero amount less than its temporal sampling interval. Also, the amount of temporal super-resolution possible will depend both on the number of periods observed and the maximum amount by which one can expect the images to be perturbed in phase across observations. 
\vspace{-0.5cm}
\section{Results}
\label{sec:results}
We demonstrate the utility and efficacy of our method on synthetic videos of a periodically expanding-contracting circular membrane in Section~\ref{sec:results:vibrating_circle} and on real b-mode ultrasound videos of the beating human heart acquired in Section~\ref{sec:results:beating_heart}. 
\vspace{-0.5cm}
\subsection{Synthetic data: expanding-contracting circular membrane}
\label{sec:results:vibrating_circle}
In this section, we discuss the results obtained using the proposed method on synthetic videos of a circular membrane that expands and contracts periodically analogous to ultrasound videos of the beating heart in short-axis view. These videos are synthesized by  generating images of a circle whose radius various according to a sinusoidal function within a user-specified range of $[r_{min}, r_{max}]$. The images are smoothened using a gaussian filter of standard deviation $\sigma_m \times (r_{max} - r_{min})$ that determines the thickness of the circular membrane and corrupted by adding gaussian noise with standard deviation $\sigma_n$ . The frame length of the video is determined by number of cycles/periods of contraction-expansion $n_p$ and the number of samples within each period $n_s$. Lastly, to simulate phenomenon analogous to device-originated sampling noise and irregularities in the heart beat, we also perturb the sample phases by an amount of $\delta_p \in (0, 1]$ times the sampling interval before perturbation. Typical baseline values for the aforementioned parameters in our experiments are as follows: $r_{min} = 20, \; r_{max} = 50, \; \sigma_m = 0.1, \; \sigma_n = 0.1, \; n_p = 30, \; n_s = 20$. The intensity range of the images is set to $[0, 1]$. Fig.~\ref{fig:de-noising}(a) depicts the de-noising result obtained using low-rank $+$ sparse decomposition on a sample dataset as described in Section~\ref{sec:method:de-noising}. Observe the way in which most of the noise was absorbed by the sparse component. Fig.~\ref{fig:phase_estimation_synthetic} depicts the results of our phase estimation strategy described in Section~\ref{sec:method:phase_estimation}. Notice the periodic nature of images depicting the detrended inter-frame similarity matrix in Fig.~\ref{fig:phase_estimation_synthetic}(a) and the sinusoidal morphology of the similarity profile in Fig.~\ref{fig:phase_estimation_synthetic}(b) picked automatically for estimating the intra-period phase. Notice the continuity of the membrane in the Mid Y-T slice of the input and low-rank videos ordered by the estimated intra-period phase in Fig.~\ref{fig:phase_estimation_heart}. The mean and standard deviation of the mean error between the estimated and groundtruth intra-period phase accross five videos generated as described above was 7.56 and 1.346 degrees, respectively. Fig.~\ref{fig:super_resolution} shows the results of super-resolution at a series of increasing magnifications.
\vspace{-0.1cm}
\subsection{Real data: ultrasound video of beating heart}
\label{sec:results:beating_heart}
In this section, we discuss the results obtained using the proposed method on real b-mode ultrasound videos of the beating heart acquired using a hand-held ultrasound device. Each video was acquired at a frame rate of 36 frames per second with 14-17 heart beats during which the subject as asked to hold breath. Fig~\ref{fig:de-noising}(b) depicts the de-noising result obtained using low-rank $+$ sparse decomposition on a sample dataset as described in Section~\ref{sec:method:de-noising}. Notice that a significant percentage of the speckle noise was again absorbed by the sparse part. Fig~\ref{fig:phase_estimation_heart} depicts the results of our phase estimation strategy described in Section~\ref{sec:method:phase_estimation}. Again, notice the periodic pattern of the detrended inter-frame similarity matrix in Fig~\ref{fig:phase_estimation_heart}(a) and the near sinusoidal morphology of the similarity profile used to estimate the intra-period phase in Fig~\ref{fig:phase_estimation_heart}(b). Lastly, notice the continuity of the chamber wall over-time and its apparent expanding/contracting motion in Fig.~\ref{fig:phase_estimation_heart}(c).
\vspace{-0.5cm}
\section{Discussion and future work}
\label{sec:discussion}
The novel contributions of this paper are: (i) to leverage the inherent temporal image-level repetition/redundancy to decouple signal from noise in a noisy periodic image sequence, (ii) the proposed strategy based on inter-frame similarity to transform the multivariate problem of instantaneous intra-period phase estimation in a noisy complex high-dimensional periodic image sequence into a univariate problem to enable the use of existing methods in univariate time series analysis, (iii) using our method it may now become possible to gate or extract measurements from a desired portion of the cardiac cycle without using any other hardware thereby enriching the value of low-cost hand-held cardiac ultrasound devices at the point-of-care, and (iv) using our phase-estimation strategy the community can now translate the mature super-resolution technology in field of computer vision~\cite{Nasrollahi2014} to enhance the value of low-cost hand-held ultrasound devices even further. Having demonstrated a proof of concept of our ideas in this paper, our method has to be better evaluated on more datasets and especially on cardiac ultrasound videos of irregular heart beats. Lastly, since this paper deals with video analysis we encourage the reader to review the supplementary material for a better understanding of the proposed method.
\vspace{-0.5cm}
\bibliographystyle{splncs03}
\bibliography{library}
\end{document}

### 2017-02-15 paper155.tex
Removed old text and created basic skeleton

@@
 We demonstrate the utility and efficacy of our method both on synthetic data and on real image sequences of the beating heart acquired using a low-cost hand-held ultrasound device.
-%\keywords{Denoising, Gating, Super-resolution, hand-held, Cardiac, Ultrasound}
+%\keywords{Cardiac, Ultrasound, Phase estimation, Gating, Temporal super-resolution}
 \end{abstract}
+
 \vspace{-0.5cm}
@@
 \label{sec:intro}
-Low-cost hand-held ultrasound devices are emerging as an increasingly effective tool for rapid non-invasive visual assessment of cardiac structure and function on the bed-side\cite{Wiley2014}.
-Their diagnostic value can be enriched further by coupling them with automated image analysis algorithms to accurately detect and quantify any cardiac abnormalities at the point-of-care potentially even in asymptomatic patients.
-However, a major challenge to this endeavour is the poor spatio-temporal resolution and high noise level of images obtained using these devices.
-Fortunately, in a periodic image sequence, such as the ones obtained in cardiac ultrasound, the expected temporal image-level redundancy/repetition can be exploited both for de-noising and super-resolution.
-These are precisely the problems that we intend to address here.
-Specifically, in this paper, we present a novel method for de-noising, gating and temporal super-resolution of noisy low frame-rate periodic image sequences such as the ones obtained using low-cost hand-held cardiac ultrasound devices.
-To our knowledge, there only one paper \cite{Makihara2011} that tries to address a similar problem on human gait sequences but the challenges posed by cardiac ultrasound videos that we intend to address here are different.
+%
+
+%
 \vspace{-0.5cm}
@@
 %
-In this section, we present the theory underlying the proposed method for de-noising, gating and temporal super-resolution of a noisy periodic image sequence.
-Specifically, we begin by describing how we de-noise a periodic image sequence using a low-rank $+$ sparse decomposition algorithm in Section \ref{sec:method:de-noising}.
-We then present a novel method for estimating the intra-period phase of each frame using the de-noised image sequence in Section \ref{sec:method:phase_estimation}.
-Finally, in Section \ref{sec:method:super-resolution}, we present a method that uses the phase-tagged image sequence to reconstruct a single-period video at higher-temporal resolution.
-%
-\begin{IonFigT}
-\centering
-%
-\makebox[1.5in]{Input}\makebox[1.5in]{LowRank}\makebox[1.5in]{Sparse}\\
-\includegraphics[width=1.5in]{figures/VibratingCircle/imInput.png}
-\includegraphics[width=1.5in]{figures/VibratingCircle/imLowRank.png}
-\includegraphics[width=1.5in]{figures/VibratingCircle/imSparse.png}\\
-\vspace{0.1cm}
-\includegraphics[width=1.5in]{figures/VibratingCircle/imInput_MidX.png}
-\includegraphics[width=1.5in]{figures/VibratingCircle/imLowRank_MidX.png}
-\includegraphics[width=1.5in]{figures/VibratingCircle/imSparse_MidX.png}
-\ionbox{4.5in}\\
-\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imInput.png}
-\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imLowRank.png}
-\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imSparse.png}\\
-\vspace{0.1cm}
-\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imInput_MidX.png}
-\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imLowRank_MidX.png}
-\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imSparse_MidX.png}
-\ionbox{4.5in}\\
-%
-\caption{Noise suppression using low-rank + sparse decomposition in (a) synthetic video of an expanding/contracting circle, and (b) real ultrasound video of beating heart.
-Images show temporal maximum intensity projections (top-row) and mid Y-T slices (bottom row) of the input (left), low-rank (middle), and sparse(right) videos.}
-\label{fig:de-noising}
-\end{IonFigT}
-\vspace{-0.5cm}
-\subsection{Noise suppression using low-rank $+$ sparse decomposition}
-\label{sec:method:de-noising}
-One of the main challenges for quantitative analysis of ultrasound image sequences in general, and the ones obtained using low-cost hand-held devices of interest to us in particular, is their high noise level.
-Fortunately, in a periodic image sequence, such as the ones seen in cardiac ultrasound, the expected image-level redundancy/repetition in time can be exploited to decouple the underlying signal from any distortions such as noise/occlusion if they do not affect the same set of pixels consistently over time.
-Inspired by their successful application to the problem of background-subtraction in surveillance videos\cite{Candes2009} by exploiting a similar kind of temporal redundancy, we use a low-rank $+$ sparse decomposition algorithm to suppress the noise in a periodic image sequence.
-Specifically, given a periodic image sequence $P = \{I_1, I_2, ..., I_n\}$ containing $n$ images with $m$ pixels each, we create an $m \times n$ matrix $D$ by stacking the images in the sequence as columns such that $i^{th}$ column of $D$ is a vectorized form of  image $I_i$ in the sequence $P$.
-We then decompose $D$ into the sum of a low-rank matrix $L$ and a sparse matrix $S$ by solving the following convex optimization problem using an augmented lagrangian approach:
-\begin{equation}
-\argmin{L,S}\left(\|L\|_{*} + \lambda \|S\|_1\right) 
-\; s.t. \; D = L + S
-\end{equation}
-where $\| \cdot \|_*$ and $\| \cdot \|_1$ are the nuclear-norm and $\ell \mhyphen 1$ norm, respectively, and $\lambda > 0$ is a balancing parameter that we set to $1/{\sqrt{max\left(m,n\right)}}$.
-Due to the expected repetition of images across periods in a periodic image sequence and assuming that the noise/occlusions corrupting the images are sparse in time, the low-rank component $L$ approximates the noise-free image sequence and the sparse component $S$ captures the residual noise.
-Prior to low-rank + sparse decomposition, we apply a median filter to suppress any impulse noise and downsize the images if necessary to reduce computational time.
-For further analysis, we use the de-noised sequence of images $\{L_1, ..., L_n\}$ in the low-rank matrix $L$.
-%
-%
-\begin{IonFigT}
-\centering
-%
-\makebox[1.5in]{Similarity}\makebox[1.5in]{Similarity - Trend}\makebox[1.5in]{Similarity - Detrended}\\
-\begin{tabular}{cc}
-	\includegraphics[width=1.5in]{figures/VibratingCircle/simMat.png}
-	\includegraphics[width=1.5in]{figures/VibratingCircle/simMat_Trend.png}
-	\includegraphics[width=1.5in]{figures/VibratingCircle/simMat_Seasonal.png}
-\end{tabular}\\
-\ionbox{4.5in}\\
-%
-\begin{tabular}{cc}
-	\begin{tabular}{c}
-		\includegraphics[width=3.5in]{figures/VibratingCircle/ts_seasonal.png}\\
-		\includegraphics[width=3.5in]{figures/VibratingCircle/ts_instaphase_comparison.png}
-	\end{tabular}&	
-	\begin{tabular}{c}
 
-		\includegraphics[width=1.0in]		
-			{figures/VibratingCircle/Input_est_phaseord_MidX.png}\\
-		\makebox[1.0in]{Input}\\				
-		\includegraphics[width=1.0in]		
-			{figures/VibratingCircle/Lowrank_est_phaseord_MidX.png}\\		
-		\makebox[1.0in]{Low-rank}				
-	\end{tabular}			
-\end{tabular}\\
-\ionbox{3.5in}\ionbox{1.0in}\\
-\caption{Intra-period phase estimation on synthetic expanding-contracting circle video: (a-left) Inter-frame similarity matrix, (a-mid) Trend computed using local regression, (a-right) Detrended similarity matrix, (b) Similarity profile of the row in the detrended similarity matrix with minimum spectral entropy (top) and the instantaneous intra-period phase (bottom) computed using the hilbert transform along with the groundtruth phaseto generate the input video, (c) Mid Y-T slice of input and low-rank videos with the frames ordered in ascending order of the estimated intra-period phase.
-Mean error in estimated phase against ground truth = 5.4 degrees.}
-\label{fig:phase_estimation_synthetic}
-\end{IonFigT}
-%
-%
-\begin{IonFigT}
-\centering
-%
-\makebox[1.5in]{Similarity}\makebox[1.5in]{Similarity - Trend}\makebox[1.5in]{Similarity - Detrended}\\
-\begin{tabular}{cc}
-	\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/simMat.png}
-	\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/simMat_Trend.png}
-	\includegraphics[width=1.5in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/simMat_Seasonal.png}
-\end{tabular}\\
-\ionbox{4.5in}\\
-%
-\begin{tabular}{cc}
-	\begin{tabular}{c}
-		\includegraphics[width=2.75in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/ts_seasonal.png}\\
-		\includegraphics[width=2.75in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/ts_instaphase.png}
-	\end{tabular}&	
-	\begin{tabular}{c}
-		\includegraphics[width=1.70in]		
-			{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/Input_est_phaseord_MidX.png}\\
-		\makebox[1.7in]{Input}\\				
-	\end{tabular}			
-\end{tabular}\\
-\ionbox{2.75in}\ionbox{1.70in}\\
-\caption{Intra-period phase estimation on cardiac ultrasound video: (a-left) Inter-frame similarity matrix, (a-mid) Trend computed using local regression, (a-right) Detrended similarity matrix, (b) Similarity profile of the row in the detrended similarity matrix with minimum spectral entropy (top) and the instantaneous intra-period phase (bottom) computed using the hilbert transform, (c) Mid Y-T slice of input and low-rank videos with the frames ordered in ascending order of the estimated intra-period phase.}
-%
-\label{fig:phase_estimation_heart}
-\end{IonFigT}
 %
 \vspace{-0.5cm}
-%
-\subsection{Intra-period phase estimation of each frame}
+\subsection{Instantaneous phase estimation of each frame}
 \label{sec:method:phase_estimation}
 %
-To facilitate gating and super-resolution, we need to estimate the intra-period phase/position of each frame in the periodic image sequence.
-When compared to the widely studied problem of super-resolution using multiple low-resolution images of the same scene\cite{Nasrollahi2014}, the process of intra-period phase estimation in a periodic image sequence is analogous to alignment of the low-resolution images at a sub-pixel precision using a registration algorithm.
-While there have been numerous efforts for the estimation of instantaneous intra-period phase\cite{Lu2013} in periodic univariate time series data (wherein a single-variable is measured at each time point), we are not aware of any methods that tackle this problem in a multivariate setting which is the case with periodic image sequences wherein upto a million variables (intensities of pixels) could be measured at each time point.
-Our strategy was to find a way to simplify this seemingly complex multi-variate problem by transforming it into a univariate one and take advantage of existing methods to solve the problem at hand.
-To accomplish this, we compute the similarity/affinity between all pairs of frames in the given periodic image sequence containing $n$ images to create a matrix $A \in R^{n \times n}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames of the given sequence.
-In principle, any of the similarity metrics used in image registration algorithms can be chosen to quantify the similarity between a given pair of frames.
-Here in, we treat the images in the given sequence as points in high-dimensional space, apply principal component analysis (PCA) to reduce their dimensionality\cite{Bishop2006} such that 99\% of the total variance is covered, and use the $-ve$ distance in the PCA reduced subspace to quantify the similarity between any given pair of frames.
-Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series that measures the similarity of the corresponding frame in the image sequence with all other frames.
-Additionally, for a periodic image sequence, we can expect them to be periodic provided that the corresponding frame is not significantly corrupted by distortions such as noise and changes in global illumination over time among others.
-To mitigate the potential problem of global illumintation changes over time, we detrend and extract the seasonal part of the time series corresponding to each row of $A$ by dividing it with the trend signal estimated using a locally-weighted regression algorithm called loess\cite{Cleveland1990}.
-Let $A_s$ be the matrix whose rows contain the seasonal part obtained by detrending the corresponding row in $A$.
-Next, we minimize the risk from noise, by finding the row $u(t)$ of $A_s$ whose power spectrum or periodogram has minimum entropy.
-In our experiments, the time series $u(t)$ was nearly sinusoidal in morphology with most of its energy in its power spectrum concentrated within a very narrow-band with the same periodicity characteristics as the image sequence it was derived from.
-Based on this observation, we decided to use its Hilbert transform \cite{Lu2013} to estimate the instantaneous intra-period phase  of each frame.
-Specifically, we compute the instantaneous phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the time series $u(t)$ as follows:
-\begin{equation}
-\phi(t) = arctan \left( \frac{H_u(t)}{u(t)}\right)
-\end{equation} 
-where $H_u(t)$ is the Hilbert transform of $u(t)$.
-Lastly, we map the instantaneous phase $\phi(t)$ to the range $\left [  0, 1\right )$ which can then be used to obtain the intra-period phase of each frame.
-%
-%
-\begin{IonFigT}
-\centering
-%
-\includegraphics[height=0.9in]{figures/VibratingCircle/phase_derivative_error.png}
-%
-\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_2_linear_interpolation.png}
-\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_4_linear_interpolation.png}
-\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_8_linear_interpolation.png}
-\hspace{.2in}
-%
-\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_2_kernel_regression.png}
-\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_4_kernel_regression.png}
-\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_8_kernel_regression.png}
-\hspace{.2in}
-%
-\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_2_bspline_registration.png}
-\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_4_bspline_registration.png}
-\includegraphics[height=0.8in]{figures/VibratingCircle/Input_Mag_8_bspline_registration.png}
-\hspace{.2in}
-\\
-\ionbox{1.0in}\ionbox{1.0in}\ionbox{1.0in}\ionbox{1.0in}
-%
-\\
-%
-%%\makebox[0.1in]{}
-%\includegraphics[height=0.8in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imHeartSuperRes_ValveMMode_4.png}
-%\includegraphics[height=0.8in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imHeartSuperRes_ValveMMode_8.png}
-%\includegraphics[height=0.8in]{figures/SHeart2015-02-20-HRes-HFPS/Heart_20150220_151146/imHeartSuperRes_ValveMMode_16.png}\\
-%\ionbox{3in}\makebox[0.1in]{}\ionbox{1in}
-%
-\caption{(a) Error with groundtruth of the phase difference between consecutive frames of phase-ordered synthetic video generated with varying levels of noise, (c-d) Mid YT slices of reconstructed one-period video of expanding-contracting circle at magnifications (left-right) of 2, 4, and 8 respectively}
-\label{fig:super_resolution}
-\end{IonFigT}
+
 %
 \vspace{-0.5cm}
+\subsection{Gating of respiratory frames}
+\label{sec:method:gating}
 %
+
+%
+\vspace{-0.5cm}
 \subsection{Reconstruction of one-period video at high temporal resolution}
-\label{sec:method:super-resolution}
+\label{sec:method:gating}
 %
-Once the intra-period phase of each frame has been estimated, it can be used (i) to facilitate the extraction of quantitative measurements from a desired part of the period, a process commonly referred to as gating, and (ii) to enable spatio-temporal super-resolution by considering the set of images within each period of a periodic image sequence as distinct observations of the same data.
-In this section, we present a method that uses a phase-tagged periodic image sequence to reconstruct a single-period video at a higher temporal resolution.
-Given a de-noised periodic sequence $\{L_1, ..., L_n\}$ of $n$ images (with $m$ pixels each) along with their estimated intra-period phases $\{\phi_1, ..., \phi_n\}$, we first use Nadarya-Watson kernel regression\cite{Bishop2006} to construct a phase-parameterized image manifold in the form of a function $M(\phi): [0, 1) \to R^m $ that maps any intra-period phase $\phi$ to an image as follows:
-\begin{equation}
-M(\phi) = \frac{\sum_{i = 1}^{n} K(\phi, \phi_i) L_i}{\sum_{i = 1}^{n} K(\phi, \phi_i)} 
-\end{equation}
-where in we define the kernel $K$ using a gausssian with standard-deviation $\sigma$ equal to a multiple of the average distance between consecutive phases after sorting them.
-Note that, we account for the periodic nature of the intra-period phase whenever we measure distance between two phase values.
-A single-period video can now be reconstructed by sampling the manifold $M(\phi)$ at any desired resolution from the range $[0, 1)$ of intra-period phase.
-Note that the aformentioned process will facilitate temporal super-resolution only if the set of images observed in each distinct period of the given image sequence differ in intra-period phase by some non-zero amount less than its temporal sampling interval.
-Also, the amount of temporal super-resolution possible will depend both on the number of periods observed and the maximum amount by which one can expect the images to be perturbed in phase across observations.
+
 \vspace{-0.5cm}
@@
 \label{sec:results}
-We demonstrate the utility and efficacy of our method on synthetic videos of a periodically expanding-contracting circular membrane in Section~\ref{sec:results:vibrating_circle} and on real b-mode ultrasound videos of the beating human heart acquired in Section~\ref{sec:results:beating_heart}.
-\vspace{-0.5cm}
-\subsection{Synthetic data: expanding-contracting circular membrane}
-\label{sec:results:vibrating_circle}
-In this section, we discuss the results obtained using the proposed method on synthetic videos of a circular membrane that expands and contracts periodically analogous to ultrasound videos of the beating heart in short-axis view.
-These videos are synthesized by  generating images of a circle whose radius various according to a sinusoidal function within a user-specified range of $[r_{min}, r_{max}]$.
-The images are smoothened using a gaussian filter of standard deviation $\sigma_m \times (r_{max} - r_{min})$ that determines the thickness of the circular membrane and corrupted by adding gaussian noise with standard deviation $\sigma_n$ .
-The frame length of the video is determined by number of cycles/periods of contraction-expansion $n_p$ and the number of samples within each period $n_s$.
-Lastly, to simulate phenomenon analogous to device-originated sampling noise and irregularities in the heart beat, we also perturb the sample phases by an amount of $\delta_p \in (0, 1]$ times the sampling interval before perturbation.
-Typical baseline values for the aforementioned parameters in our experiments are as follows: $r_{min} = 20, \; r_{max} = 50, \; \sigma_m = 0.1, \; \sigma_n = 0.1, \; n_p = 30, \; n_s = 20$.
-The intensity range of the images is set to $[0, 1]$.
-Fig.~\ref{fig:de-noising}(a) depicts the de-noising result obtained using low-rank $+$ sparse decomposition on a sample dataset as described in Section~\ref{sec:method:de-noising}.
-Observe the way in which most of the noise was absorbed by the sparse component.
-Fig.~\ref{fig:phase_estimation_synthetic} depicts the results of our phase estimation strategy described in Section~\ref{sec:method:phase_estimation}.
-Notice the periodic nature of images depicting the detrended inter-frame similarity matrix in Fig.~\ref{fig:phase_estimation_synthetic}(a) and the sinusoidal morphology of the similarity profile in Fig.~\ref{fig:phase_estimation_synthetic}(b) picked automatically for estimating the intra-period phase.
-Notice the continuity of the membrane in the Mid Y-T slice of the input and low-rank videos ordered by the estimated intra-period phase in Fig.~\ref{fig:phase_estimation_heart}.
-The mean and standard deviation of the mean error between the estimated and groundtruth intra-period phase accross five videos generated as described above was 7.56 and 1.346 degrees, respectively.
-Fig.~\ref{fig:super_resolution} shows the results of super-resolution at a series of increasing magnifications.
-\vspace{-0.1cm}
-\subsection{Real data: ultrasound video of beating heart}
-\label{sec:results:beating_heart}
-In this section, we discuss the results obtained using the proposed method on real b-mode ultrasound videos of the beating heart acquired using a hand-held ultrasound device.
-Each video was acquired at a frame rate of 36 frames per second with 14-17 heart beats during which the subject as asked to hold breath.
-Fig~\ref{fig:de-noising}(b) depicts the de-noising result obtained using low-rank $+$ sparse decomposition on a sample dataset as described in Section~\ref{sec:method:de-noising}.
-Notice that a significant percentage of the speckle noise was again absorbed by the sparse part.
-Fig~\ref{fig:phase_estimation_heart} depicts the results of our phase estimation strategy described in Section~\ref{sec:method:phase_estimation}.
-Again, notice the periodic pattern of the detrended inter-frame similarity matrix in Fig~\ref{fig:phase_estimation_heart}(a) and the near sinusoidal morphology of the similarity profile used to estimate the intra-period phase in Fig~\ref{fig:phase_estimation_heart}(b).
-Lastly, notice the continuity of the chamber wall over-time and its apparent expanding/contracting motion in Fig.~\ref{fig:phase_estimation_heart}(c).
+%
+
+%
 \vspace{-0.5cm}
@@
 \label{sec:discussion}
-The novel contributions of this paper are: (i) to leverage the inherent temporal image-level repetition/redundancy to decouple signal from noise in a noisy periodic image sequence, (ii) the proposed strategy based on inter-frame similarity to transform the multivariate problem of instantaneous intra-period phase estimation in a noisy complex high-dimensional periodic image sequence into a univariate problem to enable the use of existing methods in univariate time series analysis, (iii) using our method it may now become possible to gate or extract measurements from a desired portion of the cardiac cycle without using any other hardware thereby enriching the value of low-cost hand-held cardiac ultrasound devices at the point-of-care, and (iv) using our phase-estimation strategy the community can now translate the mature super-resolution technology in field of computer vision~\cite{Nasrollahi2014} to enhance the value of low-cost hand-held ultrasound devices even further.
-Having demonstrated a proof of concept of our ideas in this paper, our method has to be better evaluated on more datasets and especially on cardiac ultrasound videos of irregular heart beats.
-Lastly, since this paper deals with video analysis we encourage the reader to review the supplementary material for a better understanding of the proposed method.
+%
+
+%
 \vspace{-0.5cm}

### 2017-02-16 paper155.tex
Rough first draft of phase estimation method section

@@
 % first the title is needed
-\title{An Image-based instantaneous phase estimation method for gating and temporal super-resolution of cardiac ultrasound}
+\title{An Image-based method for phase estimation, gating and temporal super-resolution of \\cardiac ultrasound}
 
@@
 %
-\vspace{-0.5cm}
+
+%\vspace{-0.5cm}
 \section{Method}
@@
 %
-
+In this section, we present the theory underlying the proposed methods for instantaneous phase estimation, gating and temporal super-resolution of cardiac ultrasound videos.
+Specifically, we begin by describing our method for the estimation of instantaneous cardiac and respiratory phases in
+Section \ref{sec:method:phase_estimation}.
+We then present a method that uses these phase estimates to robustly gate video frames predominantly influenced by respiratory motion in Section \ref{sec:method:gating}.
+Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression method that uses the phase-tagged image sequence to reconstruct a single-period video at higher-temporal resolution.
 %
-\vspace{-0.5cm}
+%
+\begin{IonFigT}
+\centering
+%
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/ecg_instaphase_overlay.png}\\
+\ionbox{4.5in}\\
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/qrs_peak_to_peak.png}
+\ionbox{4.5in}\\
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instaphase_valley_to_valley.png}
+\ionbox{4.5in}\\
+%
+\caption{Illustration of correspondence between ECG and the instantaneous cardiac phase estimated using the proposed method: (a) ECG signal (blue) and the cardiac phase estimated using our method using only image data.
+Also shown are the peaks of the QRS complex (pink circle) and the minima of the estimated cardiac phase (black circle) that correspond to the QRS peaks, (b) Five video frames evenly spaced in time between the first and second QRS peaks of the ECG signal constituting one cardiac cycle, and (c) Five video frames evenly spaced in time between first and second minima of the estimated cardiac phase constituting one cardiac cycle.}
+\label{fig:instaphase_vs_ecg}
+\end{IonFigT}
+%\vspace{-0.5cm}
 \subsection{Instantaneous phase estimation of each frame}
@@
 %
-
+While there have been numerous efforts for the estimation of instantaneous intra-period phase\cite{Lu2013} in periodic univariate time series data (wherein a single-variable is measured at each time point), we are not aware of any methods that tackle this problem in a multivariate setting which is the case with periodic image sequences wherein several thousands of variables (intensities of pixels) could be measured at each time point.
+Our strategy was to find a way to simplify this seemingly complex multi-variate problem by transforming it into a univariate one and take advantage of existing methods to solve the problem at hand.
+To accomplish this, we compute the similarity/affinity between all pairs of frames in the given periodic image sequence containing $n$ images to create a matrix $A \in R^{n \times n}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames of the given sequence.
+In principle, any of the similarity metrics used in image registration algorithms can be chosen to quantify the similarity between a given pair of frames.
+Here in, we use normalization correlation also known as the pearson correlation coefficient.
+Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series that measures the similarity of the corresponding frame in the image sequence with all other frames.
+Additionally, for a periodic image sequence, we can expect them to be periodic provided that the corresponding frame is not significantly corrupted by distortions such as noise and changes in global illumination over time among others.
+To mitigate the potential problem of global illumintation changes over time, we detrend and extract the seasonal part of the time series corresponding to each row of $A$ by dividing it with the trend signal estimated using a locally-weighted regression algorithm called loess\cite{Cleveland1990}.
+Let $A_s$ be the matrix whose rows contain the seasonal part obtained by detrending the corresponding row in $A$.
+Next, we minimize the risk from noise, by finding the row $u(t)$ of $A_s$ whose power spectrum or periodogram has minimum entropy.
+In our experiments, the time series $u(t)$ was nearly sinusoidal in morphology with most of its energy in its power spectrum concentrated within a very narrow-band with the same periodicity characteristics as the image sequence it was derived from.
+Based on this observation, we decided to use its Hilbert transform \cite{Lu2013} to estimate the instantaneous intra-period phase  of each frame.
+Specifically, we compute the instantaneous phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the time series $u(t)$ as follows:
+\begin{equation}
+\phi(t) = arctan \left( \frac{H_u(t)}{u(t)}\right)
+\end{equation} 
+where $H_u(t)$ is the Hilbert transform of $u(t)$.
+Lastly, we map the instantaneous phase $\phi(t)$ to the range $\left [  0, 1\right )$ which can then be used to obtain the intra-period phase of each frame.
 %
-\vspace{-0.5cm}
+\begin{IonFigT}
+\centering
+%
+\includegraphics[width=1.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat.png}
+\includegraphics[width=1.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat_Trend.png}
+\includegraphics[width=1.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat_Seasonal.png}
+\ionbox{1.5in}\ionbox{1.5in}\ionbox{1.5in}\\
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/season_trend_decomposition.png}
+\ionbox{4.5in}
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instantaneous_phase.png}
+\ionbox{4.5in}
+%
+\caption{Illustration of the proposed phase estimation method}
+\label{fig:phase_estimation}
+\end{IonFigT}
+%\vspace{-0.5cm}
 \subsection{Gating of respiratory frames}
@@
 %
-\vspace{-0.5cm}
+%\vspace{-0.5cm}
 \subsection{Reconstruction of one-period video at high temporal resolution}
-\label{sec:method:gating}
+\label{sec:method:super_resolution}
 %
 
-\vspace{-0.5cm}
+%\vspace{-0.5cm}
 \section{Results}
@@
 %
-\vspace{-0.5cm}
+%\vspace{-0.5cm}
 \section{Discussion and future work}
@@
 %
-\vspace{-0.5cm}
+%\vspace{-0.5cm}
 \bibliographystyle{splncs03}

### 2017-02-16 paper155.tex
Updated figure properties, combined simMat images, combined phase ordered images

@@
 %
+%
+\begin{IonFigT}
+\centering
+%
+\includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat.png}
+\ionbox{4.75in}\\
+\includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/season_trend_decomposition.png}
+\ionbox{4.75in}\\
+\includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/periodogram.png}
+\ionbox{4.75in}\\
+\includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instantaneous_phase.png}
+\ionbox{4.75in}\\
+%
+\caption{Illustration of the proposed phase estimation method}
+\label{fig:phase_estimation}
+\end{IonFigT}
+%
 While there have been numerous efforts for the estimation of instantaneous intra-period phase\cite{Lu2013} in periodic univariate time series data (wherein a single-variable is measured at each time point), we are not aware of any methods that tackle this problem in a multivariate setting which is the case with periodic image sequences wherein several thousands of variables (intensities of pixels) could be measured at each time point.
@@
 Lastly, we map the instantaneous phase $\phi(t)$ to the range $\left [  0, 1\right )$ which can then be used to obtain the intra-period phase of each frame.
-%
-\begin{IonFigT}
-\centering
-%
-\includegraphics[width=1.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat.png}
-\includegraphics[width=1.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat_Trend.png}
-\includegraphics[width=1.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat_Seasonal.png}
-\ionbox{1.5in}\ionbox{1.5in}\ionbox{1.5in}\\
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/season_trend_decomposition.png}
-\ionbox{4.5in}
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instantaneous_phase.png}
-\ionbox{4.5in}
-%
-\caption{Illustration of the proposed phase estimation method}
-\label{fig:phase_estimation}
-\end{IonFigT}
+
 %\vspace{-0.5cm}
@@
 
+%
+\begin{IonFigT}
+\centering
+%
+\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/respiratory_phase_gating.png}
+\ionbox{5.0in}\\
+\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/robust_lowess_gating.png}
+\ionbox{5.0in}\\
+%
+\caption{Illustration of our method for gating respiratory frames}
+\label{fig:respiratory_gating}
+\end{IonFigT}
 %
@@
 %
-
+%
+\begin{IonFigT}
+\centering
+%
+\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/phaseordered.png}
+\ionbox{5.0in}\\
+%
+\caption{Visualization of video frames are ordered by estimated cardiac phase}
+\label{fig:phase_ordering}
+\end{IonFigT}
+%
 %

### 2017-02-16 paper155.tex
Updated paper text

@@
 In this section, we present the theory underlying the proposed methods for instantaneous phase estimation, gating and temporal super-resolution of cardiac ultrasound videos.
-Specifically, we begin by describing our method for the estimation of instantaneous cardiac and respiratory phases in
-Section \ref{sec:method:phase_estimation}.
-We then present a method that uses these phase estimates to robustly gate video frames predominantly influenced by respiratory motion in Section \ref{sec:method:gating}.
+Specifically, we begin by describing our method for the estimation of instantaneous cardiac and respiratory phases in Section \ref{sec:method:phase_estimation}.
+We then present a robust method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion in Section \ref{sec:method:gating}.
 Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression method that uses the phase-tagged image sequence to reconstruct a single-period video at higher-temporal resolution.
@@
 %
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/ecg_instaphase_overlay.png}\\
-\ionbox{4.5in}\\
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/qrs_peak_to_peak.png}
-\ionbox{4.5in}\\
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instaphase_valley_to_valley.png}
-\ionbox{4.5in}\\
+\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/ecg_instaphase_overlay.png}\\
+\ionbox{5.0in}\\
+\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/qrs_peak_to_peak.png}
+\ionbox{5.0in}\\
+\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instaphase_valley_to_valley.png}
+\ionbox{5.0in}\\
 %
@@
 %\vspace{-0.5cm}
-\subsection{Instantaneous phase estimation of each frame}
+\subsection{Estimation of instantaneous cardiac and respiratory phases}
 \label{sec:method:phase_estimation}
-%
 %
@@
 %
-\caption{Illustration of the proposed phase estimation method}
+\caption{Illustration of the proposed phase estimation method: (a-left) Inter-frame similarity matrix, (a-mid) Trend component corresponding to the respiratory motion, (a-right) Seasonal component corresponding to the beating heart motion, (b) Similarity profile with minimum spectral entropy along with the associated trend/respiration and seasonal/heart-beat components, (c) Periodogram of the heart-beat and respiration components, and (d) Instantaneous phase of heart-beat and respiration components.}
 \label{fig:phase_estimation}
@@
 %
-While there have been numerous efforts for the estimation of instantaneous intra-period phase\cite{Lu2013} in periodic univariate time series data (wherein a single-variable is measured at each time point), we are not aware of any methods that tackle this problem in a multivariate setting which is the case with periodic image sequences wherein several thousands of variables (intensities of pixels) could be measured at each time point.
-Our strategy was to find a way to simplify this seemingly complex multi-variate problem by transforming it into a univariate one and take advantage of existing methods to solve the problem at hand.
+While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data (wherein a single-variable is measured at each time point)\cite{Boashash1992,Lu2013,Luo2003}, we are not aware of any methods that tackle this problem in a multivariate setting which is the case with cardiac ultrasound videos wherein several thousands of variables (intensities of pixels) could be measured at each time point.
+Our strategy was to find a way to simplify this seemingly complex multi-variate problem by transforming it into a univariate one and take advantage of existing univariate methods to solve the problem at hand.
 To accomplish this, we compute the similarity/affinity between all pairs of frames in the given periodic image sequence containing $n$ images to create a matrix $A \in R^{n \times n}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames of the given sequence.
-In principle, any of the similarity metrics used in image registration algorithms can be chosen to quantify the similarity between a given pair of frames.
-Here in, we use normalization correlation also known as the pearson correlation coefficient.
+Here in, we use normalization correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity but, in principle, any of the similarity metrics used in image registration algorithms \cite{Goshtasby2012} can potentially be chosen.
 Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series that measures the similarity of the corresponding frame in the image sequence with all other frames.
 Additionally, for a periodic image sequence, we can expect them to be periodic provided that the corresponding frame is not significantly corrupted by distortions such as noise and changes in global illumination over time among others.
+
 To mitigate the potential problem of global illumintation changes over time, we detrend and extract the seasonal part of the time series corresponding to each row of $A$ by dividing it with the trend signal estimated using a locally-weighted regression algorithm called loess\cite{Cleveland1990}.
@@
 %\vspace{-0.5cm}
-\subsection{Gating of respiratory frames}
+\subsection{Gating out respiratory frames}
 \label{sec:method:gating}
 %
-
+Once the instantaneous cardiac and respiratory phases of each frame has been estimated, it can be used to facilitate the extraction of quantitative measurements from a desired part/point of the periodic cycle, a process commonly referred to as gating.
+In this section, we present a robust method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.
 %
@@
 %
-\caption{Illustration of our method for gating respiratory frames}
+\caption{Illustration of our method for gating out respiratory frames: (a) (b)}
 \label{fig:respiratory_gating}
@@
 %
+Once the instantaneous cardiac phase of each frame has been estimated, it can be used to enable spatio-temporal super-resolution by considering the set of images within each period of a periodic image sequence as distinct observations of the same data.
+In this section, we present a method that uses a phase-tagged periodic image sequence to reconstruct a single-period video at a higher temporal resolution.
 
+Given a de-noised periodic sequence $\{L_1, ..., L_n\}$ of $n$ images (with $m$ pixels each) along with their estimated intra-period phases $\{\phi_1, ..., \phi_n\}$, we first use Nadarya-Watson kernel regression\cite{Bishop2006} to construct a phase-parameterized image manifold in the form of a function $M(\phi): [0, 1) \to R^m $ that maps any intra-period phase $\phi$ to an image as follows:
+\begin{equation}
+M(\phi) = \frac{\sum_{i = 1}^{n} K(\phi, \phi_i) L_i}{\sum_{i = 1}^{n} K(\phi, \phi_i)} 
+\end{equation}
+where in we define the kernel $K$ using a gausssian with standard-deviation $\sigma$ equal to a multiple of the average distance between consecutive phases after sorting them.
+Note that, we account for the periodic nature of the intra-period phase whenever we measure distance between two phase values.
+A single-period video can now be reconstructed by sampling the manifold $M(\phi)$ at any desired resolution from the range $[0, 1)$ of intra-period phase.
+Note that the aformentioned process will facilitate temporal super-resolution only if the set of images observed in each distinct period of the given image sequence differ in intra-period phase by some non-zero amount less than its temporal sampling interval.
+Also, the amount of temporal super-resolution possible will depend both on the number of periods observed and the maximum amount by which one can expect the images to be perturbed in phase across observations.
 %\vspace{-0.5cm}
@@
 %
-\caption{Visualization of video frames are ordered by estimated cardiac phase}
+\caption{Visualization of m-mode frames ordered by estimated cardiac phase}
 \label{fig:phase_ordering}

### 2017-02-17 paper155.tex
Updated paper - finished phase estimation section

@@
 In an effort to address this problem, we present a novel method for de-noising, gating, and temporal super-resolution of noisy low frame-rate periodic image sequences such as the ones obtained using low-cost hand-held cardiac ultrasound devices.
-We first leverage the inherent image-level redundancy in a periodic image sequence to decouple signal from noise using a low-rank + sparse decomposition algorithm.
-Next, we transform the complex high-dimensional periodic image sequence to a narrow-band 1D time series with the same periodicity characteristics using a novel strategy based on inter-frame similarity.
-Next, we decompose this time series use a season-trend decomposition method to obtain the heart-beat (season) and respiratory (trend) signals.
-Next, we use the Hilbert transform to estimate the instantaneous cardiac and respiratory phases of each frame to facilitate gating and super-resolution.
-Lastly, we use Nadarya-Watson kernel regression to construct an image manifold parameterized by intra-period phase which can then be sampled to reconstruct a de-noised single-period image sequence at a higher temporal-resolution.
-We demonstrate the utility and efficacy of our method both on synthetic data and on real image sequences of the beating heart acquired using a low-cost hand-held ultrasound device.
+
+We first transform the complex high-dimensional periodic cardiac ultrasound image sequence to a narrow-band 1D time series with the same periodicity characteristics using a novel strategy based on inter-frame similarity.
+Next, we decompose this time series use a trend extraction technique to obtain the trend/respiration and residual/heart-beat component signals.
+Next, we use the Hilbert transform to estimate the instantaneous cardiac and respiratory phases of each frame to facilitate gating and temporal super-resolution.
+Lastly, we use kernel regression with a novel kernel to learn an image manifold parametrized by instantaneous cardiac phase that can be sampled to generate a single-cycle video at a higher temporal-resolution.
+We demonstrate and validate our methods using cardiac ultrasound videos and associated ECG recordings of 6 mice.
 %\keywords{Cardiac, Ultrasound, Phase estimation, Gating, Temporal super-resolution}
@@
 \includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat.png}
-\ionbox{4.75in}\\
+\ionbox{1.6in}\ionbox{1.6in}\ionbox{1.6in}\\
 \includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/season_trend_decomposition.png}
@@
 %
-\caption{Illustration of the proposed phase estimation method: (a-left) Inter-frame similarity matrix, (a-mid) Trend component corresponding to the respiratory motion, (a-right) Seasonal component corresponding to the beating heart motion, (b) Similarity profile with minimum spectral entropy along with the associated trend/respiration and seasonal/heart-beat components, (c) Periodogram of the heart-beat and respiration components, and (d) Instantaneous phase of heart-beat and respiration components.}
+\caption{Illustration of the proposed phase estimation method: (a) Inter-frame similarity matrix, (b) Trend component corresponding to the respiratory motion, (c) Seasonal component corresponding to the beating heart motion, (d) The similarity profile chosen for phase estimation along with the associated trend/respiration and seasonal/heart-beat components, (e) Periodogram of the heart-beat and respiration components along with the periodicity characteristics such as frequency and cycle duration, and (f) Instantaneous phase of the heart-beat and respiration components.}
 \label{fig:phase_estimation}
@@
 %
-While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data (wherein a single-variable is measured at each time point)\cite{Boashash1992,Lu2013,Luo2003}, we are not aware of any methods that tackle this problem in a multivariate setting which is the case with cardiac ultrasound videos wherein several thousands of variables (intensities of pixels) could be measured at each time point.
+While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data (wherein a single-variable is measured at each time point)\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting which is the case with periodic cardiac ultrasound videos wherein several thousands of variables (intensities of pixels) could be measured at each time point.
 Our strategy was to find a way to simplify this seemingly complex multi-variate problem by transforming it into a univariate one and take advantage of existing univariate methods to solve the problem at hand.
-To accomplish this, we compute the similarity/affinity between all pairs of frames in the given periodic image sequence containing $n$ images to create a matrix $A \in R^{n \times n}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames of the given sequence.
+
+	We first compute the similarity between all pairs of frames in the given periodic image sequence containing $N$ images/frames to create a matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames of the given sequence.
 Here in, we use normalization correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity but, in principle, any of the similarity metrics used in image registration algorithms \cite{Goshtasby2012} can potentially be chosen.
 Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series that measures the similarity of the corresponding frame in the image sequence with all other frames.
-Additionally, for a periodic image sequence, we can expect them to be periodic provided that the corresponding frame is not significantly corrupted by distortions such as noise and changes in global illumination over time among others.
+If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted we can expect this time series to encompass the
+predominant periodicity characteristics of the original image sequence.
+Two kinds of periodic motion are present in cardiac ultrasound videos, one corresponding to the beating heart motion and the other corresponding to respiratory motion.
+Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos computed using the normalized correlation metric wherein these two sources of periodicity can be clearly observed in the frame similarity signal of many rows/columns.
 
-To mitigate the potential problem of global illumintation changes over time, we detrend and extract the seasonal part of the time series corresponding to each row of $A$ by dividing it with the trend signal estimated using a locally-weighted regression algorithm called loess\cite{Cleveland1990}.
-Let $A_s$ be the matrix whose rows contain the seasonal part obtained by detrending the corresponding row in $A$.
-Next, we minimize the risk from noise, by finding the row $u(t)$ of $A_s$ whose power spectrum or periodogram has minimum entropy.
-In our experiments, the time series $u(t)$ was nearly sinusoidal in morphology with most of its energy in its power spectrum concentrated within a very narrow-band with the same periodicity characteristics as the image sequence it was derived from.
-Based on this observation, we decided to use its Hilbert transform \cite{Lu2013} to estimate the instantaneous intra-period phase  of each frame.
+	Next, we use a trend extraction technique~\cite{Alexandrov2012} called the Hodrick-Prescott (HP) filter~\cite{Hodrick1997} to decompose the frame similarity signal $u(t)$ into a sum of two components: (i) Trend component $\tau(t)$ encompassing the periodicity characteristic of only the lower-frequency respiratory motion, and (ii) Residual component $r(t)$ encompassing the periodicity characteristic of only the higher-frequency beating heart motion.
+The HP filter, that is widely used in the field of econometrics, performs the decomposition of $u(t) = \tau(t) + r(t)$ by solving the following optimization problem:
+\begin{equation}	
+\argmin{\tau(t)} \left[ \sum_{t=1}^{N}  \left(u(t) - \tau(t) \right)^2  + \lambda \sum_{t=1}^{N-1} \left( \nabla^2 \tau(t) \right)^2 \right] 
+\end{equation}
+where $\nabla^2\tau(t) = \tau(t+1) - 2 \tau(t) + \tau(t-1)$ is the second-order difference or derivative that penalizes curvature of the trend signal.
+We use a ridge-regression implementation of the Hodrick-prescott filter~\cite{Yamada2015} with $\lambda=6400$.
+Let $A_r$ and $A_h$ be the matrices whose rows contain the trend/respiratory and the residual/heart-beat components, respectively, of the frame similarity signals in the corresponding rows of the similarity matrix $A$.
+Figures~\ref{fig:phase_estimation}(b,c) show the trend ($A_r$) and residual ($A_h$) matrices for one of our cardiac ultrasound videos.
+To minimize the effect of any noise on phase estimation, we pick frame similarity signal corresponding to the row of $A_h$ whose periodogram or power-frequency distribution has minimum entropy.
+Figure~\ref{fig:phase_estimation}(d) shows the selected frame similarity signal (blue) along with the associated trend/respiration (pink) and residual/heart-beat (green) components.
+Figure~\ref{fig:phase_estimation}(e) shows the periodogram of these two components.
+Through Figures~\ref{fig:phase_estimation}(b-e), notice how well the two periodic sources have been decoupled from the raw frame similarity signal.
+Also, in Figure~\ref{fig:phase_estimation}, notice that most of the energy in the periodogram or power-frequency distribution of the respiration and heart-beat components is concentrated within a narrow frequency band.
+Furthermore, notice how these 1D signals capture the periodicity characteristics (e.g.
+frequency in beats-per-minute (bpm), period or cycle duration in frames, number of cycles) of the respective sources of motion in the input high-dimensional periodic image sequence.
+
+	Next, considering the narrow-band nature of the derived trend/respiratory and residual/heart-beat signals, we can use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous intra-period phase of each frame.
 Specifically, we compute the instantaneous phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the time series $u(t)$ as follows:
@@
 where $H_u(t)$ is the Hilbert transform of $u(t)$.
-Lastly, we map the instantaneous phase $\phi(t)$ to the range $\left [  0, 1\right )$ which can then be used to obtain the intra-period phase of each frame.
+We map $\phi(t)$ from range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
+Let $\phi_c(t)$ and $\phi_r(t)$ denote the instantaneous heart-beat/cardiac and respiratory phases computed from the trend/respiration and residual/heart-beat signals, respectively, using the aforementioned approach.
+Figure~\ref{fig:phase_estimation}(f) shows the instantaneous cardiac and respiratory phases computed from the heart-beat and respiration signals in in Figure~\ref{fig:phase_estimation}(d).
 
@@
 %
-Once the instantaneous cardiac and respiratory phases of each frame has been estimated, it can be used to facilitate the extraction of quantitative measurements from a desired part/point of the periodic cycle, a process commonly referred to as gating.
+Once the instantaneous cardiac and respiratory phases of each frame has been estimated, it can be used to facilitate the extraction of quantitative measurements from any desired part/point of the periodic cycle, a process commonly referred to as gating.
 In this section, we present a robust method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.
+
+	Firstly, based on the observation that the predominant influence of respiratory motion occurs around the troughs of the trend/respiration component (pink curve in Fig.~\ref{fig:phase_estimation}(d)), we identify the corresponding respiratory phase (equal to phase zero) and perform a rough initial gating by discarding the frames whose respiratory phase distance from it is below a specified cutoff value (we use 0.2).
+Figure~\ref{fig:respiratory_gating}(a) shows the discarded frames (blue circles) overlaid on the trend/respiration component, residual/heart-beat component, and instantaneous respiratory phase signals.
+
 %
@@
 %\vspace{-0.5cm}
-\subsection{Reconstruction of one-period video at high temporal resolution}
+\subsection{Build model to generate image at any phase for super-resolution}
 \label{sec:method:super_resolution}

### 2017-02-20 paper155.tex
Finished gating section and edited abstract

@@
 \newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\operatorname{min}}\;}
+%
 \mathchardef\mhyphen="2D
@@
 \begin{abstract}
-Point-of-care low-cost hand-held ultrasound devices are emerging as an increasingly effective tool for rapid non-invasive visual assessment of cardiac structure and function on the bed-side.
-Their diagnostic value can be enriched further by coupling them with automated image analysis algorithms to accurately detect and quantify any cardiac abnormalities at the point-of-care potentially even in asymptomatic patients.
-However, a major challenge to this is the poor spatio-temporal resolution and high noise level of images obtained using these devices.
-In an effort to address this problem, we present a novel method for de-noising, gating, and temporal super-resolution of noisy low frame-rate periodic image sequences such as the ones obtained using low-cost hand-held cardiac ultrasound devices.
-
-We first transform the complex high-dimensional periodic cardiac ultrasound image sequence to a narrow-band 1D time series with the same periodicity characteristics using a novel strategy based on inter-frame similarity.
-Next, we decompose this time series use a trend extraction technique to obtain the trend/respiration and residual/heart-beat component signals.
-Next, we use the Hilbert transform to estimate the instantaneous cardiac and respiratory phases of each frame to facilitate gating and temporal super-resolution.
-Lastly, we use kernel regression with a novel kernel to learn an image manifold parametrized by instantaneous cardiac phase that can be sampled to generate a single-cycle video at a higher temporal-resolution.
+In this paper, we present a novel image-based method for the estimation of instantaneous cardiac and respiratory phases from cardiac ultrasound videos and demonstrate its utility in gating and temporal super-resolution.
+We first use a novel strategy based on inter-frame similarity to transform the complex high-dimensional periodic cardiac ultrasound image sequence into a 1D time series with the same periodicity characteristics.
+Next, we use a trend extraction technique to decompose this time series into a  trend component encoding respiratory motion alone and a residual component encoding the beating heart motion along.
+Next, we use the Hilbert transform to estimate the instantaneous cardiac and respiratory phases of each frame from these signals.
+Next, we present a robust non-parametric regression technique that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.
+Lastly, we use a kernel regression with a novel kernel to learn an image manifold parameterized by instantaneous cardiac phase that can be sampled to generate a single-cycle video at a higher temporal-resolution.
 We demonstrate and validate our methods using cardiac ultrasound videos and associated ECG recordings of 6 mice.
@@
 
-\vspace{-0.5cm}
+%\vspace{-0.5cm}
 \section{Introduction}
@@
 %
-
+Ultrasound is emerging as an increasingly effective tool for rapid non-invasive visual assessment of cardiac structure and function. 
 %
@@
 Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series that measures the similarity of the corresponding frame in the image sequence with all other frames.
-If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted we can expect this time series to encompass the
-predominant periodicity characteristics of the original image sequence.
+If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted we can expect this time series to encompass the predominant periodicity characteristics of the original image sequence.
 Two kinds of periodic motion are present in cardiac ultrasound videos, one corresponding to the beating heart motion and the other corresponding to respiratory motion.
@@
 
-	Next, we use a trend extraction technique~\cite{Alexandrov2012} called the Hodrick-Prescott (HP) filter~\cite{Hodrick1997} to decompose the frame similarity signal $u(t)$ into a sum of two components: (i) Trend component $\tau(t)$ encompassing the periodicity characteristic of only the lower-frequency respiratory motion, and (ii) Residual component $r(t)$ encompassing the periodicity characteristic of only the higher-frequency beating heart motion.
-The HP filter, that is widely used in the field of econometrics, performs the decomposition of $u(t) = \tau(t) + r(t)$ by solving the following optimization problem:
+	Next, we use a trend extraction technique~\cite{Alexandrov2012} called the Hodrick-Prescott (HP) filter~\cite{Hodrick1997}, widely used in econometrics, to decompose the frame similarity signal $u(t)$ corresponding to each row of $A$ into a sum of two components: (i) Trend component $\tau_{resp}(t)$ encompassing the periodicity characteristic of only the lower-frequency respiratory motion, and (ii) Residual component $r_{heart}(t)$ encompassing the periodicity characteristic of only the higher-frequency beating heart motion.
+The HP filter performs the decomposition of $u(t) = \tau_{resp}(t) + r_{heart}(t)$ by solving the following optimization problem:
 \begin{equation}	
-\argmin{\tau(t)} \left[ \sum_{t=1}^{N}  \left(u(t) - \tau(t) \right)^2  + \lambda \sum_{t=1}^{N-1} \left( \nabla^2 \tau(t) \right)^2 \right] 
+\argmin{\tau(t)} \left[ \sum_{t=1}^{N}  \left(u(t) - \tau_{resp}(t) \right)^2  + \lambda \sum_{t=1}^{N-1} \left( \nabla^2 \tau_{resp}(t) \right)^2 \right] 
 \end{equation}
-where $\nabla^2\tau(t) = \tau(t+1) - 2 \tau(t) + \tau(t-1)$ is the second-order difference or derivative that penalizes curvature of the trend signal.
-We use a ridge-regression implementation of the Hodrick-prescott filter~\cite{Yamada2015} with $\lambda=6400$.
-Let $A_r$ and $A_h$ be the matrices whose rows contain the trend/respiratory and the residual/heart-beat components, respectively, of the frame similarity signals in the corresponding rows of the similarity matrix $A$.
-Figures~\ref{fig:phase_estimation}(b,c) show the trend ($A_r$) and residual ($A_h$) matrices for one of our cardiac ultrasound videos.
-To minimize the effect of any noise on phase estimation, we pick frame similarity signal corresponding to the row of $A_h$ whose periodogram or power-frequency distribution has minimum entropy.
+where $\nabla^2\tau_{resp}(t) = \tau_{resp}(t+1) - 2 \tau_{resp}(t) + \tau_{resp}(t-1)$ is the second-order difference or derivative that penalizes curvature of the trend signal.
+We use a ridge-regression implementation of the HP filter~\cite{Yamada2015} with $\lambda=6400$.
+Let $A_{resp}$ and $A_{heart}$ be the matrices whose rows contain the trend/respiratory and residual/heart-beat components, respectively, of the frame similarity signals in the corresponding rows of the similarity matrix $A$.
+Figures~\ref{fig:phase_estimation}(b,c) show the trend $A_{res[}$ and residual $A_{heart}$ matrices for one of our cardiac ultrasound videos.
+To minimize the effect of any noise on phase estimation, we pick frame similarity signal corresponding to the row of $A_{heart}$ whose periodogram or power-frequency distribution has minimum entropy.
+Let $\hat{u}(t)$ be the chosen frame similarity signal and let $\hat{\tau}_{resp}(t)$ and $\hat{r}_{heart}(t)$  be its associated trend/respiration and residual/heart-beat component signals, respectively.
 Figure~\ref{fig:phase_estimation}(d) shows the selected frame similarity signal (blue) along with the associated trend/respiration (pink) and residual/heart-beat (green) components.
@@
 
-	Next, considering the narrow-band nature of the derived trend/respiratory and residual/heart-beat signals, we can use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous intra-period phase of each frame.
-Specifically, we compute the instantaneous phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the time series $u(t)$ as follows:
+	Next, considering the narrow-band nature of the derived trend/respiratory and residual/heart-beat signals, we use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous phase of each frame.
+Specifically, we compute the instantaneous intra-period phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows:
 \begin{equation}
-\phi(t) = arctan \left( \frac{H_u(t)}{u(t)}\right)
+\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)
 \end{equation} 
-where $H_u(t)$ is the Hilbert transform of $u(t)$.
-We map $\phi(t)$ from range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
-Let $\phi_c(t)$ and $\phi_r(t)$ denote the instantaneous heart-beat/cardiac and respiratory phases computed from the trend/respiration and residual/heart-beat signals, respectively, using the aforementioned approach.
+and map $\phi(t)$ from the range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
+Let $\hat{\phi}_{heart}(t)$ and $\hat{\phi}_{resp}(t)$ denote the instantaneous cardiac and respiratory phases computed from the trend/respiration $\hat{\tau}_{resp}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signals, respectively, using the aforementioned approach.
 Figure~\ref{fig:phase_estimation}(f) shows the instantaneous cardiac and respiratory phases computed from the heart-beat and respiration signals in in Figure~\ref{fig:phase_estimation}(d).
@@
 %
-Once the instantaneous cardiac and respiratory phases of each frame has been estimated, it can be used to facilitate the extraction of quantitative measurements from any desired part/point of the periodic cycle, a process commonly referred to as gating.
-In this section, we present a robust method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.
+Once the instantaneous cardiac and respiratory phases of each frame has been estimated, it can be used to facilitate the extraction of quantitative measurements from a desired part/point of the periodic cycle, a process commonly referred to as gating.
+In this section, we present a robust two-step method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.
 
-	Firstly, based on the observation that the predominant influence of respiratory motion occurs around the troughs of the trend/respiration component (pink curve in Fig.~\ref{fig:phase_estimation}(d)), we identify the corresponding respiratory phase (equal to phase zero) and perform a rough initial gating by discarding the frames whose respiratory phase distance from it is below a specified cutoff value (we use 0.2).
-Figure~\ref{fig:respiratory_gating}(a) shows the discarded frames (blue circles) overlaid on the trend/respiration component, residual/heart-beat component, and instantaneous respiratory phase signals.
+	In the first step, based on the observation that the predominant influence of respiratory motion occurs around the troughs/minima of the trend/respiration signal $\hat{\tau}_{resp}(t)$ (see pink curve in Fig.~\ref{fig:phase_estimation}(d)) corresponding to the respiratory phase $\hat{\phi}_{resp}(t) = 0$, we perform a rough initial gating by discarding the frames $F_{cutoff} = \left \{ t \mid \hat{\phi}_{resp}(t) < c \parallel \hat{\phi}_{resp}(t) > (1 - c) \right \}$ whose respiratory phase distance from $\hat{\phi}_{resp}(t) = 0$ is below a specified cutoff value $c$ that we set to 0.2.
+Figure~\ref{fig:respiratory_gating}(a) shows the discarded frames $F_{cutoff}$ overlaid (blue circles) on the trend/respiration $\hat{\tau}_{resp}(t)$, residual/heart-beat  $\hat{r}_{heart}(t)$, and respiratory phase $\hat{\phi}_{resp}(t)$ signals.
 
+	In the second step, we learn a robust mapping $S(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the residual/heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called Locally weighted regression (LOESS) to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signal values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above.
+Next, we compute the median absolution value $\gamma_{MAD}$ of the difference between the LOESS fit $S( \hat{\phi}_{heart}(t) )$ and residual/heart-beat signal $\hat{r}_{heart}(t)$ for all frames as a robust estimate of the dispersion of non-respiratory frames around the LOESS fit.
+Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert S( \hat{\phi}_{heart}(t) ) - \hat{r}_{heart}(t)  \rvert   > k \times \gamma_{mad}  \right \}$ for which the difference between the LOESS fit and heart-beat signal value is greater than $k \times \gamma_{mad}$ (we set $k = 2.0$).
 %
@@
 %
-\caption{Illustration of our method for gating out respiratory frames: (a) (b)}
+\caption{Illustration of our method for gating out respiratory frames: (a) Frames $F_{cutoff}$ discarded (blue circles) in the first step overlaid with the trend/respiration, residual/heart-beat, and instantaneous respiratory phase signals (b) }
 \label{fig:respiratory_gating}
@@
 %\vspace{-0.5cm}
-\subsection{Build model to generate image at any phase for super-resolution}
+\subsection{Learning to reconstruct image at any cardiac phase}
 \label{sec:method:super_resolution}
@@
 \end{equation}
-where in we define the kernel $K$ using a gausssian with standard-deviation $\sigma$ equal to a multiple of the average distance between consecutive phases after sorting them.
+where in we define the kernel $K$ using a gaussian with standard-deviation $\sigma$ equal to a multiple of the average distance between consecutive phases after sorting them.
 Note that, we account for the periodic nature of the intra-period phase whenever we measure distance between two phase values.

### 2017-02-21 paper155.tex
Updated paper and figures -- finished kernel regression section

@@
 \begin{abstract}
-In this paper, we present a novel image-based method for the estimation of instantaneous cardiac and respiratory phases from cardiac ultrasound videos and demonstrate its utility in gating and temporal super-resolution.
-We first use a novel strategy based on inter-frame similarity to transform the complex high-dimensional periodic cardiac ultrasound image sequence into a 1D time series with the same periodicity characteristics.
-Next, we use a trend extraction technique to decompose this time series into a  trend component encoding respiratory motion alone and a residual component encoding the beating heart motion along.
-Next, we use the Hilbert transform to estimate the instantaneous cardiac and respiratory phases of each frame from these signals.
+In this paper, we present a novel purely image-based method for the estimation of instantaneous cardiac and respiratory phases from cardiac ultrasound videos and demonstrate its utility in gating and temporal super-resolution.
+We first use a novel strategy based on inter-frame similarity to transform the complex high-dimensional image sequence into a 1D time series with the same periodicity characteristics.
+Next, we use a trend extraction technique to decompose this time series into a trend component encoding respiratory motion alone and a residual component encoding the beating heart motion alone.
+Next, we use the Hilbert transform to estimate the instantaneous cardiac and respiratory phases of each frame from these components.
 Next, we present a robust non-parametric regression technique that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.
@@
 %
-\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/ecg_instaphase_overlay.png}\\
-\ionbox{5.0in}\\
-\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/qrs_peak_to_peak.png}
-\ionbox{5.0in}\\
-\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instaphase_valley_to_valley.png}
-\ionbox{5.0in}\\
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/ecg_instaphase_overlay.png}\\
+\ionbox{4.5in}\\
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/qrs_peak_to_peak.png}\\
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instaphase_valley_to_valley.png}
+\ionbox{4.5in}\\
 %
-\caption{Illustration of correspondence between ECG and the instantaneous cardiac phase estimated using the proposed method: (a) ECG signal (blue) and the cardiac phase estimated using our method using only image data.
-Also shown are the peaks of the QRS complex (pink circle) and the minima of the estimated cardiac phase (black circle) that correspond to the QRS peaks, (b) Five video frames evenly spaced in time between the first and second QRS peaks of the ECG signal constituting one cardiac cycle, and (c) Five video frames evenly spaced in time between first and second minima of the estimated cardiac phase constituting one cardiac cycle.}
+\caption{Illustration of correspondence between ECG and the instantaneous cardiac phase estimated using the proposed method: (a) ECG signal (blue) and the cardiac phase estimated using our method overlaid with peaks of the QRS complex (red circle) and the corresponding cardiac phase minima (black circle), (b) Five video frames evenly spaced in time between the first and second QRS peaks of the ECG signal (top-row) and the corresponding cardiac phase minima (bottom-row) constituting one cardiac cycle.}
 \label{fig:instaphase_vs_ecg}
@@
 %
-\includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat.png}
-\ionbox{1.6in}\ionbox{1.6in}\ionbox{1.6in}\\
-\includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/season_trend_decomposition.png}
-\ionbox{4.75in}\\
-\includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/periodogram.png}
-\ionbox{4.75in}\\
-\includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instantaneous_phase.png}
-\ionbox{4.75in}\\
+\includegraphics[width=4.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat.png}
+\ionbox{1.33in}\ionbox{1.33in}\ionbox{1.33in}\\
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/season_trend_decomposition.png}
+\ionbox{4.5in}\\
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/periodogram.png}
+\ionbox{4.5in}\\
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instantaneous_phase.png}
+\ionbox{4.5in}\\
 %
-\caption{Illustration of the proposed phase estimation method: (a) Inter-frame similarity matrix, (b) Trend component corresponding to the respiratory motion, (c) Seasonal component corresponding to the beating heart motion, (d) The similarity profile chosen for phase estimation along with the associated trend/respiration and seasonal/heart-beat components, (e) Periodogram of the heart-beat and respiration components along with the periodicity characteristics such as frequency and cycle duration, and (f) Instantaneous phase of the heart-beat and respiration components.}
+\caption{Illustration of our phase estimation method: (a) Inter-frame similarity matrix, (b) Trend component matrix corresponding to respiratory motion, (c) Residual component matrix corresponding to beating heart motion, (d) Similarity profile chosen for phase estimation with associated trend/respiration and residual/heart-beat components, (e) Periodogram of heart-beat and respiration components along with periodicity characteristics (e.g.
+frequency, cycle duration) calculated for the peak frequency, and (f) Instantaneous cardiac and respiratory phases estimated using the Hilbert transform.}
 \label{fig:phase_estimation}
@@
 %
-While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data (wherein a single-variable is measured at each time point)\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting which is the case with periodic cardiac ultrasound videos wherein several thousands of variables (intensities of pixels) could be measured at each time point.
-Our strategy was to find a way to simplify this seemingly complex multi-variate problem by transforming it into a univariate one and take advantage of existing univariate methods to solve the problem at hand.
+While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data (wherein a single-variable is measured at each time point)\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting which is the case with periodic cardiac ultrasound videos wherein several thousands of variables (pixel intensities) could be measured at each time point.
+Our strategy was to find a way to simplify this complex multi-variate problem by transforming it into a univariate one and take advantage of existing univariate methods to solve the problem.
 
@@
 Two kinds of periodic motion are present in cardiac ultrasound videos, one corresponding to the beating heart motion and the other corresponding to respiratory motion.
-Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos computed using the normalized correlation metric wherein these two sources of periodicity can be clearly observed in the frame similarity signal of many rows/columns.
+Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos computed using the normalized correlation metric wherein these two sources of periodicity can be clearly observed in many rows/columns.
 
@@
 Figure~\ref{fig:phase_estimation}(d) shows the selected frame similarity signal (blue) along with the associated trend/respiration (pink) and residual/heart-beat (green) components.
-Figure~\ref{fig:phase_estimation}(e) shows the periodogram of these two components.
-Through Figures~\ref{fig:phase_estimation}(b-e), notice how well the two periodic sources have been decoupled from the raw frame similarity signal.
-Also, in Figure~\ref{fig:phase_estimation}, notice that most of the energy in the periodogram or power-frequency distribution of the respiration and heart-beat components is concentrated within a narrow frequency band.
-Furthermore, notice how these 1D signals capture the periodicity characteristics (e.g.
-frequency in beats-per-minute (bpm), period or cycle duration in frames, number of cycles) of the respective sources of motion in the input high-dimensional periodic image sequence.
+Figure~\ref{fig:phase_estimation}(e) shows the periodogram of these two components that are nicely segregated in frequency domain.
+The periodicity characteristics (e.g.
+frequency in beats-per-minute (bpm), cycle duration in frames, number of cycles) of heart-beat and respiration can be calculated based on the maximum-power frequency of their periodograms.
 
-	Next, considering the narrow-band nature of the derived trend/respiratory and residual/heart-beat signals, we use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous phase of each frame.
-Specifically, we compute the instantaneous intra-period phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows:
-\begin{equation}
-\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)
-\end{equation} 
-and map $\phi(t)$ from the range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
-Let $\hat{\phi}_{heart}(t)$ and $\hat{\phi}_{resp}(t)$ denote the instantaneous cardiac and respiratory phases computed from the trend/respiration $\hat{\tau}_{resp}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signals, respectively, using the aforementioned approach.
-Figure~\ref{fig:phase_estimation}(f) shows the instantaneous cardiac and respiratory phases computed from the heart-beat and respiration signals in in Figure~\ref{fig:phase_estimation}(d).
-
+	Next, considering the narrow-band nature (see  Figure~\ref{fig:phase_estimation}) of the derived trend/respiratory and residual/heart-beat signals, we use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous phase of each frame.
+Specifically, we compute the instantaneous intra-period phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows: $\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)$ and map $\phi(t)$ from the range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
+Let $\hat{\phi}_{heart}(t)$ and $\hat{\phi}_{resp}(t)$ denote the instantaneous cardiac and respiratory phases computed from the trend/respiration $\hat{\tau}_{resp}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signals, respectively.
+Figure~\ref{fig:phase_estimation}(f) shows the instantaneous cardiac and respiratory phases computed from the heart-beat and respiration signals in Figure~\ref{fig:phase_estimation}(d).
+%
 %\vspace{-0.5cm}
@@
 %
-Once the instantaneous cardiac and respiratory phases of each frame has been estimated, it can be used to facilitate the extraction of quantitative measurements from a desired part/point of the periodic cycle, a process commonly referred to as gating.
-In this section, we present a robust two-step method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.
-
-	In the first step, based on the observation that the predominant influence of respiratory motion occurs around the troughs/minima of the trend/respiration signal $\hat{\tau}_{resp}(t)$ (see pink curve in Fig.~\ref{fig:phase_estimation}(d)) corresponding to the respiratory phase $\hat{\phi}_{resp}(t) = 0$, we perform a rough initial gating by discarding the frames $F_{cutoff} = \left \{ t \mid \hat{\phi}_{resp}(t) < c \parallel \hat{\phi}_{resp}(t) > (1 - c) \right \}$ whose respiratory phase distance from $\hat{\phi}_{resp}(t) = 0$ is below a specified cutoff value $c$ that we set to 0.2.
-Figure~\ref{fig:respiratory_gating}(a) shows the discarded frames $F_{cutoff}$ overlaid (blue circles) on the trend/respiration $\hat{\tau}_{resp}(t)$, residual/heart-beat  $\hat{r}_{heart}(t)$, and respiratory phase $\hat{\phi}_{resp}(t)$ signals.
-
-	In the second step, we learn a robust mapping $S(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the residual/heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called Locally weighted regression (LOESS) to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signal values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above.
-Next, we compute the median absolution value $\gamma_{MAD}$ of the difference between the LOESS fit $S( \hat{\phi}_{heart}(t) )$ and residual/heart-beat signal $\hat{r}_{heart}(t)$ for all frames as a robust estimate of the dispersion of non-respiratory frames around the LOESS fit.
-Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert S( \hat{\phi}_{heart}(t) ) - \hat{r}_{heart}(t)  \rvert   > k \times \gamma_{mad}  \right \}$ for which the difference between the LOESS fit and heart-beat signal value is greater than $k \times \gamma_{mad}$ (we set $k = 2.0$).
 %
@@
 %
-\caption{Illustration of our method for gating out respiratory frames: (a) Frames $F_{cutoff}$ discarded (blue circles) in the first step overlaid with the trend/respiration, residual/heart-beat, and instantaneous respiratory phase signals (b) }
+\caption{Illustration of our two-step method for gating out respiratory frames: (a) Frames $F_{cutoff}$ discarded (blue circles) in the first step overlaid with the trend/respiration, residual/heart-beat, and instantaneous respiratory phase signals (b) residual/heart-beat signal $\hat{r}_{heart}(t)$ vs cardiac phase $\hat{\phi}_{heart}(t)$ overlaid with frames $F_{cutoff}$ discarded in step-1 (blue circle), LOWESS fit $L(\phi)$ (black solid), upper and lower bounds or 95\% confidence interval $(L(\phi) \pm 2.0 * \hat{\sigma}_L)$ of valid non-respiratory frames (black dotted), and frames $F_{resp}$ (red circles) gated at the end of step-2.}
 \label{fig:respiratory_gating}
@@
 %
+Once the cardiac and respiratory phases of each frame has been estimated, it can be used to guide the extraction of quantitative measurements to a desired part/point of the periodic cycle, a process commonly referred to as gating.
+In this section, we present a robust two-step method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.
+
+	In the first step, based on the observation that the predominant influence of respiratory motion occurs around the troughs/minima of the trend/respiration signal $\hat{\tau}_{resp}(t)$ (see pink curve in Fig.~\ref{fig:phase_estimation}(d)) corresponding to the respiratory phase $\hat{\phi}_{resp}(t) = 0$, we perform a rough initial gating by discarding the frames $F_{cutoff} = \left \{ t \mid \hat{\phi}_{resp}(t) < c \parallel \hat{\phi}_{resp}(t) > (1 - c) \right \}$ whose respiratory phase distance from $\hat{\phi}_{resp}(t) = 0$ is below a specified cutoff value $c$ that we set to 0.2.
+Figure~\ref{fig:respiratory_gating}(a) shows the discarded frames $F_{cutoff}$ overlaid (blue circles) on the trend/respiration $\hat{\tau}_{resp}(t)$, residual/heart-beat  $\hat{r}_{heart}(t)$, and respiratory phase $\hat{\phi}_{resp}(t)$ signals.
+
+	In the second step, we learn a robust mapping $L(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the residual/heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called Locally weighted regression (LOWESS) to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signal values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above.
+Next, we compute a robust estimate of the standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit based on the median absolute value of the difference between the LOWESS fit $L( \hat{\phi}_{heart}(t) )$ and residual/heart-beat signal $\hat{r}_{heart}(t)$ for all frames.
+Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert L( \hat{\phi}_{heart}(t) ) - \hat{r}_{heart}(t)  \rvert   > k \times \hat{\sigma}_{L}  \right \}$ for which the absolute  difference between the LOWESS fit and heart-beat signal value is greater than $k = 0.2$ times the standard deviation $\hat{\sigma}_{L}$.
+%
 %\vspace{-0.5cm}
-\subsection{Learning to reconstruct image at any cardiac phase}
+\subsection{Model to generate images by cardiac phase for super-resolution}
 \label{sec:method:super_resolution}
 %
-Once the instantaneous cardiac phase of each frame has been estimated, it can be used to enable spatio-temporal super-resolution by considering the set of images within each period of a periodic image sequence as distinct observations of the same data.
-In this section, we present a method that uses a phase-tagged periodic image sequence to reconstruct a single-period video at a higher temporal resolution.
+In this section, we present a kernel regression model to reconstruct the image at any cardiac phase.
+This model can then be used to generate a single-cycle video representative of the subject's heart-beat at a desired higher temporal resolution to achieve temporal super-resolution.
 
-Given a de-noised periodic sequence $\{L_1, ..., L_n\}$ of $n$ images (with $m$ pixels each) along with their estimated intra-period phases $\{\phi_1, ..., \phi_n\}$, we first use Nadarya-Watson kernel regression\cite{Bishop2006} to construct a phase-parameterized image manifold in the form of a function $M(\phi): [0, 1) \to R^m $ that maps any intra-period phase $\phi$ to an image as follows:
+Given a cardiac ultrasound video $I(t) : \{1, ..., N\} \to R^m$ of $N$ images with $m$ pixels each, we extract the residual/heart-beat component $\hat{r}_{heart}(t)$ of the frame similarity signal, estimate the instantaneous cardiac phase $\hat{\phi}_{heart}(t)$ of each frame, and compute the robust LOWESS fit $L(\phi) : \left [  0, 1\right ) \to R$ that maps cardiac phase to the heart-beat signal as  described in Sections~\ref{sec:method:phase_estimation} and ~\ref{sec:method:gating}.
+We then use Nadarya-Watson kernel regression\cite{Bishop2006} to learn a phase-parametrized image manifold in the form of a function $M(\phi): [0, 1) \to R^m $ that maps any cardiac phase $\phi$ to an image using kernel-weighted local average as follows:
 \begin{equation}
-M(\phi) = \frac{\sum_{i = 1}^{n} K(\phi, \phi_i) L_i}{\sum_{i = 1}^{n} K(\phi, \phi_i)} 
+M(\phi) = \frac{\sum_{t = 1}^{N} K \left( \phi, \hat{\phi}_{heart}(t) \right) I(t)}{\sum_{t = 1}^{n} K \left( \phi, \hat{\phi}_{heart}(t) \right)} 
 \end{equation}
-where in we define the kernel $K$ using a gaussian with standard-deviation $\sigma$ equal to a multiple of the average distance between consecutive phases after sorting them.
-Note that, we account for the periodic nature of the intra-period phase whenever we measure distance between two phase values.
-A single-period video can now be reconstructed by sampling the manifold $M(\phi)$ at any desired resolution from the range $[0, 1)$ of intra-period phase.
-Note that the aformentioned process will facilitate temporal super-resolution only if the set of images observed in each distinct period of the given image sequence differ in intra-period phase by some non-zero amount less than its temporal sampling interval.
-Also, the amount of temporal super-resolution possible will depend both on the number of periods observed and the maximum amount by which one can expect the images to be perturbed in phase across observations.
+where kernel $K\left( \phi, \hat{\phi}_{heart}(t) \right) = exp\left \{ -\frac{ \mid \phi - \hat{\phi}_{heart}(t) \mid^2}{2  \sigma^2_\phi} \right \} \times exp\left \{ -\frac{ \mid L(\phi) - \hat{r}_{heart}(t) \mid^2}{2  \sigma^2_{L}} \right \}$ is defined as the product of two radial-basis function (RBF) kernels.
+The first RBF kernel  weighs images inversely proportional to their distance (accounting for periodicity) in the cardiac phase space.
+Its bandwidth $\sigma_\phi$ is set equal to a constant $k_\phi = 0.4$ times the median difference in cardiac phase between consecutive frames of the given image sequence.
+The second RBF kernel weighs images inversely proportional to their deviation from the LOWESS prediction $L(\phi)$ in the residual/heart-beat signal space.
+Its bandwidth $\sigma_{L}$ is set equal to a constant $k_L = 2.0$ times the robust estimate of standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit as described in Section~\ref{sec:method:gating}.
+
+	A representative single-cycle video of the subject's heart-beat can now be reconstructed by sampling the manifold $M(\phi)$ at any desired resolution from the range $[0, 1)$ of the cardiac phase.
+Note that this approach will facilitate temporal super-resolution only if the set of images observed in each distinct period of the given image sequence differ in cardiac phase by some non-zero amount.
+Also, the amount of temporal super-resolution possible will depend both on the number of cycles captured and the amount by which one can expect the corresponding images to be perturbed in phase across cycles.
 %\vspace{-0.5cm}

### 2017-02-21 paper155.tex
Added tables into latex file

@@
 %
+\begin{table}[h]
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+VID & $R^2$ & Mean $\pm$ Std & Median $\pm$ IQR \\ \hline
+1 & 0.9995 & 1.55 $\pm$ 0.89 & 1.00 $\pm$ 1.00 \\ \hline
+2 & 0.9995 & 1.36 $\pm$ 1.15 & 1.00 $\pm$ 2.00 \\ \hline
+3 & 0.9996 & 1.45 $\pm$ 0.89 & 2.00 $\pm$ 1.00 \\ \hline
+4 & 0.9998 & 0.73 $\pm$ 0.62 & 1.00 $\pm$ 1.00 \\ \hline
+5 & 0.9990 & 2.09 $\pm$ 1.78 & 1.00 $\pm$ 2.00 \\ \hline
+6 & 0.9992 & 1.40 $\pm$ 1.74 & 1.00 $\pm$ 0.75 \\ \hline
+\end{tabular}
+\caption{My caption}
+\label{my-label}
+\end{table}
+%
+\begin{table}[h]
+\centering
+\begin{tabular}{|c|c|c|}
+\hline
+\begin{tabular}[c]{@{}c@{}}\\ VID \end{tabular} & \begin{tabular}[c]{@{}c@{}}LOOCV (50) NCORR\\ Mean $\pm$ Std\end{tabular} & \begin{tabular}[c]{@{}c@{}}QRS Peak Frame NCORR\\ Mean $\pm$ Std\end{tabular} \\ \hline
+1 & 0.8562 $\pm$ 0.0366 & 0.7981 $\pm$ 0.0532 \\ \hline
+2 & 0.8558 $\pm$ 0.0525 & 0.7877 $\pm$ 0.0503 \\ \hline
+3 & 0.8592 $\pm$ 0.0348 & 0.7849 $\pm$ 0.0457 \\ \hline
+4 & 0.8713 $\pm$ 0.0348 & 0.8395 $\pm$ 0.0486 \\ \hline
+5 & 0.8655 $\pm$ 0.0312 & 0.8196 $\pm$ 0.0476 \\ \hline
+6 & 0.8736 $\pm$ 0.0278 & 0.8270 $\pm$ 0.0778 \\ \hline
+\end{tabular}
+\caption{My caption}
+\label{my-label}
+\end{table}
 %

### 2017-02-21 paper155.tex
Updated text of results section

--- 2017-02-21 paper155.tex
@@
 %
+\begin{IonFigT}
+\centering
+%
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/phaseordered.png}
+\ionbox{1.0in}\ionbox{1.0in}\ionbox{1.0in}\ionbox{1.0in}\\
+%
+\caption{Visualization of m-mode frames ordered by estimated cardiac phase}
+\label{fig:phase_ordering}
+\end{IonFigT}
+%
 %\vspace{-0.5cm}
@@
 %
+We use cardiac ultrasound videos and associated ECG recordings of 6 mice to validate the proposed methods.
+The ultrasound videos were acquired using the VisualSonics Vevo 2100 scanner at 233 frames per second.
+Furthermore, each video consisted of approximately 300 frames, 11 cardiac cycles, and 2 respiratory cycles.
+
+	We validate our cardiac phase estimation method by comparing its phase estimates with the ECG signal.
+Figure~\ref{fig:instaphase_vs_ecg}(a) shows a head-to-head comparison between the ECG signal and the instantaneous cardiac phase obtained using the method described in Section~\ref{sec:method:phase_estimation} for one of the 6 videos.
+
 %
-\begin{IonFigT}
+\begin{table}[h]
 \centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+\begin{tabular}[c]{@{}c@{}}VID\\ $\;$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}$R^2$\\ $\;$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}QRS Peak Prediction Error\\ Mean $\pm$ Std\end{tabular} \\ \hline
+1 & 0.9995 & 1.55 $\pm$ 0.89 \\ \hline
+2 & 0.9995 & 1.36 $\pm$ 1.15 \\ \hline
+3 & 0.9996 & 1.45 $\pm$ 0.89 \\ \hline
+4 & 0.9998 & 0.73 $\pm$ 0.62 \\ \hline
+5 & 0.9990 & 2.09 $\pm$ 1.78 \\ \hline
+6 & 0.9992 & 1.40 $\pm$ 1.74 \\ \hline
+\end{tabular}
+\caption{Comparision between frame positions of QRS peaks in the ECG signal and the corresponding minima of the cardiac phase estimated using our method for 6 videos}
+\label{table:phase_estimation}
+\end{table}
 %
-\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/phaseordered.png}
-\ionbox{5.0in}\\
+\begin{table}[h]
+\centering
+\begin{tabular}{|c|c|c|}
+\hline
+\begin{tabular}[c]{@{}c@{}}VID \\ $\;$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}LOOCV (50) NCORR\\ Mean $\pm$ Std\end{tabular} & \begin{tabular}[c]{@{}c@{}}QRS Peak Frame NCORR\\ Mean $\pm$ Std\end{tabular} \\ \hline
+1 & 0.8562 $\pm$ 0.0366 & 0.7981 $\pm$ 0.0532 \\ \hline
+2 & 0.8558 $\pm$ 0.0525 & 0.7877 $\pm$ 0.0503 \\ \hline
+3 & 0.8592 $\pm$ 0.0348 & 0.7849 $\pm$ 0.0457 \\ \hline
+4 & 0.8713 $\pm$ 0.0348 & 0.8395 $\pm$ 0.0486 \\ \hline
+5 & 0.8655 $\pm$ 0.0312 & 0.8196 $\pm$ 0.0476 \\ \hline
+6 & 0.8736 $\pm$ 0.0278 & 0.8270 $\pm$ 0.0778 \\ \hline
+\end{tabular}
+\caption{My caption}
+\label{table:image_reconstruction}
+\end{table}
 %
-\caption{Visualization of m-mode frames ordered by estimated cardiac phase}
-\label{fig:phase_ordering}
-\end{IonFigT}
-%
+
 %

### 2017-02-21 paper155.tex
Finished first draft of results section

--- 2017-02-21 paper155.tex
@@
 \usepackage{multirow}
+\usepackage{wrapfig}
 %
@@
 %
-Ultrasound is emerging as an increasingly effective tool for rapid non-invasive visual assessment of cardiac structure and function. 
+Cardiovascular disease is the leading cause of death worldwide and ultrasound is emerging as an increasingly effective tool for rapid non-invasive visual assessment of cardiac structure and function. 
 %
 
-%\vspace{-0.5cm}
+\vspace{-0.3cm}
 \section{Method}
@@
 %
-\caption{Illustration of correspondence between ECG and the instantaneous cardiac phase estimated using the proposed method: (a) ECG signal (blue) and the cardiac phase estimated using our method overlaid with peaks of the QRS complex (red circle) and the corresponding cardiac phase minima (black circle), (b) Five video frames evenly spaced in time between the first and second QRS peaks of the ECG signal (top-row) and the corresponding cardiac phase minima (bottom-row) constituting one cardiac cycle.}
+\caption{Illustration of correspondence between ECG and the instantaneous cardiac phase estimated using our method: (a) ECG signal (blue) and the cardiac phase estimated using our method overlaid with peaks of the QRS complex (red circle) and the corresponding cardiac phase minima (black circle), (b) Five video frames evenly spaced in time between the first and second QRS peaks of the ECG signal (top-row) and the corresponding cardiac phase minima (bottom-row) constituting one cardiac cycle.}
 \label{fig:instaphase_vs_ecg}
 \end{IonFigT}
-%\vspace{-0.5cm}
+\vspace{-0.3cm}
 \subsection{Estimation of instantaneous cardiac and respiratory phases}
@@
 %
-%\vspace{-0.5cm}
+\vspace{-0.3cm}
 \subsection{Gating out respiratory frames}
@@
 %
-%\vspace{-0.5cm}
+\begin{IonFigT}
+\centering
+%
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/phaseordered.png}
+\ionbox{1.2in}\ionbox{1.1in}\ionbox{1.1in}\ionbox{1.1in}\\
+%
+\caption{Visualization of m-mode frames ordered by estimated cardiac phase: (a) One of the frames in the video overlaid with the M-mode line shown in the next three images, (b) M-mode frames in the order they appear in the input video, (c) M-mode frames ordered by estimated cardiac phase, and (d) Non-respiratory m-mode frames ordered by estimated cardiac phase.}
+\label{fig:phase_ordering}
+\end{IonFigT}
+%
+\vspace{-0.3cm}
 \subsection{Model to generate images by cardiac phase for super-resolution}
@@
 Also, the amount of temporal super-resolution possible will depend both on the number of cycles captured and the amount by which one can expect the corresponding images to be perturbed in phase across cycles.
-%\vspace{-0.5cm}
+\vspace{-0.3cm}
 \section{Results}
@@
 %
+We use cardiac ultrasound videos and associated ECG recordings of 6 mice to validate our methods.
+The ultrasound videos were acquired using the VisualSonics Vevo 2100 scanner at 233 frames per second.
+Each video consists of approximately 300 frames, 11 cardiac cycles, and 2 respiratory cycles.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+\begin{tabular}[c]{@{}c@{}}VID\\ $\;$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}Correlation \\ $R^2$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}QRS peak prediction error\\ $mean \pm std$\end{tabular} \\ \hline
+1 & 0.9995 & 1.55 $\pm$ 0.89 \\ \hline
+2 & 0.9995 & 1.36 $\pm$ 1.15 \\ \hline
+3 & 0.9996 & 1.45 $\pm$ 0.89 \\ \hline
+4 & 0.9998 & 0.73 $\pm$ 0.62 \\ \hline
+5 & 0.9990 & 2.09 $\pm$ 1.78 \\ \hline
+6 & 0.9992 & 1.40 $\pm$ 1.74 \\ \hline
+\end{tabular}
+\caption{Comparision between frame positions of QRS peaks in the ECG signal and the corresponding minima of the estimated cardiac phase for 6 videos}
+\label{table:phase_estimation}
+\end{table}
+	We validate our cardiac phase estimation method by comparing its phase estimates with the ECG signal which is the gold standard for cardiac gating.
+Figure~\ref{fig:instaphase_vs_ecg}(a) shows a head-to-head comparison between the ECG signal and the instantaneous cardiac phase signal obtained using the method described in Section~\ref{sec:method:phase_estimation} for one of the 6 videos.
+Notice how well the locations of the peaks of the QRS complex (red circles) in the ECG signal and the corresponding minima (black circles) of the cardiac phase signal in each cycle match.
+Table 1 reports the $R^2$ correlation and the $mean \pm std$ error between the locations of the QRS peaks and the corresponding minima of the cardiac phase signals for all 6 videos in our dataset.
+Figure~\ref{fig:instaphase_vs_ecg}(b) shows a head-to-head comparison of five video frames evenly spaced in time between two consecutive QRS peaks (top-row) and corresponding minima of the cardiac phase signal (bottom-row).
+Notice the high-level of similarity in the images of each column.
+Figure~\ref{fig:phase_ordering}(b,c) shows the frames of an m-mode line before and after ordering with the estimated cardiac phase.
+Figure~\ref{fig:phase_ordering}(d) shows the phase-ordered m-mode frames after performing respiratory gating as described in Section~\ref{sec:method:gating}.
+
+	We validate the accuracy of our kernel regression model for reconstructing the image at any cardiac phase using leave-one-out-cross-validation (LOOCV).
+In each round of cross-validation, we randomly exclude one of the non-respiratory frames in the video, fit our kernel-regression model on the remaining frames, use the fitted model to reconstruct the image at the cardiac phase of the excluded frame, and compute the similarity between the reconstructed and original image using normalized correlation.
+The second column in Table~\ref{table:image_reconstruction} shows the mean and standard deviation of normalized correlation between the reconstructed and the original images over 50 rounds of LOCCV for each of the 6 videos in our dataset.
+As a baseline, the third column of  Table~\ref{table:image_reconstruction} shows the mean and standard deviation of the normalized correlation between frames at QRS peaks of the ECG signal for each video.
+
+	We encourage the readers to watch the supplementary videos for a better assessment of the results of the proposed methods.
 %
-\begin{IonFigT}
+\begin{table}[h]
 \centering
+\begin{tabular}{|c|c|c|}
+\hline
+\begin{tabular}[c]{@{}c@{}}VID \\ $\;$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}LOOCV (50) NCORR\\ $mean \pm std$\end{tabular} & \begin{tabular}[c]{@{}c@{}}QRS Peak Frame NCORR\\ $mean \pm std$\end{tabular} \\ \hline
+1 & 0.8562 $\pm$ 0.0366 & 0.7981 $\pm$ 0.0532 \\ \hline
+2 & 0.8558 $\pm$ 0.0525 & 0.7877 $\pm$ 0.0503 \\ \hline
+3 & 0.8592 $\pm$ 0.0348 & 0.7849 $\pm$ 0.0457 \\ \hline
+4 & 0.8713 $\pm$ 0.0348 & 0.8395 $\pm$ 0.0486 \\ \hline
+5 & 0.8655 $\pm$ 0.0312 & 0.8196 $\pm$ 0.0476 \\ \hline
+6 & 0.8736 $\pm$ 0.0278 & 0.8270 $\pm$ 0.0778 \\ \hline
+\end{tabular}
+\caption{Evaluation of our kernel regression model for reconstructing the image at any cardiac phase using 50 rounds of Leave one out cross-validation (LOOCV)}
+\label{table:image_reconstruction}
+\end{table}
 %
-\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/phaseordered.png}
-\ionbox{5.0in}\\
+
 %
-\caption{Visualization of m-mode frames ordered by estimated cardiac phase}
-\label{fig:phase_ordering}
-\end{IonFigT}
-%
-%
-%\vspace{-0.5cm}
+\vspace{-0.3cm}
 \section{Discussion and future work}
@@
 %
-%\vspace{-0.5cm}
+\vspace{-0.3cm}
 \bibliographystyle{splncs03}

### 2017-02-21 paper155.tex
Started writing the introduction section

@@
 Cardiovascular disease is the leading cause of death worldwide and ultrasound is emerging as an increasingly effective tool for rapid non-invasive visual assessment of cardiac structure and function.
+Two kinds of periodic motion are present in cardiac ultrasound videos, one that corresponds to the beating heart motion and the other that corresponds to respiratory motion.
+Knowing the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications.
+Cardiac motion is usually tracked by simultaneously acquiring an ECG signal and respiratory motion is tracked through the motion of markers placed on the subject's body.
 %

### 2017-02-21 paper155.tex
Updated introduction section

--- 2017-02-21 paper155.tex
@@
 \newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\operatorname{min}}\;}
+\setlength{\belowcaptionskip}{-15pt}
+\setlength{\abovecaptionskip}{-2pt}
 %
@@
 Next, we present a robust non-parametric regression technique that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.
-Lastly, we use a kernel regression with a novel kernel to learn an image manifold parameterized by instantaneous cardiac phase that can be sampled to generate a single-cycle video at a higher temporal-resolution.
-We demonstrate and validate our methods using cardiac ultrasound videos and associated ECG recordings of 6 mice.
+Lastly, we use a kernel regression method with a novel kernel to learn an image manifold parameterized by instantaneous cardiac phase that can be sampled to generate a single-cycle video at a higher temporal-resolution.
+We demonstrate and validate our methods using 2D cardiac ultrasound videos and associated ECG recordings of 6 mice.
 %\keywords{Cardiac, Ultrasound, Phase estimation, Gating, Temporal super-resolution}
 \end{abstract}
-
-%\vspace{-0.5cm}
+%
+\vspace{-0.5cm}
 \section{Introduction}
@@
 Cardiovascular disease is the leading cause of death worldwide and ultrasound is emerging as an increasingly effective tool for rapid non-invasive visual assessment of cardiac structure and function.
+Cardiac ultrasound videos consists of two kinds of periodic motion, one that corresponds to the beating heart motion and the other that corresponds to respiratory motion.
+Knowing the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating, quiescence detection, 3D reconstruction, and robotic beating heart surgery\cite{VonBirgelen1997,Wick2013,Wachinger2012,Kurz2015}.
+Typically, the position within the cardiac cycle is tracked through ECG data and the position within respiratory cycle is tracked by motion of markers placed on the subject's body\cite{Khamene2004}.
+
+	In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the images of the cardiac ultrasound video.
+We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression to reconstruct images at any cardiac phase to facilitate temporal super-resolution.
 %
-
 \vspace{-0.3cm}
@@
 We then present a robust method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion in Section \ref{sec:method:gating}.
-Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression method that uses the phase-tagged image sequence to reconstruct a single-period video at higher-temporal resolution.
+Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression model to generate the image at any cardiac phase that can be used to achieve temporal super-resolution.
 %
@@
 %
-While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data (wherein a single-variable is measured at each time point)\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting which is the case with periodic cardiac ultrasound videos wherein several thousands of variables (pixel intensities) could be measured at each time point.
-Our strategy was to find a way to simplify this complex multi-variate problem by transforming it into a univariate one and take advantage of existing univariate methods to solve the problem.
+While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting such as the case of periodic cardiac ultrasound videos wherein several thousands of variables (pixel intensities) could be measured at each time point.
+Our strategy was to find a way to transforming this complex multi-variate problem into a univariate one and take advantage of existing univariate methods to solve the problem.
 

### 2017-02-21 paper155.tex
Updated paper and figures

@@
 \begin{abstract}
-In this paper, we present a novel purely image-based method for the estimation of instantaneous cardiac and respiratory phases from cardiac ultrasound videos and demonstrate its utility in gating and temporal super-resolution.
-We first use a novel strategy based on inter-frame similarity to transform the complex high-dimensional image sequence into a 1D time series with the same periodicity characteristics.
-Next, we use a trend extraction technique to decompose this time series into a trend component encoding respiratory motion alone and a residual component encoding the beating heart motion alone.
-Next, we use the Hilbert transform to estimate the instantaneous cardiac and respiratory phases of each frame from these components.
-Next, we present a robust non-parametric regression technique that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.
-Lastly, we use a kernel regression method with a novel kernel to learn an image manifold parameterized by instantaneous cardiac phase that can be sampled to generate a single-cycle video at a higher temporal-resolution.
+Ultrasound is emerging as an increasingly effective tool for rapid non-invasive visual assessment of cardiac structure and function.
+Knowing the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications.
+In this paper, we present a novel method for the estimation of instantaneous cardiac and respiratory phases directly from cardiac ultrasound videos.
+The method: (i) transforms the complex high-dimensional image sequence into a univariate time series with the same periodicity characteristics, (ii) decouples the sources of beating heart and respiratory motion using a trend extraction technique, and (iii) estimates the cardiac and respiratory phases using Hilbert transform.
+We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression to reconstruct images at any cardiac phase to facilitate temporal super-resolution.
 We demonstrate and validate our methods using 2D cardiac ultrasound videos and associated ECG recordings of 6 mice.
@@
 
-	In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the images of the cardiac ultrasound video.
+	In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the images of the cardiac ultrasound video and demonstrate its utility in respiratory gating and temporal super-resolution.
 We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression to reconstruct images at any cardiac phase to facilitate temporal super-resolution.
@@
 	We first compute the similarity between all pairs of frames in the given periodic image sequence containing $N$ images/frames to create a matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames of the given sequence.
-Here in, we use normalization correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity but, in principle, any of the similarity metrics used in image registration algorithms \cite{Goshtasby2012} can potentially be chosen.
+Here in, we use normalization correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity but, in principle, any of the similarity metrics used in image registration algorithms can potentially be chosen.
 Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series that measures the similarity of the corresponding frame in the image sequence with all other frames.
@@
 
-	We encourage the readers to watch the supplementary videos for a better assessment of the results of the proposed methods.
+	Lastly, since this paper deals with video analysis we encourage the readers to review the supplementary material for a better understanding of the utility and potential of the proposed methods.
 %
@@
 %
-
 %
 \vspace{-0.3cm}
-\section{Discussion and future work}
-\label{sec:discussion}
+\section{Conclusion}
+\label{sec:conclusion}
 %
-
+In this paper, we have presented a novel method to estimate the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video.
+We have also presented a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase that can be used to facilitate temporal super-resolution.
+Having demonstrated a proof of concept, in the future, we plan to evaluate our methods on more datasets, especially on cardiac ultrasound videos of subjects with irregular heart beats, and address any remaining pitfalls.
+One drawback of our phase estimation method is a strong assumption of periodicity which may breakdown in the case of subjects with cardiac arrhythmia.
+To address this, we will look into generalizations of the hilbert transform\cite{Lu2013} that are more susceptible noise.
+Another drawback is that the use normalized correlation to measure inter-frame similarity relies on an inherent assumption that a large part of the image is pulsating.
+To address this, we will try to devise more sophisticated patch-based similarity measures.
 %

### 2017-02-22 paper155.tex
Incorporated corrections from proof reading

@@
 In this paper, we present a novel method for the estimation of instantaneous cardiac and respiratory phases directly from cardiac ultrasound videos.
-The method: (i) transforms the complex high-dimensional image sequence into a univariate time series with the same periodicity characteristics, (ii) decouples the sources of beating heart and respiratory motion using a trend extraction technique, and (iii) estimates the cardiac and respiratory phases using Hilbert transform.
-We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression to reconstruct images at any cardiac phase to facilitate temporal super-resolution.
-We demonstrate and validate our methods using 2D cardiac ultrasound videos and associated ECG recordings of 6 mice.
+The method transforms the complex high-dimensional image sequence into a univariate time series with the same periodicity characteristics, decouples the periodic sources of beating heart and respiratory motion using a trend extraction technique, and estimates the cardiac and respiratory phases using Hilbert transform.
+We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
+We validate our methods using 2D cardiac ultrasound videos and associated ECG recordings of 6 mice.
 %\keywords{Cardiac, Ultrasound, Phase estimation, Gating, Temporal super-resolution}
@@
 Cardiac ultrasound videos consists of two kinds of periodic motion, one that corresponds to the beating heart motion and the other that corresponds to respiratory motion.
-Knowing the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating, quiescence detection, 3D reconstruction, and robotic beating heart surgery\cite{VonBirgelen1997,Wick2013,Wachinger2012,Kurz2015}.
+Knowing the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection\cite{Wick2013}, 3D reconstruction\cite{Wachinger2012}, temporal super-resolution and robotic beating heart surgery\cite{Kurz2015}.
 Typically, the position within the cardiac cycle is tracked through ECG data and the position within respiratory cycle is tracked by motion of markers placed on the subject's body\cite{Khamene2004}.
 
-	In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the images of the cardiac ultrasound video and demonstrate its utility in respiratory gating and temporal super-resolution.
+	In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the the cardiac ultrasound video.
 We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression to reconstruct images at any cardiac phase to facilitate temporal super-resolution.
@@
 Specifically, we begin by describing our method for the estimation of instantaneous cardiac and respiratory phases in Section \ref{sec:method:phase_estimation}.
-We then present a robust method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion in Section \ref{sec:method:gating}.
+We then present a robust method that uses these phase estimates to gate out video frames with predominant respiratory motion in Section \ref{sec:method:gating}.
 Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression model to generate the image at any cardiac phase that can be used to achieve temporal super-resolution.
@@
 %
-\caption{Illustration of our phase estimation method: (a) Inter-frame similarity matrix, (b) Trend component matrix corresponding to respiratory motion, (c) Residual component matrix corresponding to beating heart motion, (d) Similarity profile chosen for phase estimation with associated trend/respiration and residual/heart-beat components, (e) Periodogram of heart-beat and respiration components along with periodicity characteristics (e.g.
-frequency, cycle duration) calculated for the peak frequency, and (f) Instantaneous cardiac and respiratory phases estimated using the Hilbert transform.}
+\caption{Illustration of our phase estimation method: (a) Inter-frame similarity matrix, (b) Matrix of trend components corresponding to respiratory motion, (c) Matrix of residual components corresponding to beating heart motion, (d) Similarity profile chosen for phase estimation with associated trend/respiration and residual/heart-beat components, (e) Periodogram of heart-beat and respiration components along with periodicity characteristics (e.g.
+frequency, cycle duration) calculated from the dominant frequency, and (f) Instantaneous cardiac and respiratory phases estimated using the Hilbert transform.}
 \label{fig:phase_estimation}
@@
 %
-While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting such as the case of periodic cardiac ultrasound videos wherein several thousands of variables (pixel intensities) could be measured at each time point.
+While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting such as the case of periodic cardiac ultrasound videos wherein several thousands of variables (pixel intensities) are involved.
 Our strategy was to find a way to transforming this complex multi-variate problem into a univariate one and take advantage of existing univariate methods to solve the problem.
 
-	We first compute the similarity between all pairs of frames in the given periodic image sequence containing $N$ images/frames to create a matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames of the given sequence.
+	We first compute the similarity between all pairs of frames in the given periodic image sequence containing $N$ images/frames to create a matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames.
 Here in, we use normalization correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity but, in principle, any of the similarity metrics used in image registration algorithms can potentially be chosen.
-Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series that measures the similarity of the corresponding frame in the image sequence with all other frames.
-If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted we can expect this time series to encompass the predominant periodicity characteristics of the original image sequence.
+Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series.
+Furthermore, If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted, this time series can be expected to preserve the periodicity characteristics of the original image sequence.
 Two kinds of periodic motion are present in cardiac ultrasound videos, one corresponding to the beating heart motion and the other corresponding to respiratory motion.
-Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos computed using the normalized correlation metric wherein these two sources of periodicity can be clearly observed in many rows/columns.
+Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos wherein these two sources of periodicity can be clearly observed in many rows/columns.
 
@@
 
-	Next, considering the narrow-band nature (see  Figure~\ref{fig:phase_estimation}) of the derived trend/respiratory and residual/heart-beat signals, we use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous phase of each frame.
+	Next, considering the narrow-band nature of the trend/respiration and residual/heart-beat signals (see  Figure~\ref{fig:phase_estimation}(e)), we use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous phase of each frame.
 Specifically, we compute the instantaneous intra-period phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows: $\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)$ and map $\phi(t)$ from the range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
@@
 %
-Once the cardiac and respiratory phases of each frame has been estimated, it can be used to guide the extraction of quantitative measurements to a desired part/point of the periodic cycle, a process commonly referred to as gating.
+Once the cardiac and respiratory phases of each frame have been estimated, it can be used to filter out or guide the extraction of quantitative measurements from a desired part/point of the periodic cycle, a process commonly referred to as gating.
 In this section, we present a robust two-step method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.
 
-	In the first step, based on the observation that the predominant influence of respiratory motion occurs around the troughs/minima of the trend/respiration signal $\hat{\tau}_{resp}(t)$ (see pink curve in Fig.~\ref{fig:phase_estimation}(d)) corresponding to the respiratory phase $\hat{\phi}_{resp}(t) = 0$, we perform a rough initial gating by discarding the frames $F_{cutoff} = \left \{ t \mid \hat{\phi}_{resp}(t) < c \parallel \hat{\phi}_{resp}(t) > (1 - c) \right \}$ whose respiratory phase distance from $\hat{\phi}_{resp}(t) = 0$ is below a specified cutoff value $c$ that we set to 0.2.
+	In the first step, based on the observation that respiratory motion mainly occurs around the minima of the trend/respiration signal $\hat{\tau}_{resp}(t)$ (pink curve in Fig.~\ref{fig:phase_estimation}(d)) with respiratory phase $\hat{\phi}_{resp}(t) = 0$, we perform a rough initial gating by discarding the frames $F_{cutoff} = \left \{ t \mid \hat{\phi}_{resp}(t) < c \parallel \hat{\phi}_{resp}(t) > (1 - c) \right \}$ whose respiratory phase distance from $\hat{\phi}_{resp}(t) = 0$ is below a specified cutoff value $c=0.2$.
 Figure~\ref{fig:respiratory_gating}(a) shows the discarded frames $F_{cutoff}$ overlaid (blue circles) on the trend/respiration $\hat{\tau}_{resp}(t)$, residual/heart-beat  $\hat{r}_{heart}(t)$, and respiratory phase $\hat{\phi}_{resp}(t)$ signals.
@@
 In this section, we present a kernel regression model to reconstruct the image at any cardiac phase.
-This model can then be used to generate a single-cycle video representative of the subject's heart-beat at a desired higher temporal resolution to achieve temporal super-resolution.
+This model can then be used to generate a single-cycle video representative of the subject's heart-beat at a higher temporal resolution.
 
@@
 Note that this approach will facilitate temporal super-resolution only if the set of images observed in each distinct period of the given image sequence differ in cardiac phase by some non-zero amount.
-Also, the amount of temporal super-resolution possible will depend both on the number of cycles captured and the amount by which one can expect the corresponding images to be perturbed in phase across cycles.
+Also, the amount of temporal super-resolution possible will depend both on the number of cycles captured and the amount corresponding images are perturbed in phase across cycles.
 \vspace{-0.3cm}
@@
 \end{table}
-	We validate our cardiac phase estimation method by comparing its phase estimates with the ECG signal which is the gold standard for cardiac gating.
+	We validate the cardiac phase estimates of our method by comparing them with the ECG signal which is the gold standard for cardiac gating.
 Figure~\ref{fig:instaphase_vs_ecg}(a) shows a head-to-head comparison between the ECG signal and the instantaneous cardiac phase signal obtained using the method described in Section~\ref{sec:method:phase_estimation} for one of the 6 videos.
@@
 Table 1 reports the $R^2$ correlation and the $mean \pm std$ error between the locations of the QRS peaks and the corresponding minima of the cardiac phase signals for all 6 videos in our dataset.
-Figure~\ref{fig:instaphase_vs_ecg}(b) shows a head-to-head comparison of five video frames evenly spaced in time between two consecutive QRS peaks (top-row) and corresponding minima of the cardiac phase signal (bottom-row).
+Figure~\ref{fig:instaphase_vs_ecg}(b) presents a visual comparison of five video frames evenly spaced in time between two consecutive QRS peaks (top-row) and corresponding minima of the cardiac phase signal (bottom-row).
 Notice the high-level of similarity in the images of each column.
@@
 
-	Lastly, since this paper deals with video analysis we encourage the readers to review the supplementary material for a better understanding of the utility and potential of the proposed methods.
+	Lastly, since this paper deals with video analysis we encourage the readers to review the supplementary material for a better understanding of the utility and effectiveness of the proposed methods.
 %
@@
 %
-%
 \vspace{-0.3cm}
@@
 In this paper, we have presented a novel method to estimate the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video.
-We have also presented a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase that can be used to facilitate temporal super-resolution.
-Having demonstrated a proof of concept, in the future, we plan to evaluate our methods on more datasets, especially on cardiac ultrasound videos of subjects with irregular heart beats, and address any remaining pitfalls.
+We have also presented a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
+Having demonstrated a proof of concept, we next plan to evaluate our methods on more datasets, especially of subjects with irregular heart beats, and address any remaining pitfalls.
 One drawback of our phase estimation method is a strong assumption of periodicity which may breakdown in the case of subjects with cardiac arrhythmia.
@@
 Another drawback is that the use normalized correlation to measure inter-frame similarity relies on an inherent assumption that a large part of the image is pulsating.
-To address this, we will try to devise more sophisticated patch-based similarity measures.
+To relax this, we will try to devise more sophisticated local patch-based similarity measures.
 %

### 2017-02-22 paper155.tex
Added reference for super-resolution and put validation tables side by side

@@
 % a short form should be given in case it is too long for the running head
-\titlerunning{Image-based Phase Estimation Method for Cardiac Ultrasound}
+\titlerunning{An Image-based Phase Estimation Method for Cardiac Ultrasound}
 
@@
 We validate our methods using 2D cardiac ultrasound videos and associated ECG recordings of 6 mice.
-%\keywords{Cardiac, Ultrasound, Phase estimation, Gating, Temporal super-resolution}
+%\keywords{Cardiac, Ultrasound, Phase estimation, Gating, Temporal Super-resolution}
 \end{abstract}
@@
 Cardiac ultrasound videos consists of two kinds of periodic motion, one that corresponds to the beating heart motion and the other that corresponds to respiratory motion.
-Knowing the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection\cite{Wick2013}, 3D reconstruction\cite{Wachinger2012}, temporal super-resolution and robotic beating heart surgery\cite{Kurz2015}.
+Knowing the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection\cite{Wick2013}, 3D reconstruction\cite{Wachinger2012}, temporal super-resolution\cite{Cherin2006} and robotic beating heart surgery\cite{Kurz2015}.
 Typically, the position within the cardiac cycle is tracked through ECG data and the position within respiratory cycle is tracked by motion of markers placed on the subject's body\cite{Khamene2004}.
@@
 %
-\caption{Illustration of our phase estimation method: (a) Inter-frame similarity matrix, (b) Matrix of trend components corresponding to respiratory motion, (c) Matrix of residual components corresponding to beating heart motion, (d) Similarity profile chosen for phase estimation with associated trend/respiration and residual/heart-beat components, (e) Periodogram of heart-beat and respiration components along with periodicity characteristics (e.g.
+\caption{Illustration of our phase estimation method: (a) Inter-frame similarity matrix, (b) Trend matrix corresponding to respiratory motion, (c) Residual matrix corresponding to beating heart motion, (d) Similarity profile chosen for phase estimation with associated trend/respiration and residual/heart-beat components, (e) Periodogram of heart-beat and respiration components along with periodicity characteristics (e.g.
 frequency, cycle duration) calculated from the dominant frequency, and (f) Instantaneous cardiac and respiratory phases estimated using the Hilbert transform.}
@@
 
-	Next, considering the narrow-band nature of the trend/respiration and residual/heart-beat signals (see  Figure~\ref{fig:phase_estimation}(e)), we use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous phase of each frame.
+	Next, considering the narrow-band nature of the respiration and heart-beat signals (see  Figure~\ref{fig:phase_estimation}(e)), we use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous phase of each frame.
 Specifically, we compute the instantaneous intra-period phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows: $\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)$ and map $\phi(t)$ from the range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
@@
 Once the cardiac and respiratory phases of each frame have been estimated, it can be used to filter out or guide the extraction of quantitative measurements from a desired part/point of the periodic cycle, a process commonly referred to as gating.
-In this section, we present a robust two-step method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.
+In this section, we present a robust two-step method that uses these phase estimates to gate out video frames with significant respiratory motion.
 
@@
 
-\begin{table}[h]
+	We validate the cardiac phase estimates of our method by comparing them with the ECG signal which is the gold standard for cardiac gating.
+Figure~\ref{fig:instaphase_vs_ecg}(a) shows a head-to-head comparison between the ECG signal and the instantaneous cardiac phase signal obtained using the method described in Section~\ref{sec:method:phase_estimation} for one of the 6 videos.
+Notice how well the locations of the  QRS peaks (red circles) in the ECG signal and the corresponding minima (black circles) of the cardiac phase signal in each cycle match.
+Table 1 reports the $R^2$ correlation and the $mean \pm std$ error between the locations of the QRS peaks and the corresponding minima of the cardiac phase signals for all 6 videos in our dataset.
+Figure~\ref{fig:instaphase_vs_ecg}(b) presents a visual comparison of five video frames evenly spaced in time between two consecutive QRS peaks (top-row) and corresponding minima of the cardiac phase signal (bottom-row).
+Notice the high-level of similarity in the images of each column.
+Figure~\ref{fig:phase_ordering}(b,c) shows the frames of an m-mode line before and after ordering with the estimated cardiac phase.
+Figure~\ref{fig:phase_ordering}(d) shows m-mode frames ordered by cardiac phase after gating out respiratory frames.
+
+	We validate the accuracy of our kernel regression model for reconstructing the image at any cardiac phase using leave-one-out-cross-validation (LOOCV).
+In each round of cross-validation, we randomly exclude one of the non-respiratory frames in the video, fit our kernel-regression model on the remaining frames, use the fitted model to reconstruct the image at the cardiac phase of the excluded frame, and compute the similarity between the reconstructed and original image using normalized correlation.
+The second column in Table~\ref{table:image_reconstruction} shows the mean and standard deviation of normalized correlation between the reconstructed and the original images over 50 rounds of LOCCV for each of the 6 videos in our dataset.
+As a baseline, the third column of  Table~\ref{table:image_reconstruction} shows the mean and standard deviation of the normalized correlation between frames at QRS peaks of the ECG signal for each video.
+We encourage the readers to watch the supplementary videos for a better understanding of the effectiveness and utility of the proposed methods.
+%
+%
+\begin{table}[t]
+\begin{minipage}[b]{0.45\linewidth}
 \centering
@@
 \hline
-\begin{tabular}[c]{@{}c@{}}VID\\ $\;$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}Correlation \\ $R^2$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}QRS peak prediction error\\ $mean \pm std$\end{tabular} \\ \hline
+\begin{tabular}[c]{@{}c@{}}VID\\ $\;$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}Correlation \\ $R^2$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}QRS peak prediction\\ $mean \pm std \; error$\end{tabular} \\ \hline
 1 & 0.9995 & 1.55 $\pm$ 0.89 \\ \hline
@@
 \end{tabular}
-\caption{Comparision between frame positions of QRS peaks in the ECG signal and the corresponding minima of the estimated cardiac phase for 6 videos}
+\caption{Comparision between frame positions of the QRS peaks in ECG signal and the corresponding minima of the estimated cardiac phase for 6 videos.}
 \label{table:phase_estimation}
-\end{table}
-	We validate the cardiac phase estimates of our method by comparing them with the ECG signal which is the gold standard for cardiac gating.
-Figure~\ref{fig:instaphase_vs_ecg}(a) shows a head-to-head comparison between the ECG signal and the instantaneous cardiac phase signal obtained using the method described in Section~\ref{sec:method:phase_estimation} for one of the 6 videos.
-Notice how well the locations of the peaks of the QRS complex (red circles) in the ECG signal and the corresponding minima (black circles) of the cardiac phase signal in each cycle match.
-Table 1 reports the $R^2$ correlation and the $mean \pm std$ error between the locations of the QRS peaks and the corresponding minima of the cardiac phase signals for all 6 videos in our dataset.
-Figure~\ref{fig:instaphase_vs_ecg}(b) presents a visual comparison of five video frames evenly spaced in time between two consecutive QRS peaks (top-row) and corresponding minima of the cardiac phase signal (bottom-row).
-Notice the high-level of similarity in the images of each column.
-Figure~\ref{fig:phase_ordering}(b,c) shows the frames of an m-mode line before and after ordering with the estimated cardiac phase.
-Figure~\ref{fig:phase_ordering}(d) shows the phase-ordered m-mode frames after performing respiratory gating as described in Section~\ref{sec:method:gating}.
-
-	We validate the accuracy of our kernel regression model for reconstructing the image at any cardiac phase using leave-one-out-cross-validation (LOOCV).
-In each round of cross-validation, we randomly exclude one of the non-respiratory frames in the video, fit our kernel-regression model on the remaining frames, use the fitted model to reconstruct the image at the cardiac phase of the excluded frame, and compute the similarity between the reconstructed and original image using normalized correlation.
-The second column in Table~\ref{table:image_reconstruction} shows the mean and standard deviation of normalized correlation between the reconstructed and the original images over 50 rounds of LOCCV for each of the 6 videos in our dataset.
-As a baseline, the third column of  Table~\ref{table:image_reconstruction} shows the mean and standard deviation of the normalized correlation between frames at QRS peaks of the ECG signal for each video.
-
-	Lastly, since this paper deals with video analysis we encourage the readers to review the supplementary material for a better understanding of the utility and effectiveness of the proposed methods.
+\end{minipage}
 %
-\begin{table}[h]
+\hspace{0.5cm}
+%
+\begin{minipage}[b]{0.45\linewidth}
 \centering
@@
 \hline
-\begin{tabular}[c]{@{}c@{}}VID \\ $\;$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}LOOCV (50) NCORR\\ $mean \pm std$\end{tabular} & \begin{tabular}[c]{@{}c@{}}QRS Peak Frame NCORR\\ $mean \pm std$\end{tabular} \\ \hline
+\begin{tabular}[c]{@{}c@{}}VID \\ $\;$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}LOOCV (50)\\ $mean \pm std \; ncorr$\end{tabular} & \begin{tabular}[c]{@{}c@{}}QRS Peak Frames\\ $mean \pm std \; ncorr$\end{tabular} \\ \hline
 1 & 0.8562 $\pm$ 0.0366 & 0.7981 $\pm$ 0.0532 \\ \hline
@@
 \end{tabular}
-\caption{Evaluation of our kernel regression model for reconstructing the image at any cardiac phase using 50 rounds of Leave one out cross-validation (LOOCV)}
+\caption{Evaluation of our kernel regression model for reconstructing the image at any cardiac phase using 50 rounds of Leave one out cross-validation (LOOCV).}
 \label{table:image_reconstruction}
+\end{minipage}
 \end{table}
+%
 %

### 2017-02-22 paper155.tex
Moved table captions to the top
Increased negative vspace to try and fit into 8 pages

@@
 %
-\vspace{-0.3cm}
+\vspace{-0.5cm}
 \section{Method}
@@
 \end{IonFigT}
-\vspace{-0.3cm}
+\vspace{-0.5cm}
 \subsection{Estimation of instantaneous cardiac and respiratory phases}
@@
 %
-\vspace{-0.3cm}
+\vspace{-0.5cm}
 \subsection{Gating out respiratory frames}
@@
 %
-\vspace{-0.3cm}
+\vspace{-0.5cm}
 \subsection{Model to generate images by cardiac phase for super-resolution}
@@
 Its bandwidth $\sigma_{L}$ is set equal to a constant $k_L = 2.0$ times the robust estimate of standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit as described in Section~\ref{sec:method:gating}.
-
-	A representative single-cycle video of the subject's heart-beat can now be reconstructed by sampling the manifold $M(\phi)$ at any desired resolution from the range $[0, 1)$ of the cardiac phase.
-Note that this approach will facilitate temporal super-resolution only if the set of images observed in each distinct period of the given image sequence differ in cardiac phase by some non-zero amount.
-Also, the amount of temporal super-resolution possible will depend both on the number of cycles captured and the amount corresponding images are perturbed in phase across cycles.
-\vspace{-0.3cm}
+A single-cycle video representative of the subject's heart-beat can now be reconstructed by sampling the manifold $M(\phi)$ at any desired resolution.
+\vspace{-0.5cm}
 \section{Results}
@@
 \centering
+\caption{Comparision between frame positions of the QRS peaks in ECG signal and the corresponding minima of the estimated cardiac phase for 6 videos.}
 \begin{tabular}{|c|c|c|c|}
@@
 \end{tabular}
-\caption{Comparision between frame positions of the QRS peaks in ECG signal and the corresponding minima of the estimated cardiac phase for 6 videos.}
 \label{table:phase_estimation}
@@
 \begin{minipage}[b]{0.45\linewidth}
+\caption{Evaluation of our kernel regression model for reconstructing the image at any cardiac phase using 50 rounds of Leave one out cross-validation (LOOCV).}
 \centering
@@
 \end{tabular}
-\caption{Evaluation of our kernel regression model for reconstructing the image at any cardiac phase using 50 rounds of Leave one out cross-validation (LOOCV).}
 \label{table:image_reconstruction}
 \end{minipage}
+\vspace{-0.3cm}
 \end{table}
@@
 %
-\vspace{-0.3cm}
+\vspace{-0.5cm}
 \section{Conclusion}
@@
 %
-\vspace{-0.3cm}
+\vspace{-0.5cm}
 \bibliographystyle{splncs03}

### 2017-02-22 paper155.tex
Updated paper

@@
 	Next, considering the narrow-band nature of the respiration and heart-beat signals (see  Figure~\ref{fig:phase_estimation}(e)), we use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous phase of each frame.
-Specifically, we compute the instantaneous intra-period phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows: $\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)$ and map $\phi(t)$ from the range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
+Specifically, we compute the instantaneous intra-period phase $\phi(t) \in \left [  -\pi, \pi\right )$ of a periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows: $\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)$ and map $\phi(t)$ from the range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
 Let $\hat{\phi}_{heart}(t)$ and $\hat{\phi}_{resp}(t)$ denote the instantaneous cardiac and respiratory phases computed from the trend/respiration $\hat{\tau}_{resp}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signals, respectively.

### 2017-02-22 paper155.tex
Made a few corrections and removed a few lines to reduce text

@@
 \begin{abstract}
-Ultrasound is emerging as an increasingly effective tool for rapid non-invasive visual assessment of cardiac structure and function.
-Knowing the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications.
-In this paper, we present a novel method for the estimation of instantaneous cardiac and respiratory phases directly from cardiac ultrasound videos.
-The method transforms the complex high-dimensional image sequence into a univariate time series with the same periodicity characteristics, decouples the periodic sources of beating heart and respiratory motion using a trend extraction technique, and estimates the cardiac and respiratory phases using Hilbert transform.
-We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
-We validate our methods using 2D cardiac ultrasound videos and associated ECG recordings of 6 mice.
+Ultrasound is emerging as an increasingly effective tool for rapid non-invasive assessment of cardiac structure and function.
+Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in many applications.
+In this paper, we present a novel method for estimation of instantaneous cardiac and respiratory phases directly from cardiac ultrasound videos.
+The method transforms the complex high-dimensional image sequence into a univariate time series with the same periodicity characteristics, decouples the periodic sources of beating heart and respiratory motion using a trend extraction technique, and estimates the cardiac and respiratory phases using the Hilbert transform.
+We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model that reconstructs images at any cardiac phase and facilitates temporal super-resolution.
+We validate our methods using 2D cardiac ultrasound videos and ECG recordings of 6 mice.
 %\keywords{Cardiac, Ultrasound, Phase estimation, Gating, Temporal Super-resolution}
@@
 %
-Cardiovascular disease is the leading cause of death worldwide and ultrasound is emerging as an increasingly effective tool for rapid non-invasive visual assessment of cardiac structure and function.
-Cardiac ultrasound videos consists of two kinds of periodic motion, one that corresponds to the beating heart motion and the other that corresponds to respiratory motion.
-Knowing the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection\cite{Wick2013}, 3D reconstruction\cite{Wachinger2012}, temporal super-resolution\cite{Cherin2006} and robotic beating heart surgery\cite{Kurz2015}.
-Typically, the position within the cardiac cycle is tracked through ECG data and the position within respiratory cycle is tracked by motion of markers placed on the subject's body\cite{Khamene2004}.
+Cardiovascular disease is the leading cause of death worldwide and ultrasound is emerging as an increasingly effective tool for rapid non-invasive assessment of cardiac structure and function.
+Cardiac ultrasound videos consist of two kinds of periodic motion, one corresponding to beating heart motion and the other corresponding to respiratory motion.
+Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection\cite{Wick2013}, 3D reconstruction\cite{Wachinger2012}, temporal super-resolution\cite{Cherin2006} and robotic beating heart surgery\cite{Kurz2015}.
+Typically, the position within cardiac cycle is tracked through ECG data and the position within respiratory cycle is tracked by motion of markers placed on the subject's body\cite{Khamene2004}.
 
@@
 %
-In this section, we present the theory underlying the proposed methods for instantaneous phase estimation, gating and temporal super-resolution of cardiac ultrasound videos.
+In this section, we present the theory underlying the proposed methods.
 Specifically, we begin by describing our method for the estimation of instantaneous cardiac and respiratory phases in Section \ref{sec:method:phase_estimation}.
 We then present a robust method that uses these phase estimates to gate out video frames with predominant respiratory motion in Section \ref{sec:method:gating}.
-Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression model to generate the image at any cardiac phase that can be used to achieve temporal super-resolution.
+Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression model for reconstructing images at any cardiac phase and facilitate temporal super-resolution.
 %
@@
 
-	We first compute the similarity between all pairs of frames in the given periodic image sequence containing $N$ images/frames to create a matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames.
+	We first compute the similarity between all pairs of images/frames in the given periodic image sequence containing $N$ frames to create a symmetric matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames.
 Here in, we use normalization correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity but, in principle, any of the similarity metrics used in image registration algorithms can potentially be chosen.
 Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series.
-Furthermore, If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted, this time series can be expected to preserve the periodicity characteristics of the original image sequence.
+If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted, this time series can be expected to preserve the periodicity characteristics of the original image sequence.
 Two kinds of periodic motion are present in cardiac ultrasound videos, one corresponding to the beating heart motion and the other corresponding to respiratory motion.
@@
 
-	Next, we use a trend extraction technique~\cite{Alexandrov2012} called the Hodrick-Prescott (HP) filter~\cite{Hodrick1997}, widely used in econometrics, to decompose the frame similarity signal $u(t)$ corresponding to each row of $A$ into a sum of two components: (i) Trend component $\tau_{resp}(t)$ encompassing the periodicity characteristic of only the lower-frequency respiratory motion, and (ii) Residual component $r_{heart}(t)$ encompassing the periodicity characteristic of only the higher-frequency beating heart motion.
+	Next, we use a trend extraction technique~\cite{Alexandrov2012} called the Hodrick-Prescott (HP) filter~\cite{Yamada2015}, widely used in econometrics, to decompose the frame similarity signal $u(t)$ corresponding to each row of $A$ into a sum of two components: (i) Trend component $\tau_{resp}(t)$ encompassing the periodicity characteristic of only the lower-frequency respiratory motion, and (ii) Residual component $r_{heart}(t)$ encompassing the periodicity characteristic of only the higher-frequency beating heart motion.
 The HP filter performs the decomposition of $u(t) = \tau_{resp}(t) + r_{heart}(t)$ by solving the following optimization problem:
@@
 \end{equation}
-where $\nabla^2\tau_{resp}(t) = \tau_{resp}(t+1) - 2 \tau_{resp}(t) + \tau_{resp}(t-1)$ is the second-order difference or derivative that penalizes curvature of the trend signal.
-We use a ridge-regression implementation of the HP filter~\cite{Yamada2015} with $\lambda=6400$.
+where $\nabla^2\tau_{resp}(t) = \tau_{resp}(t+1) - 2 \tau_{resp}(t) + \tau_{resp}(t-1)$ is the second-order difference or derivative that penalizes curvature of the trend signal and $\lambda=6400$ specifies the strength of this penalty.
 Let $A_{resp}$ and $A_{heart}$ be the matrices whose rows contain the trend/respiratory and residual/heart-beat components, respectively, of the frame similarity signals in the corresponding rows of the similarity matrix $A$.
-Figures~\ref{fig:phase_estimation}(b,c) show the trend $A_{res[}$ and residual $A_{heart}$ matrices for one of our cardiac ultrasound videos.
+Figures~\ref{fig:phase_estimation}(b,c) show the trend $A_{resp}$ and residual $A_{heart}$ matrices for one of our cardiac ultrasound videos.
 To minimize the effect of any noise on phase estimation, we pick frame similarity signal corresponding to the row of $A_{heart}$ whose periodogram or power-frequency distribution has minimum entropy.
@@
 %
-\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/respiratory_phase_gating.png}
-\ionbox{5.0in}\\
-\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/robust_lowess_gating.png}
-\ionbox{5.0in}\\
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/respiratory_phase_gating.png}
+\ionbox{4.5in}\\
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/robust_lowess_gating.png}
+\ionbox{4.5in}\\
 %
-\caption{Illustration of our two-step method for gating out respiratory frames: (a) Frames $F_{cutoff}$ discarded (blue circles) in the first step overlaid with the trend/respiration, residual/heart-beat, and instantaneous respiratory phase signals (b) residual/heart-beat signal $\hat{r}_{heart}(t)$ vs cardiac phase $\hat{\phi}_{heart}(t)$ overlaid with frames $F_{cutoff}$ discarded in step-1 (blue circle), LOWESS fit $L(\phi)$ (black solid), upper and lower bounds or 95\% confidence interval $(L(\phi) \pm 2.0 * \hat{\sigma}_L)$ of valid non-respiratory frames (black dotted), and frames $F_{resp}$ (red circles) gated at the end of step-2.}
+\caption{Illustration of our method for gating out respiratory frames: (a) Frames $F_{cutoff}$ discarded (blue circles) in the first step overlaid with the trend/respiration, residual/heart-beat, and instantaneous respiratory phase signals (b) residual/heart-beat signal $\hat{r}_{heart}(t)$ vs cardiac phase $\hat{\phi}_{heart}(t)$ overlaid with frames $F_{cutoff}$ discarded in step-1 (blue circle), LOWESS fit $L(\phi)$ (black solid), upper and lower bounds or 95\% confidence interval $(L(\phi) \pm 2.0 * \hat{\sigma}_L)$ of valid non-respiratory frames (black dotted), and frames $F_{resp}$ (red circles) gated at the end of step-2.}
 \label{fig:respiratory_gating}
@@
 %
-Once the cardiac and respiratory phases of each frame have been estimated, it can be used to filter out or guide the extraction of quantitative measurements from a desired part/point of the periodic cycle, a process commonly referred to as gating.
+Once the cardiac and respiratory phases of each frame have been estimated, they can be used to filter out or guide the extraction of quantitative measurements from a desired part/point of the periodic cycle, a process commonly referred to as gating.
 In this section, we present a robust two-step method that uses these phase estimates to gate out video frames with significant respiratory motion.
 
-	In the first step, based on the observation that respiratory motion mainly occurs around the minima of the trend/respiration signal $\hat{\tau}_{resp}(t)$ (pink curve in Fig.~\ref{fig:phase_estimation}(d)) with respiratory phase $\hat{\phi}_{resp}(t) = 0$, we perform a rough initial gating by discarding the frames $F_{cutoff} = \left \{ t \mid \hat{\phi}_{resp}(t) < c \parallel \hat{\phi}_{resp}(t) > (1 - c) \right \}$ whose respiratory phase distance from $\hat{\phi}_{resp}(t) = 0$ is below a specified cutoff value $c=0.2$.
-Figure~\ref{fig:respiratory_gating}(a) shows the discarded frames $F_{cutoff}$ overlaid (blue circles) on the trend/respiration $\hat{\tau}_{resp}(t)$, residual/heart-beat  $\hat{r}_{heart}(t)$, and respiratory phase $\hat{\phi}_{resp}(t)$ signals.
+	In the first step, based on the observation that respiratory motion mainly occurs around the minima of the trend/respiration signal $\hat{\tau}_{resp}(t)$ (pink curve in Fig.~\ref{fig:phase_estimation}(d)) with respiratory phase $\hat{\phi}_{resp}(t) = 0$, we perform a rough initial gating by discarding the frames $F_{cutoff} = \left \{ t \mid \hat{\phi}_{resp}(t) < c \parallel \hat{\phi}_{resp}(t) > (1 - c) \right \}$ whose phase distance from $\hat{\phi}_{resp}(t) = 0$ is below a specified cutoff value $c=0.2$.
+Figure~\ref{fig:respiratory_gating}(a) shows the discarded frames $F_{cutoff}$ overlaid on the trend/respiration $\hat{\tau}_{resp}(t)$, residual/heart-beat  $\hat{r}_{heart}(t)$, and respiratory phase $\hat{\phi}_{resp}(t)$ signals.
 
@@
 Figure~\ref{fig:phase_ordering}(d) shows m-mode frames ordered by cardiac phase after gating out respiratory frames.
+The supplementary material includes a video showing the estimated cardiac and respiratory phases of each frame for one of our datasets.
 
@@
 As a baseline, the third column of  Table~\ref{table:image_reconstruction} shows the mean and standard deviation of the normalized correlation between frames at QRS peaks of the ECG signal for each video.
-We encourage the readers to watch the supplementary videos for a better understanding of the effectiveness and utility of the proposed methods.
+The supplementary material includes single-cycle videos at 2x and 4x temporal resolution generated using the kernel regression model wherein the motion of the heart chambers and valves are much clear compared to the original video.
 %
@@
 \begin{minipage}[b]{0.45\linewidth}
-\caption{Evaluation of our kernel regression model for reconstructing the image at any cardiac phase using 50 rounds of Leave one out cross-validation (LOOCV).}
+\caption{Evaluation of our kernel regression model for reconstructing images at any cardiac phase using 50 rounds of Leave one out cross-validation (LOOCV).}
 \centering
@@
 To relax this, we will try to devise more sophisticated local patch-based similarity measures.
+We encourage the readers to watch the supplementary videos for a better understanding of the effectiveness and utility of the proposed methods.
 %

### 2017-02-22 paper155.tex
Updated paper

@@
 \ionbox{1.33in}\ionbox{1.33in}\ionbox{1.33in}\\
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/season_trend_decomposition.png}
-\ionbox{4.5in}\\
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/periodogram.png}
-\ionbox{4.5in}\\
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instantaneous_phase.png}
-\ionbox{4.5in}\\
+\includegraphics[width=4.25in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/season_trend_decomposition.png}
+\ionbox{4.25in}\\
+\includegraphics[width=4.25in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/periodogram.png}
+\ionbox{4.25in}\\
+\includegraphics[width=4.25in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instantaneous_phase.png}
+\ionbox{4.25in}\\
 %
@@
 %
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/respiratory_phase_gating.png}
-\ionbox{4.5in}\\
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/robust_lowess_gating.png}
-\ionbox{4.5in}\\
+\includegraphics[width=4.25in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/respiratory_phase_gating.png}
+\ionbox{4.25in}\\
+\includegraphics[width=4.25in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/robust_lowess_gating.png}
+\ionbox{4.25in}\\
 %
@@
 %
-\caption{Visualization of m-mode frames ordered by estimated cardiac phase: (a) One of the frames in the video overlaid with the M-mode line shown in the next three images, (b) M-mode frames in the order they appear in the input video, (c) M-mode frames ordered by estimated cardiac phase, and (d) Non-respiratory m-mode frames ordered by estimated cardiac phase.}
+\caption{Visualization of m-mode frames ordered by cardiac phase: (a) One of the frames in the video overlaid with the M-mode line shown in the next three images, (b) M-mode frames in the order they appear in the input video, (c) M-mode frames ordered by cardiac phase, and (d) Non-respiratory m-mode frames ordered by cardiac phase.}
 \label{fig:phase_ordering}
@@
 	We validate the cardiac phase estimates of our method by comparing them with the ECG signal which is the gold standard for cardiac gating.
-Figure~\ref{fig:instaphase_vs_ecg}(a) shows a head-to-head comparison between the ECG signal and the instantaneous cardiac phase signal obtained using the method described in Section~\ref{sec:method:phase_estimation} for one of the 6 videos.
+Figure~\ref{fig:instaphase_vs_ecg}(a) shows a head-to-head comparison between the ECG signal and the cardiac phase signal obtained using the method described in Section~\ref{sec:method:phase_estimation} for one of the 6 videos.
 Notice how well the locations of the  QRS peaks (red circles) in the ECG signal and the corresponding minima (black circles) of the cardiac phase signal in each cycle match.
-Table 1 reports the $R^2$ correlation and the $mean \pm std$ error between the locations of the QRS peaks and the corresponding minima of the cardiac phase signals for all 6 videos in our dataset.
+Table 1 reports the $R^2$ correlation and the $mean \pm std$ error between the locations of the QRS peaks and the corresponding minima of the cardiac phase signals for all 6 videos.
 Figure~\ref{fig:instaphase_vs_ecg}(b) presents a visual comparison of five video frames evenly spaced in time between two consecutive QRS peaks (top-row) and corresponding minima of the cardiac phase signal (bottom-row).
@@
 
-	We validate the accuracy of our kernel regression model for reconstructing the image at any cardiac phase using leave-one-out-cross-validation (LOOCV).
+	We validate the accuracy of our kernel regression model for reconstructing images at any cardiac phase using leave-one-out-cross-validation (LOOCV).
 In each round of cross-validation, we randomly exclude one of the non-respiratory frames in the video, fit our kernel-regression model on the remaining frames, use the fitted model to reconstruct the image at the cardiac phase of the excluded frame, and compute the similarity between the reconstructed and original image using normalized correlation.
-The second column in Table~\ref{table:image_reconstruction} shows the mean and standard deviation of normalized correlation between the reconstructed and the original images over 50 rounds of LOCCV for each of the 6 videos in our dataset.
-As a baseline, the third column of  Table~\ref{table:image_reconstruction} shows the mean and standard deviation of the normalized correlation between frames at QRS peaks of the ECG signal for each video.
-The supplementary material includes single-cycle videos at 2x and 4x temporal resolution generated using the kernel regression model wherein the motion of the heart chambers and valves are much clear compared to the original video.
+The second column of Table~\ref{table:image_reconstruction} shows the $mean \pm std$ of normalized correlation between the reconstructed and the original images over 50 rounds of LOCCV for all 6 videos.
+As a baseline, the third column of  Table~\ref{table:image_reconstruction} shows the $mean \pm std$ of the normalized correlation between frames at QRS peaks of the ECG signal.
+The supplementary material includes single-cycle videos at 2x and 4x temporal resolution generated using the kernel regression model wherein the motion of heart chambers and valves is much clear than the original video.
 %
@@
 \begin{minipage}[b]{0.45\linewidth}
-\caption{Evaluation of our kernel regression model for reconstructing images at any cardiac phase using 50 rounds of Leave one out cross-validation (LOOCV).}
+\caption{Evaluation of our kernel regression model for reconstructing images at any cardiac phase using 50 rounds of Leave one out cross-validation (LOOCV) for 6 videos.}
 \centering

### 2017-02-22 paper155.tex
Reviewed some previous work at the end of the introductory section

@@
 We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model that reconstructs images at any cardiac phase and facilitates temporal super-resolution.
-We validate our methods using 2D cardiac ultrasound videos and ECG recordings of 6 mice.
-%\keywords{Cardiac, Ultrasound, Phase estimation, Gating, Temporal Super-resolution}
+We validate our methods using cardiac ultrasound or echocardiography videos and ECG recordings of 6 mice.
+%\keywords{Cardiac, Ultrasound, Echocardiography, Phase estimation, Gating, Temporal Super-resolution}
+\vspace{-0.5cm}
 \end{abstract}
 %
-\vspace{-0.5cm}
 \section{Introduction}
@@
 Cardiac ultrasound videos consist of two kinds of periodic motion, one corresponding to beating heart motion and the other corresponding to respiratory motion.
-Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection\cite{Wick2013}, 3D reconstruction\cite{Wachinger2012}, temporal super-resolution\cite{Cherin2006} and robotic beating heart surgery\cite{Kurz2015}.
-Typically, the position within cardiac cycle is tracked through ECG data and the position within respiratory cycle is tracked by motion of markers placed on the subject's body\cite{Khamene2004}.
+Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection\cite{Wick2013}, 3D reconstruction\cite{Wachinger2012}, and temporal super-resolution\cite{Cherin2006}.
+Typically, the position within cardiac cycle is tracked through ECG data and position within respiratory cycle is tracked by motion of markers placed on the subject's body\cite{Khamene2004}.
 
-	In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the the cardiac ultrasound video.
+	In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video.
 We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression to reconstruct images at any cardiac phase to facilitate temporal super-resolution.
+Previous work on the estimation of cardiac and/or respiratory phases directly from ultrasound videos is very limited.
+In~\cite{Karadayi2006}, the authors compute a signal measuring the x- or y-coordinate of the center of mass of each frame, use bandpass filtering to remove components outside typical cardiac frequency range, find the dominant frequency and apply matched sine/cosine filters to estimate the instantaneous phase.
+In~\cite{Sundar2009}, the authors compute a signal measuring phase correlation between consecutive frames, use a bandpass filter centered at 1Hz to recover the cardiac signal, and use a low pass filter to recover the respiratory signal.
 %
@@
 Specifically, we begin by describing our method for the estimation of instantaneous cardiac and respiratory phases in Section \ref{sec:method:phase_estimation}.
-We then present a robust method that uses these phase estimates to gate out video frames with predominant respiratory motion in Section \ref{sec:method:gating}.
+We then present a robust method that uses these phase estimates to gate out video frames with significant respiratory motion in Section \ref{sec:method:gating}.
 Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression model for reconstructing images at any cardiac phase and facilitate temporal super-resolution.
@@
 	We first compute the similarity between all pairs of images/frames in the given periodic image sequence containing $N$ frames to create a symmetric matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames.
-Here in, we use normalization correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity but, in principle, any of the similarity metrics used in image registration algorithms can potentially be chosen.
+Here in, we use normalized correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity but, in principle, any of the similarity metrics used in image registration algorithms can potentially be chosen.
 Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series.
@@
 \begin{equation}	
-\argmin{\tau(t)} \left[ \sum_{t=1}^{N}  \left(u(t) - \tau_{resp}(t) \right)^2  + \lambda \sum_{t=1}^{N-1} \left( \nabla^2 \tau_{resp}(t) \right)^2 \right] 
+\argmin{\tau_{resp}(t)} \left[ \sum_{t=1}^{N}  \left(u(t) - \tau_{resp}(t) \right)^2  + \lambda \sum_{t=1}^{N-1} \left( \nabla^2 \tau_{resp}(t) \right)^2 \right] 
 \end{equation}
-where $\nabla^2\tau_{resp}(t) = \tau_{resp}(t+1) - 2 \tau_{resp}(t) + \tau_{resp}(t-1)$ is the second-order difference or derivative that penalizes curvature of the trend signal and $\lambda=6400$ specifies the strength of this penalty.
+where $\nabla^2\tau_{resp}(t) = \tau_{resp}(t+1) - 2 \tau_{resp}(t) + \tau_{resp}(t-1)$ is the second-order difference or derivative of the trend signal and $\lambda=6400$ is penalty parameter.
 Let $A_{resp}$ and $A_{heart}$ be the matrices whose rows contain the trend/respiratory and residual/heart-beat components, respectively, of the frame similarity signals in the corresponding rows of the similarity matrix $A$.
@@
 Next, we compute a robust estimate of the standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit based on the median absolute value of the difference between the LOWESS fit $L( \hat{\phi}_{heart}(t) )$ and residual/heart-beat signal $\hat{r}_{heart}(t)$ for all frames.
-Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert L( \hat{\phi}_{heart}(t) ) - \hat{r}_{heart}(t)  \rvert   > k \times \hat{\sigma}_{L}  \right \}$ for which the absolute  difference between the LOWESS fit and heart-beat signal value is greater than $k = 0.2$ times the standard deviation $\hat{\sigma}_{L}$.
+Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert L( \hat{\phi}_{heart}(t) ) - \hat{r}_{heart}(t)  \rvert   > k \times \hat{\sigma}_{L}  \right \}$ for which the absolute  difference between the LOWESS fit and heart-beat signal value is greater than $k = 2.0$ (95\% confidence interval) times the standard deviation $\hat{\sigma}_{L}$.
 %

### 2017-02-23 paper155.tex
Made QRS peak identification table look better
Playing with periodogram figue
Updated paper

@@
 The method transforms the complex high-dimensional image sequence into a univariate time series with the same periodicity characteristics, decouples the periodic sources of beating heart and respiratory motion using a trend extraction technique, and estimates the cardiac and respiratory phases using the Hilbert transform.
-We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model that reconstructs images at any cardiac phase and facilitates temporal super-resolution.
+We also present a robust non-parametric regression technique for gating out respiratory frames and a kernel regression model that reconstructs images at any cardiac phase and facilitates temporal super-resolution.
 We validate our methods using cardiac ultrasound or echocardiography videos and ECG recordings of 6 mice.
@@
 Cardiac ultrasound videos consist of two kinds of periodic motion, one corresponding to beating heart motion and the other corresponding to respiratory motion.
-Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection\cite{Wick2013}, 3D reconstruction\cite{Wachinger2012}, and temporal super-resolution\cite{Cherin2006}.
+Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection~\cite{Wick2013}, 3D reconstruction~\cite{Wachinger2012}, and temporal super-resolution~\cite{Cherin2006}.
 Typically, the position within cardiac cycle is tracked through ECG data and position within respiratory cycle is tracked by motion of markers placed on the subject's body\cite{Khamene2004}.
@@
 In~\cite{Karadayi2006}, the authors compute a signal measuring the x- or y-coordinate of the center of mass of each frame, use bandpass filtering to remove components outside typical cardiac frequency range, find the dominant frequency and apply matched sine/cosine filters to estimate the instantaneous phase.
-In~\cite{Sundar2009}, the authors compute a signal measuring phase correlation between consecutive frames, use a bandpass filter centered at 1Hz to recover the cardiac signal, and use a low pass filter to recover the respiratory signal.
+In~\cite{Sundar2009}, the authors compute a signal measuring phase correlation between consecutive frames, use a bandpass filter to recover the cardiac signal and a low pass filter to recover the respiratory signal.
 %
@@
 In this section, we present the theory underlying the proposed methods.
-Specifically, we begin by describing our method for the estimation of instantaneous cardiac and respiratory phases in Section \ref{sec:method:phase_estimation}.
-We then present a robust method that uses these phase estimates to gate out video frames with significant respiratory motion in Section \ref{sec:method:gating}.
+Specifically, we begin by describing our method for estimation of instantaneous cardiac and respiratory phases in Section \ref{sec:method:phase_estimation}.
+We then present a robust method to gate out video frames with significant respiratory motion in Section \ref{sec:method:gating}.
 Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression model for reconstructing images at any cardiac phase and facilitate temporal super-resolution.
@@
 %
-While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting such as the case of periodic cardiac ultrasound videos wherein several thousands of variables (pixel intensities) are involved.
-Our strategy was to find a way to transforming this complex multi-variate problem into a univariate one and take advantage of existing univariate methods to solve the problem.
+While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data~\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting such as the case of periodic cardiac ultrasound videos wherein several thousands of variables (pixel intensities) are involved.
+Our strategy is to transform this complex multi-variate problem into a univariate one and take advantage of existing methods to solve the problem.
 
@@
 Next, we compute a robust estimate of the standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit based on the median absolute value of the difference between the LOWESS fit $L( \hat{\phi}_{heart}(t) )$ and residual/heart-beat signal $\hat{r}_{heart}(t)$ for all frames.
-Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert L( \hat{\phi}_{heart}(t) ) - \hat{r}_{heart}(t)  \rvert   > k \times \hat{\sigma}_{L}  \right \}$ for which the absolute  difference between the LOWESS fit and heart-beat signal value is greater than $k = 2.0$ (95\% confidence interval) times the standard deviation $\hat{\sigma}_{L}$.
-%
-\begin{IonFigT}
-\centering
-%
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/phaseordered.png}
-\ionbox{1.2in}\ionbox{1.1in}\ionbox{1.1in}\ionbox{1.1in}\\
-%
-\caption{Visualization of m-mode frames ordered by cardiac phase: (a) One of the frames in the video overlaid with the M-mode line shown in the next three images, (b) M-mode frames in the order they appear in the input video, (c) M-mode frames ordered by cardiac phase, and (d) Non-respiratory m-mode frames ordered by cardiac phase.}
-\label{fig:phase_ordering}
-\end{IonFigT}
+Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert \hat{r}_{heart}(t) - L( \hat{\phi}_{heart}(t) ) \rvert   > k \times \hat{\sigma}_{L}  \right \}$ for which the absolute  difference between the LOWESS fit and heart-beat signal value is greater than $k = 2.0$ (95\% confidence interval) times the standard deviation $\hat{\sigma}_{L}$.
 %
@@
 A single-cycle video representative of the subject's heart-beat can now be reconstructed by sampling the manifold $M(\phi)$ at any desired resolution.
+%
+%
+\begin{IonFigT}
+\centering
+%
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/phaseordered.png}
+\ionbox{1.2in}\ionbox{1.1in}\ionbox{1.1in}\ionbox{1.1in}\\
+%
+\caption{Visualization of m-mode frames ordered by cardiac phase: (a) One of the frames in the video overlaid with the M-mode line shown in the next three images, (b) M-mode frames in the order they appear in the input video, (c) M-mode frames ordered by cardiac phase, and (d) Non-respiratory m-mode frames ordered by cardiac phase.}
+\label{fig:phase_ordering}
+\end{IonFigT}
+%
+%
 \vspace{-0.5cm}
@@
 \begin{table}[t]
-\begin{minipage}[b]{0.45\linewidth}
+\begin{minipage}[t]{0.45\linewidth}
 \centering
 \caption{Comparision between frame positions of the QRS peaks in ECG signal and the corresponding minima of the estimated cardiac phase for 6 videos.}
-\begin{tabular}{|c|c|c|c|}
+\begin{tabular}{|c|c|c|}
 \hline
-\begin{tabular}[c]{@{}c@{}}VID\\ $\;$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}Correlation \\ $R^2$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}QRS peak prediction\\ $mean \pm std \; error$\end{tabular} \\ \hline
-1 & 0.9995 & 1.55 $\pm$ 0.89 \\ \hline
-2 & 0.9995 & 1.36 $\pm$ 1.15 \\ \hline
-3 & 0.9996 & 1.45 $\pm$ 0.89 \\ \hline
-4 & 0.9998 & 0.73 $\pm$ 0.62 \\ \hline
-5 & 0.9990 & 2.09 $\pm$ 1.78 \\ \hline
-6 & 0.9992 & 1.40 $\pm$ 1.74 \\ \hline
+\multicolumn{1}{|l|}{} & \multicolumn{2}{l|}{QRS peak frame identification} \\ \hline
+VID & $R^2$ & $mean \pm std \; error$ \\ \hline
+1 & \; 0.9995 \; & 1.55 $\pm$ 0.89 \\ \hline
+2 & \; 0.9995 \; & 1.36 $\pm$ 1.15 \\ \hline
+3 & \; 0.9996 \; & 1.45 $\pm$ 0.89 \\ \hline
+4 & \; 0.9998 \; & 0.73 $\pm$ 0.62 \\ \hline
+5 & \; 0.9990 \; & 2.09 $\pm$ 1.78 \\ \hline
+6 & \; 0.9992 \; & 1.40 $\pm$ 1.74 \\ \hline
 \end{tabular}
@@
 %
-\begin{minipage}[b]{0.45\linewidth}
+\begin{minipage}[t]{0.45\linewidth}
 \caption{Evaluation of our kernel regression model for reconstructing images at any cardiac phase using 50 rounds of Leave one out cross-validation (LOOCV) for 6 videos.}

### 2017-02-23 paper155.tex
ENH: Edits

--- 2017-02-22 paper155.tex
@@
 Ultrasound is emerging as an increasingly effective tool for rapid non-invasive assessment of cardiac structure and function.
-Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in many applications.
+Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in many applications and is particularly challenging in pre-clinical studies involving small animals with rapid heartrates.
+
 In this paper, we present a novel method for estimation of instantaneous cardiac and respiratory phases directly from cardiac ultrasound videos.
@@
 Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection\cite{Wick2013}, 3D reconstruction\cite{Wachinger2012}, and temporal super-resolution\cite{Cherin2006}.
+These tasks are particularly common and challenging in pre-clinical studies involving small animals such as mice that have heartrates greater than 400 beats per minute (BPM) and yet are difficult to instrument with electronic cardiac and respirator monitors.
 Typically, the position within cardiac cycle is tracked through ECG data and position within respiratory cycle is tracked by motion of markers placed on the subject's body\cite{Khamene2004}.
@@
 We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression to reconstruct images at any cardiac phase to facilitate temporal super-resolution.
+
   Previous work on the estimation of cardiac and/or respiratory phases directly from ultrasound videos is very limited.
-In~\cite{Karadayi2006}, the authors compute a signal measuring the x- or y-coordinate of the center of mass of each frame, use bandpass filtering to remove components outside typical cardiac frequency range, find the dominant frequency and apply matched sine/cosine filters to estimate the instantaneous phase.
-In~\cite{Sundar2009}, the authors compute a signal measuring phase correlation between consecutive frames, use a bandpass filter centered at 1Hz to recover the cardiac signal, and use a low pass filter to recover the respiratory signal.
+In~\cite{Karadayi2006}, the authors compute a signal from the x- or y-coordinate of the center of mass of each frame, use bandpass filtering to remove components outside typical cardiac frequency range, find the dominant frequency and apply matched sine/cosine filters to estimate the instantaneous phase.
+In~\cite{Sundar2009}, the authors compute a signal from the phase correlation between consecutive frames, use a bandpass filter centered at 1Hz to recover the embedded cardiac signal, and use a low pass filter to recover the respiratory signal.
 %
@@
 %
-In this section, we present the theory underlying the proposed methods.
-Specifically, we begin by describing our method for the estimation of instantaneous cardiac and respiratory phases in Section \ref{sec:method:phase_estimation}.
-We then present a robust method that uses these phase estimates to gate out video frames with significant respiratory motion in Section \ref{sec:method:gating}.
-Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression model for reconstructing images at any cardiac phase and facilitate temporal super-resolution.
+In this section, we present the methods and their underlying theoretical basis.
+We begin by describing our method for the estimation of instantaneous cardiac and respiratory phases in Section \ref{sec:method:phase_estimation}.
+We then present a robust method that uses these phase estimates to exclude video frames with significant respiratory motion in Section \ref{sec:method:gating}.
+In Section \ref{sec:method:super_resolution}, we present a kernel regression model for reconstructing images at any cardiac phase and facilitate temporal super-resolution.
 %
@@
 %
-\caption{Illustration of our phase estimation method: (a) Inter-frame similarity matrix, (b) Trend matrix corresponding to respiratory motion, (c) Residual matrix corresponding to beating heart motion, (d) Similarity profile chosen for phase estimation with associated trend/respiration and residual/heart-beat components, (e) Periodogram of heart-beat and respiration components along with periodicity characteristics (e.g.
-frequency, cycle duration) calculated from the dominant frequency, and (f) Instantaneous cardiac and respiratory phases estimated using the Hilbert transform.}
+\caption{Illustration of our phase estimation method: (a) Inter-frame similarity matrix, (b) Trend matrix corresponding to respiratory motion, (c) Residual matrix corresponding to beating heart motion, (d) Similarity profile chosen for phase estimation with associated trend/respiration and residual/heart-beat components, (e) Periodogram of heart-beat and respiration components along with periodicity characteristics (e.g., frequency, cycle duration) calculated from the dominant frequency, and (f) Instantaneous cardiac and respiratory phases estimated from the trend and residual signals using the Hilbert transform.}
 \label{fig:phase_estimation}
@@
 While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting such as the case of periodic cardiac ultrasound videos wherein several thousands of variables (pixel intensities) are involved.
-Our strategy was to find a way to transforming this complex multi-variate problem into a univariate one and take advantage of existing univariate methods to solve the problem.
+Our strategy was to find a way to transforming this complex multi-variate problem into a univariate one and to take advantage of existing univariate methods to solve the problem.
 
-	We first compute the similarity between all pairs of images/frames in the given periodic image sequence containing $N$ frames to create a symmetric matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames.
-Here in, we use normalized correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity but, in principle, any of the similarity metrics used in image registration algorithms can potentially be chosen.
-Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series.
-If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted, this time series can be expected to preserve the periodicity characteristics of the original image sequence.
+	We first compute the similarity between all pairs of images/frames in the given cardiac image sequence containing $N$ frames to create a symmetric matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames.
+Here in, we use the Peason normalized correlation coefficient to quantify inter-frame similarity; but, in practice, other image similarity metrics could have been used.
+Each row in the inter-frame similarity matrix $A$ can be considered as a univariate time series.
+For a given similarity metric, if the row's corresponding image frame is not significantly corrupted, the time-series similarity signal along a row will preserve the periodicity characteristics of the original image sequence.
 Two kinds of periodic motion are present in cardiac ultrasound videos, one corresponding to the beating heart motion and the other corresponding to respiratory motion.
@@
 
-	Next, we use a trend extraction technique~\cite{Alexandrov2012} called the Hodrick-Prescott (HP) filter~\cite{Yamada2015}, widely used in econometrics, to decompose the frame similarity signal $u(t)$ corresponding to each row of $A$ into a sum of two components: (i) Trend component $\tau_{resp}(t)$ encompassing the periodicity characteristic of only the lower-frequency respiratory motion, and (ii) Residual component $r_{heart}(t)$ encompassing the periodicity characteristic of only the higher-frequency beating heart motion.
+	Next, we use a trend extraction technique~\cite{Alexandrov2012} called the Hodrick-Prescott (HP) filter~\cite{Yamada2015}, which is widely used in econometrics, to decompose the frame similarity signal $u(t)$ corresponding to each row of $A$ into a sum of two components: (i) Trend component $\tau_{resp}(t)$ encompassing the periodicity characteristic of only the lower-frequency respiratory motion, and (ii) Residual component $r_{heart}(t)$ encompassing the periodicity characteristic of only the higher-frequency beating heart motion.
 The HP filter performs the decomposition of $u(t) = \tau_{resp}(t) + r_{heart}(t)$ by solving the following optimization problem:
@@
 Figures~\ref{fig:phase_estimation}(b,c) show the trend $A_{resp}$ and residual $A_{heart}$ matrices for one of our cardiac ultrasound videos.
-To minimize the effect of any noise on phase estimation, we pick frame similarity signal corresponding to the row of $A_{heart}$ whose periodogram or power-frequency distribution has minimum entropy.
+To select a single similarity signal that minimizes the effect of noise on phase estimation, we select the frame similarity signal that corresponds to the row of $A_{heart}$ whose periodogram or power-frequency distribution has minimum entropy.
 Let $\hat{u}(t)$ be the chosen frame similarity signal and let $\hat{\tau}_{resp}(t)$ and $\hat{r}_{heart}(t)$ be its associated trend/respiration and residual/heart-beat component signals, respectively.
@@
 Figure~\ref{fig:phase_estimation}(e) shows the periodogram of these two components that are nicely segregated in frequency domain.
-The periodicity characteristics (e.g.
-frequency in beats-per-minute (bpm), cycle duration in frames, number of cycles) of heart-beat and respiration can be calculated based on the maximum-power frequency of their periodograms.
+The periodicity characteristics (e.g., frequency in BPM, cycle duration in frames, number of cycles) of heart-beat and respiration can be calculated based on the maximum-power frequency of their periodograms.
 
@@
 
-	In the second step, we learn a robust mapping $L(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the residual/heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called Locally weighted regression (LOWESS) to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signal values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above.
+	In the second step, we learn a robust mapping $L(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the residual/heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called locally weighted regression (LOWESS) to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signal values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above.
 Next, we compute a robust estimate of the standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit based on the median absolute value of the difference between the LOWESS fit $L( \hat{\phi}_{heart}(t) )$ and residual/heart-beat signal $\hat{r}_{heart}(t)$ for all frames.
 Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert L( \hat{\phi}_{heart}(t) ) - \hat{r}_{heart}(t)  \rvert   > k \times \hat{\sigma}_{L}  \right \}$ for which the absolute  difference between the LOWESS fit and heart-beat signal value is greater than $k = 2.0$ (95\% confidence interval) times the standard deviation $\hat{\sigma}_{L}$.
-%
-\begin{IonFigT}
-\centering
-%
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/phaseordered.png}
-\ionbox{1.2in}\ionbox{1.1in}\ionbox{1.1in}\ionbox{1.1in}\\
-%
-\caption{Visualization of m-mode frames ordered by cardiac phase: (a) One of the frames in the video overlaid with the M-mode line shown in the next three images, (b) M-mode frames in the order they appear in the input video, (c) M-mode frames ordered by cardiac phase, and (d) Non-respiratory m-mode frames ordered by cardiac phase.}
-\label{fig:phase_ordering}
-\end{IonFigT}
 %
@@
 A single-cycle video representative of the subject's heart-beat can now be reconstructed by sampling the manifold $M(\phi)$ at any desired resolution.
+%
+%
+\begin{IonFigT}
+\centering
+%
+\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/phaseordered.png}
+\ionbox{1.2in}\ionbox{1.1in}\ionbox{1.1in}\ionbox{1.1in}\\
+%
+\caption{Visualization of m-mode frames ordered by cardiac phase: (a) One of the frames in the video overlaid with the M-mode line shown in the next three images, (b) M-mode frames in the order they appear in the input video, (c) M-mode frames ordered by cardiac phase, and (d) Non-respiratory m-mode frames ordered by cardiac phase.}
+\label{fig:phase_ordering}
+\end{IonFigT}
+%
+%
 \vspace{-0.5cm}
-\section{Results}
+\section{Experiments and Results}
 \label{sec:results}
@@
 We use cardiac ultrasound videos and associated ECG recordings of 6 mice to validate our methods.
+One scan was used during parameter selection for the system, and five scans were reserved for this testing.
 The ultrasound videos were acquired using the VisualSonics Vevo 2100 scanner at 233 frames per second.
@@
 %
-In this paper, we have presented a novel method to estimate the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video.
+In this paper, we have presented a novel method to estimate the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video, thereby eliminating the need for ECG and respiratory monitoring in, for example, small animal studies.
 We have also presented a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
-Having demonstrated a proof of concept, we next plan to evaluate our methods on more datasets, especially of subjects with irregular heart beats, and address any remaining pitfalls.
-One drawback of our phase estimation method is a strong assumption of periodicity which may breakdown in the case of subjects with cardiac arrhythmia.
-To address this, we will look into generalizations of the hilbert transform\cite{Lu2013} that are more susceptible noise.
+Having demonstrated a proof of concept, we next plan to evaluate our methods on more datasets, especially of subjects with irregular heart beats.
+One drawback of our phase estimation method is a strong assumption of periodicity which may be used to flag infrequent arrhythmia or that may breakdown in the case of persistent cardiac arrhythmia.
+To address persisten arrhythmia, we will look into generalizations of the Hilbert transform\cite{Lu2013} that are more susceptible noise.
 Another drawback is that the use normalized correlation to measure inter-frame similarity relies on an inherent assumption that a large part of the image is pulsating.

### 2017-02-23 paper155.tex
Reduce text by making sentences less verbose

--- 2017-02-23 paper155.tex
@@
 We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression to reconstruct images at any cardiac phase to facilitate temporal super-resolution.
-Previous work on the estimation of cardiac and/or respiratory phases directly from ultrasound videos is very limited.
-In~\cite{Karadayi2006}, the authors compute a signal measuring the x- or y-coordinate of the center of mass of each frame, use bandpass filtering to remove components outside typical cardiac frequency range, find the dominant frequency and apply matched sine/cosine filters to estimate the instantaneous phase.
-In~\cite{Sundar2009}, the authors compute a signal measuring phase correlation between consecutive frames, use a bandpass filter to recover the cardiac signal and a low pass filter to recover the respiratory signal.
+Previous work on estimation of cardiac and/or respiratory phases directly from ultrasound videos is very limited.
+In~\cite{Karadayi2006}, the authors compute a signal measuring the x- or y-coordinate of the center of mass of each frame, use bandpass filtering to remove components outside the cardiac frequency range, find the dominant frequency and apply matched sine/cosine filters to estimate the instantaneous phase.
+In~\cite{Sundar2009}, the authors compute a signal of phase correlation between consecutive frames and use bandpass and lowpass filtering to recover the cardiac and respiratory signals, respectively.
 %
@@
 We then present a robust method to gate out video frames with significant respiratory motion in Section \ref{sec:method:gating}.
-Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression model for reconstructing images at any cardiac phase and facilitate temporal super-resolution.
+Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
 %
@@
 %
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/ecg_instaphase_overlay.png}\\
-\ionbox{4.5in}\\
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/qrs_peak_to_peak.png}\\
-\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instaphase_valley_to_valley.png}
-\ionbox{4.5in}\\
+\includegraphics[width=4.25in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/ecg_instaphase_overlay.png}\\
+\ionbox{4.25in}\\
+\includegraphics[width=4.25in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/qrs_peak_to_peak.png}\\
+\includegraphics[width=4.25in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instaphase_valley_to_valley.png}
+\ionbox{4.25in}\\
 %
@@
 %
-While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data~\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting such as the case of periodic cardiac ultrasound videos wherein several thousands of variables (pixel intensities) are involved.
+While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data~\cite{Boashash1992}, there are not many methods that tackle this problem in a multivariate setting such as the case of cardiac ultrasound videos wherein thousands of variables (pixel intensities) are involved.
 Our strategy is to transform this complex multi-variate problem into a univariate one and take advantage of existing methods to solve the problem.
@@
 If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted, this time series can be expected to preserve the periodicity characteristics of the original image sequence.
-Two kinds of periodic motion are present in cardiac ultrasound videos, one corresponding to the beating heart motion and the other corresponding to respiratory motion.
-Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos wherein these two sources of periodicity can be clearly observed in many rows/columns.
+Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos wherein the periodicity of low-frequency respiratory motion and high-frequency beating heart motion can observed in many rows/columns.
 
-	Next, we use a trend extraction technique~\cite{Alexandrov2012} called the Hodrick-Prescott (HP) filter~\cite{Yamada2015}, widely used in econometrics, to decompose the frame similarity signal $u(t)$ corresponding to each row of $A$ into a sum of two components: (i) Trend component $\tau_{resp}(t)$ encompassing the periodicity characteristic of only the lower-frequency respiratory motion, and (ii) Residual component $r_{heart}(t)$ encompassing the periodicity characteristic of only the higher-frequency beating heart motion.
+	Next, we use a trend extraction technique~\cite{Alexandrov2012} called the Hodrick-Prescott (HP) filter~\cite{Yamada2015} to decompose the frame similarity signal $u(t)$ corresponding to each row of $A$ into a sum of two components: (i) Trend component $\tau_{resp}(t)$ with periodicity characteristic of only the lower-frequency respiratory motion, and (ii) Residual component $r_{heart}(t)$ with periodicity characteristic of only the higher-frequency beating heart motion.
 The HP filter performs the decomposition of $u(t) = \tau_{resp}(t) + r_{heart}(t)$ by solving the following optimization problem:
@@
 where $\nabla^2\tau_{resp}(t) = \tau_{resp}(t+1) - 2 \tau_{resp}(t) + \tau_{resp}(t-1)$ is the second-order difference or derivative of the trend signal and $\lambda=6400$ is penalty parameter.
-Let $A_{resp}$ and $A_{heart}$ be the matrices whose rows contain the trend/respiratory and residual/heart-beat components, respectively, of the frame similarity signals in the corresponding rows of the similarity matrix $A$.
-Figures~\ref{fig:phase_estimation}(b,c) show the trend $A_{resp}$ and residual $A_{heart}$ matrices for one of our cardiac ultrasound videos.
+Let $A_{resp}$ and $A_{heart}$ be the matrices whose rows contain the trend/respiratory and residual/heart-beat components, respectively (Figures~\ref{fig:phase_estimation}(b,c)).
 To minimize the effect of any noise on phase estimation, we pick frame similarity signal corresponding to the row of $A_{heart}$ whose periodogram or power-frequency distribution has minimum entropy.
-Let $\hat{u}(t)$ be the chosen frame similarity signal and let $\hat{\tau}_{resp}(t)$ and $\hat{r}_{heart}(t)$  be its associated trend/respiration and residual/heart-beat component signals, respectively.
-Figure~\ref{fig:phase_estimation}(d) shows the selected frame similarity signal (blue) along with the associated trend/respiration (pink) and residual/heart-beat (green) components.
-Figure~\ref{fig:phase_estimation}(e) shows the periodogram of these two components that are nicely segregated in frequency domain.
-The periodicity characteristics (e.g.
-frequency in beats-per-minute (bpm), cycle duration in frames, number of cycles) of heart-beat and respiration can be calculated based on the maximum-power frequency of their periodograms.
+Let $\hat{u}(t)$ be the chosen frame similarity signal and let $\hat{\tau}_{resp}(t)$ and $\hat{r}_{heart}(t)$  be its associated trend/respiration and residual/heart-beat component signals, respectively (Figure~\ref{fig:phase_estimation}(d)).
 
-	Next, considering the narrow-band nature of the respiration and heart-beat signals (see  Figure~\ref{fig:phase_estimation}(e)), we use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous phase of each frame.
+	Next, considering the narrow-band nature of the respiration and heart-beat signals (see  Figure~\ref{fig:phase_estimation}(e)), we use the Hilbert transform~\cite{Boashash1992} to estimate the instantaneous phase of each frame.
 Specifically, we compute the instantaneous intra-period phase $\phi(t) \in \left [  -\pi, \pi\right )$ of a periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows: $\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)$ and map $\phi(t)$ from the range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
-Let $\hat{\phi}_{heart}(t)$ and $\hat{\phi}_{resp}(t)$ denote the instantaneous cardiac and respiratory phases computed from the trend/respiration $\hat{\tau}_{resp}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signals, respectively.
-Figure~\ref{fig:phase_estimation}(f) shows the instantaneous cardiac and respiratory phases computed from the heart-beat and respiration signals in Figure~\ref{fig:phase_estimation}(d).
+Let $\hat{\phi}_{heart}(t)$ and $\hat{\phi}_{resp}(t)$ denote the instantaneous cardiac and respiratory phases (Figure~\ref{fig:phase_estimation}(f)) computed from the respiration $\hat{\tau}_{resp}(t)$ and heart-beat $\hat{r}_{heart}(t)$ signals, respectively.
 %
-\vspace{-0.5cm}
+\vspace{-0.3cm}
 \subsection{Gating out respiratory frames}
@@
 %
-\caption{Illustration of our method for gating out respiratory frames: (a) Frames $F_{cutoff}$ discarded (blue circles) in the first step overlaid with the trend/respiration, residual/heart-beat, and instantaneous respiratory phase signals (b) residual/heart-beat signal $\hat{r}_{heart}(t)$ vs cardiac phase $\hat{\phi}_{heart}(t)$ overlaid with frames $F_{cutoff}$ discarded in step-1 (blue circle), LOWESS fit $L(\phi)$ (black solid), upper and lower bounds or 95\% confidence interval $(L(\phi) \pm 2.0 * \hat{\sigma}_L)$ of valid non-respiratory frames (black dotted), and frames $F_{resp}$ (red circles) gated at the end of step-2.}
+\caption{Illustration of our method for gating out respiratory frames: (a) Frames $F_{cutoff}$ discarded (blue circles) in the first step overlaid with the respiration, heart-beat, and instantaneous respiratory phase signals (b) residual/heart-beat signal $\hat{r}_{heart}(t)$ vs cardiac phase $\hat{\phi}_{heart}(t)$ overlaid with frames $F_{cutoff}$ discarded in step-1 (blue circle), LOWESS fit $L(\phi)$ (black solid), upper and lower bounds or 95\% confidence interval $(L(\phi) \pm 2.0 * \hat{\sigma}_L)$ of valid non-respiratory frames (black dotted), and frames $F_{resp}$ (red circles) gated at the end of step-2.}
 \label{fig:respiratory_gating}
@@
 
-	In the first step, based on the observation that respiratory motion mainly occurs around the minima of the trend/respiration signal $\hat{\tau}_{resp}(t)$ (pink curve in Fig.~\ref{fig:phase_estimation}(d)) with respiratory phase $\hat{\phi}_{resp}(t) = 0$, we perform a rough initial gating by discarding the frames $F_{cutoff} = \left \{ t \mid \hat{\phi}_{resp}(t) < c \parallel \hat{\phi}_{resp}(t) > (1 - c) \right \}$ whose phase distance from $\hat{\phi}_{resp}(t) = 0$ is below a specified cutoff value $c=0.2$.
-Figure~\ref{fig:respiratory_gating}(a) shows the discarded frames $F_{cutoff}$ overlaid on the trend/respiration $\hat{\tau}_{resp}(t)$, residual/heart-beat  $\hat{r}_{heart}(t)$, and respiratory phase $\hat{\phi}_{resp}(t)$ signals.
+	In the first step, based on the observation that respiratory motion mainly occurs around the minima of the respiration signal $\hat{\tau}_{resp}(t)$ (pink curve in Fig.~\ref{fig:phase_estimation}(d)) with respiratory phase $\hat{\phi}_{resp}(t) = 0$, we perform a rough initial gating by discarding the frames $F_{cutoff} = \left \{ t \mid \hat{\phi}_{resp}(t) < c \parallel \hat{\phi}_{resp}(t) > (1 - c) \right \}$ whose phase distance from $\hat{\phi}_{resp}(t) = 0$ is below a specified cutoff value $c=0.2$.
+Figure~\ref{fig:respiratory_gating}(a) shows the discarded frames $F_{cutoff}$ overlaid on the respiration $\hat{\tau}_{resp}(t)$, heart-beat  $\hat{r}_{heart}(t)$, and respiratory phase $\hat{\phi}_{resp}(t)$ signals.
 
-	In the second step, we learn a robust mapping $L(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the residual/heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called Locally weighted regression (LOWESS) to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signal values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above.
-Next, we compute a robust estimate of the standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit based on the median absolute value of the difference between the LOWESS fit $L( \hat{\phi}_{heart}(t) )$ and residual/heart-beat signal $\hat{r}_{heart}(t)$ for all frames.
-Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert \hat{r}_{heart}(t) - L( \hat{\phi}_{heart}(t) ) \rvert   > k \times \hat{\sigma}_{L}  \right \}$ for which the absolute  difference between the LOWESS fit and heart-beat signal value is greater than $k = 2.0$ (95\% confidence interval) times the standard deviation $\hat{\sigma}_{L}$.
-%
-\vspace{-0.5cm}
-\subsection{Model to generate images by cardiac phase for super-resolution}
-\label{sec:method:super_resolution}
-%
-In this section, we present a kernel regression model to reconstruct the image at any cardiac phase.
-This model can then be used to generate a single-cycle video representative of the subject's heart-beat at a higher temporal resolution.
-
-Given a cardiac ultrasound video $I(t) : \{1, ..., N\} \to R^m$ of $N$ images with $m$ pixels each, we extract the residual/heart-beat component $\hat{r}_{heart}(t)$ of the frame similarity signal, estimate the instantaneous cardiac phase $\hat{\phi}_{heart}(t)$ of each frame, and compute the robust LOWESS fit $L(\phi) : \left [  0, 1\right ) \to R$ that maps cardiac phase to the heart-beat signal as  described in Sections~\ref{sec:method:phase_estimation} and ~\ref{sec:method:gating}.
-We then use Nadarya-Watson kernel regression\cite{Bishop2006} to learn a phase-parametrized image manifold in the form of a function $M(\phi): [0, 1) \to R^m $ that maps any cardiac phase $\phi$ to an image using kernel-weighted local average as follows:
-\begin{equation}
-M(\phi) = \frac{\sum_{t = 1}^{N} K \left( \phi, \hat{\phi}_{heart}(t) \right) I(t)}{\sum_{t = 1}^{n} K \left( \phi, \hat{\phi}_{heart}(t) \right)} 
-\end{equation}
-where kernel $K\left( \phi, \hat{\phi}_{heart}(t) \right) = exp\left \{ -\frac{ \mid \phi - \hat{\phi}_{heart}(t) \mid^2}{2  \sigma^2_\phi} \right \} \times exp\left \{ -\frac{ \mid L(\phi) - \hat{r}_{heart}(t) \mid^2}{2  \sigma^2_{L}} \right \}$ is defined as the product of two radial-basis function (RBF) kernels.
-The first RBF kernel  weighs images inversely proportional to their distance (accounting for periodicity) in the cardiac phase space.
-Its bandwidth $\sigma_\phi$ is set equal to a constant $k_\phi = 0.4$ times the median difference in cardiac phase between consecutive frames of the given image sequence.
-The second RBF kernel weighs images inversely proportional to their deviation from the LOWESS prediction $L(\phi)$ in the residual/heart-beat signal space.
-Its bandwidth $\sigma_{L}$ is set equal to a constant $k_L = 2.0$ times the robust estimate of standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit as described in Section~\ref{sec:method:gating}.
-A single-cycle video representative of the subject's heart-beat can now be reconstructed by sampling the manifold $M(\phi)$ at any desired resolution.
+	In the second step, we learn a robust mapping $L(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called Locally weighted regression (LOWESS)~\cite{Cleveland1988} to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and heart-beat signal $\hat{r}_{heart}(t)$  values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above.
+Next, we compute a robust estimate of the standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit based on the median absolute value of the difference between the LOWESS fit $L( \hat{\phi}_{heart}(t) )$ and heart-beat signal $\hat{r}_{heart}(t)$ for all frames.
+Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert \hat{r}_{heart}(t) - L( \hat{\phi}_{heart}(t) ) \rvert   > k \times \hat{\sigma}_{L}  \right \}$
+whose heart-beat signal value deviates from the LOWESS fit is greater than $k = 2.0$ (95\% confidence interval) times the standard deviation $\hat{\sigma}_{L}$.	
 %
@@
 %
+\vspace{-0.3cm}
+\subsection{Model to generate images by cardiac phase for super-resolution}
+\label{sec:method:super_resolution}
 %
-\vspace{-0.5cm}
+In this section, we present a kernel regression model to reconstruct the image at any cardiac phase.
+This model can then be used to generate a single-cycle video representative of the subject's heart-beat at a higher temporal resolution.
+
+Given a cardiac ultrasound video $I(t) : \{1, ..., N\} \to R^m$ of $N$ images with $m$ pixels each, we extract the heart-beat signal $\hat{r}_{heart}(t)$, estimate the instantaneous cardiac phase $\hat{\phi}_{heart}(t)$ of each frame, and compute the robust LOWESS fit $L(\phi) : \left [  0, 1\right ) \to R$ that maps cardiac phase to the heart-beat signal as described in Sections~\ref{sec:method:phase_estimation} and ~\ref{sec:method:gating}.
+We then use Nadarya-Watson kernel regression\cite{Bishop2006} to learn a function $M(\phi): [0, 1) \to R^m $ that reconstructs the image for any cardiac phase $\phi$ using a kernel-weighted local average as follows:
+\begin{equation}
+M(\phi) = \frac{\sum_{t = 1}^{N} K \left( \phi, \hat{\phi}_{heart}(t) \right) I(t)}{\sum_{t = 1}^{n} K \left( \phi, \hat{\phi}_{heart}(t) \right)} 
+\end{equation}
+where kernel $K\left( \phi, \hat{\phi}_{heart}(t) \right) = exp\left \{ -\frac{ \mid \phi - \hat{\phi}_{heart}(t) \mid^2}{2  \sigma^2_\phi} \right \} \times exp\left \{ -\frac{ \mid L(\phi) - \hat{r}_{heart}(t) \mid^2}{2  \sigma^2_{L}} \right \}$ is defined as the product of two radial-basis function (RBF) kernels.
+The first RBF kernel  weighs images inversely proportional to their distance (accounting for periodicity) in the cardiac phase space.
+Its bandwidth $\sigma_\phi$ is set equal to a constant $k_\phi = 0.4$ times the median difference in cardiac phase between consecutive frames of the given image sequence.
+The second RBF kernel weighs images inversely proportional to the deviation of their residual/heart-beat signal value from the LOWESS prediction $L(\phi)$.
+Its bandwidth $\sigma_{L}$ is set equal to a constant $k_L = 2.0$ times the robust estimate of standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames ($t \notin F_{resp}$) around the LOWESS fit.
+The function $M(\phi)$ can now be used to reconstruct a single-cycle video representative of the subject's heart-beat at any desired resolution/sampling of phase in the range $[0, 1)$.
+%
+\vspace{-0.3cm}
 \section{Results}
@@
 	We validate the cardiac phase estimates of our method by comparing them with the ECG signal which is the gold standard for cardiac gating.
-Figure~\ref{fig:instaphase_vs_ecg}(a) shows a head-to-head comparison between the ECG signal and the cardiac phase signal obtained using the method described in Section~\ref{sec:method:phase_estimation} for one of the 6 videos.
+Figure~\ref{fig:instaphase_vs_ecg}(a) shows a head-to-head comparison between the ECG signal and the estimated cardiac phase signal for one of the 6 videos.
 Notice how well the locations of the QRS peaks (red circles) in the ECG signal and the corresponding minima (black circles) of the cardiac phase signal in each cycle match.
@@
 \centering
-\caption{Comparision between frame positions of the QRS peaks in ECG signal and the corresponding minima of the estimated cardiac phase for 6 videos.}
+\caption{Match between frame positions of QRS peaks in ECG signal and corresponding cardiac phase minima.}
 \begin{tabular}{|c|c|c|}
@@
 \begin{minipage}[t]{0.45\linewidth}
-\caption{Evaluation of our kernel regression model for reconstructing images at any cardiac phase using 50 rounds of Leave one out cross-validation (LOOCV) for 6 videos.}
 \centering
+\caption{Evaluation of our kernel regression model Leave one out cross-validation (LOOCV) of 50 rounds.}
 \begin{tabular}{|c|c|c|}
@@
 We have also presented a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
-Having demonstrated a proof of concept, we next plan to evaluate our methods on more datasets, especially of subjects with irregular heart beats, and address any remaining pitfalls.
+Having demonstrated a proof of concept, we next plan to evaluate our methods on more datasets and address any remaining pitfalls.
 One drawback of our phase estimation method is a strong assumption of periodicity which may breakdown in the case of subjects with cardiac arrhythmia.
-To address this, we will look into generalizations of the hilbert transform\cite{Lu2013} that are more susceptible noise.
+To address this, we will look into univariate methods for phase estimation in quasi-periodic signals that are less susceptible noise.
 Another drawback is that the use normalized correlation to measure inter-frame similarity relies on an inherent assumption that a large part of the image is pulsating.
 To relax this, we will try to devise more sophisticated local patch-based similarity measures.
-We encourage the readers to watch the supplementary videos for a better understanding of the effectiveness and utility of the proposed methods.
 %

### 2017-02-23 paper155.tex
Integrated stephens edits in abstract

@@
 Ultrasound is emerging as an increasingly effective tool for rapid non-invasive assessment of cardiac structure and function.
-Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in many applications.
+Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in many applications and is particularly challenging in pre-clinical studies involving small animals with rapid heart and respiration rates.
 In this paper, we present a novel method for estimation of instantaneous cardiac and respiratory phases directly from cardiac ultrasound videos.
@@
 Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection~\cite{Wick2013}, 3D reconstruction~\cite{Wachinger2012}, and temporal super-resolution~\cite{Cherin2006}.
+These tasks are particularly common and challenging in pre-clinical studies involving small animals such as mice that have rapid heart (310-840 BPM) and respiration (80-230 BPM) rates and yet are difficult to instrument with electronic cardiac and respirator monitors.
 Typically, the position within cardiac cycle is tracked through ECG data and position within respiratory cycle is tracked by motion of markers placed on the subject's body\cite{Khamene2004}.

### 2017-02-23 paper155.tex
Incorporated stephen's edits and almost down to 8 pages

@@
 %\keywords{Cardiac, Ultrasound, Echocardiography, Phase estimation, Gating, Temporal Super-resolution}
-\vspace{-0.5cm}
+\vspace{-0.3cm}
 \end{abstract}
@@
 Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection~\cite{Wick2013}, 3D reconstruction~\cite{Wachinger2012}, and temporal super-resolution~\cite{Cherin2006}.
-These tasks are particularly common and challenging in pre-clinical studies involving small animals such as mice that have rapid heart (310-840 BPM) and respiration (80-230 BPM) rates and yet are difficult to instrument with electronic cardiac and respirator monitors.
+These tasks are particularly challenging in pre-clinical studies involving small animals such as mice that have rapid heart (310-840 BPM) and respiration (80-230 BPM) rates.
 Typically, the position within cardiac cycle is tracked through ECG data and position within respiratory cycle is tracked by motion of markers placed on the subject's body\cite{Khamene2004}.
@@
 	In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video.
-We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression to reconstruct images at any cardiac phase to facilitate temporal super-resolution.
+We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
+
 	Previous work on estimation of cardiac and/or respiratory phases directly from ultrasound videos is very limited.
-In~\cite{Karadayi2006}, the authors compute a signal measuring the x- or y-coordinate of the center of mass of each frame, use bandpass filtering to remove components outside the cardiac frequency range, find the dominant frequency and apply matched sine/cosine filters to estimate the instantaneous phase.
-In~\cite{Sundar2009}, the authors compute a signal of phase correlation between consecutive frames and use bandpass and lowpass filtering to recover the cardiac and respiratory signals, respectively.
+In~\cite{Karadayi2006}, the authors compute a signal of the x- or y-coordinate of the center-of-mass of each frame, use bandpass filtering to remove components outside the cardiac frequency range, and apply matched sine/cosine filters to estimate the instantaneous phase.
+In~\cite{Sundar2009}, the authors compute a signal of phase correlation between consecutive frames and use an approach similar to~\cite{Karadayi2006} to recover the cardiac and respiratory signals.
+The center-of-mass signals are not reliable and phase correlation finds global translation that does not model beating heart and out-of-plane motion well.
 %
-\vspace{-0.5cm}
+\vspace{-0.3cm}
 \section{Method}
 \label{sec:method}
+\vspace{-0.3cm}
 %
 In this section, we present the theory underlying the proposed methods.
-Specifically, we begin by describing our method for estimation of instantaneous cardiac and respiratory phases in Section \ref{sec:method:phase_estimation}.
-We then present a robust method to gate out video frames with significant respiratory motion in Section \ref{sec:method:gating}.
-Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
+In Section~\ref{sec:method:phase_estimation}, we begin by describing our method for estimation of instantaneous cardiac and respiratory phases.
+In Section~\ref{sec:method:gating}, we present a robust method to exclude video frames with significant respiratory motion.
+In Section~\ref{sec:method:super_resolution}, we present a kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
 %
@@
 \end{IonFigT}
-\vspace{-0.5cm}
+\vspace{-0.3cm}
 \subsection{Estimation of instantaneous cardiac and respiratory phases}
@@
 %
-\includegraphics[width=4.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat.png}
-\ionbox{1.33in}\ionbox{1.33in}\ionbox{1.33in}\\
+\includegraphics[width=3.9in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat.png}
+\ionbox{1.3in}\ionbox{1.3in}\ionbox{1.3in}\\
 \includegraphics[width=4.25in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/season_trend_decomposition.png}
@@
 %
-\caption{Illustration of our phase estimation method: (a) Inter-frame similarity matrix, (b) Trend matrix corresponding to respiratory motion, (c) Residual matrix corresponding to beating heart motion, (d) Similarity profile chosen for phase estimation with associated trend/respiration and residual/heart-beat components, (e) Periodogram of heart-beat and respiration components along with periodicity characteristics (e.g.
+\caption{Illustration of our phase estimation method: (a) Inter-frame similarity matrix, (b) Trend matrix corresponding to respiratory motion, (c) Residual matrix corresponding to beating heart motion, (d) Frame similarity $\hat{u}(t)$ selected for phase estimation with associated trend/respiration $\hat{\tau}_{resp}$ and residual/heart-beat $\hat{r}_{heart}$ signals, (e) Periodogram of heart-beat and respiration signals along with periodicity characteristics (e.g.
 frequency, cycle duration) calculated from the dominant frequency, and (f) Instantaneous cardiac and respiratory phases estimated using the Hilbert transform.}
@@
 	We first compute the similarity between all pairs of images/frames in the given periodic image sequence containing $N$ frames to create a symmetric matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames.
-Here in, we use normalized correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity but, in principle, any of the similarity metrics used in image registration algorithms can potentially be chosen.
+Here in, we use normalized correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity; but in principle, other image similarity metrics can be used.
 Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series.
-If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted, this time series can be expected to preserve the periodicity characteristics of the original image sequence.
-Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos wherein the periodicity of low-frequency respiratory motion and high-frequency beating heart motion can observed in many rows/columns.
+If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted, this time series will preserve the periodicity characteristics of the original image sequence.
+Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos wherein the periodicity of low-frequency respiratory motion and high-frequency beating heart motion can observed.
 
-	Next, we use a trend extraction technique~\cite{Alexandrov2012} called the Hodrick-Prescott (HP) filter~\cite{Yamada2015} to decompose the frame similarity signal $u(t)$ corresponding to each row of $A$ into a sum of two components: (i) Trend component $\tau_{resp}(t)$ with periodicity characteristic of only the lower-frequency respiratory motion, and (ii) Residual component $r_{heart}(t)$ with periodicity characteristic of only the higher-frequency beating heart motion.
-The HP filter performs the decomposition of $u(t) = \tau_{resp}(t) + r_{heart}(t)$ by solving the following optimization problem:
+	Next, we use a trend extraction technique called the Hodrick-Prescott (HP) filter~\cite{Alexandrov2012} to decompose the frame similarity signal $u^i(t)$ corresponding to each row $i$ of the matrix $A$ into a sum of two components: (i) Trend component $\tau^i_{resp}(t)$ with periodicity characteristic of only respiratory motion, and (ii) Residual component $r^i_{heart}(t)$ with periodicity characteristic of only beating heart motion.
+The HP filter performs the decomposition of $u^i(t) = \tau^i_{resp}(t) + r^i_{heart}(t)$ by solving the following optimization problem:
 \begin{equation}	
-\argmin{\tau_{resp}(t)} \left[ \sum_{t=1}^{N}  \left(u(t) - \tau_{resp}(t) \right)^2  + \lambda \sum_{t=1}^{N-1} \left( \nabla^2 \tau_{resp}(t) \right)^2 \right] 
+\argmin{\tau^i_{resp}(t)} \left[ \sum_{t=1}^{N}  \left(u^i(t) - \tau^i_{resp}(t) \right)^2  + \lambda \sum_{t=1}^{N-1} \left( \nabla^2 \tau^i_{resp}(t) \right)^2 \right] 
 \end{equation}
-where $\nabla^2\tau_{resp}(t) = \tau_{resp}(t+1) - 2 \tau_{resp}(t) + \tau_{resp}(t-1)$ is the second-order difference or derivative of the trend signal and $\lambda=6400$ is penalty parameter.
-Let $A_{resp}$ and $A_{heart}$ be the matrices whose rows contain the trend/respiratory and residual/heart-beat components, respectively (Figures~\ref{fig:phase_estimation}(b,c)).
-To minimize the effect of any noise on phase estimation, we pick frame similarity signal corresponding to the row of $A_{heart}$ whose periodogram or power-frequency distribution has minimum entropy.
-Let $\hat{u}(t)$ be the chosen frame similarity signal and let $\hat{\tau}_{resp}(t)$ and $\hat{r}_{heart}(t)$  be its associated trend/respiration and residual/heart-beat component signals, respectively (Figure~\ref{fig:phase_estimation}(d)).
+where $\nabla^2\tau^i_{resp}(t) = \tau^i_{resp}(t+1) - 2 \tau^i_{resp}(t) + \tau^i_{resp}(t-1)$ is the second-order difference or derivative of the trend signal and $\lambda=6400$ is penalty parameter.
+Let $A_{resp}$ and $A_{heart}$ be the matrices whose rows contain the trend/respiratory and residual/heart-beat components (Figures~\ref{fig:phase_estimation}(b,c)), respectively, of the frame similarity signal in the corresponding rows of matrix $A$.
+We now need to select the signals corresponding to one of the rows for phase estimation.
+To minimize the effect of any noise, we pick the row for which the periodogram or power-frequency distribution of the heart-beat component in $A_{resp}$ has minimum entropy.
+Let $\hat{u}(t)$ be the selected frame similarity signal and let $\hat{\tau}_{resp}(t)$ and $\hat{r}_{heart}(t)$  be its associated trend and residual components  (Figure~\ref{fig:phase_estimation}(d)) that we will henceforth refer to as respiration and heart-beat signals, respectively.
 
 	Next, considering the narrow-band nature of the respiration and heart-beat signals (see  Figure~\ref{fig:phase_estimation}(e)), we use the Hilbert transform~\cite{Boashash1992} to estimate the instantaneous phase of each frame.
-Specifically, we compute the instantaneous intra-period phase $\phi(t) \in \left [  -\pi, \pi\right )$ of a periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows: $\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)$ and map $\phi(t)$ from the range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
-Let $\hat{\phi}_{heart}(t)$ and $\hat{\phi}_{resp}(t)$ denote the instantaneous cardiac and respiratory phases (Figure~\ref{fig:phase_estimation}(f)) computed from the respiration $\hat{\tau}_{resp}(t)$ and heart-beat $\hat{r}_{heart}(t)$ signals, respectively.
+Specifically, we compute the instantaneous phase $\phi(t) \in \left [  -\pi, \pi\right )$ of a periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows: $\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)$ and map $\phi(t)$ from $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
+Let $\hat{\phi}_{heart}(t)$ and $\hat{\phi}_{resp}(t)$ denote the instantaneous cardiac and respiratory phases (Figure~\ref{fig:phase_estimation}(f)) computed from the respiration and heart-beat signals, respectively.
 %
@@
 %
-\caption{Illustration of our method for gating out respiratory frames: (a) Frames $F_{cutoff}$ discarded (blue circles) in the first step overlaid with the respiration, heart-beat, and instantaneous respiratory phase signals (b) residual/heart-beat signal $\hat{r}_{heart}(t)$ vs cardiac phase $\hat{\phi}_{heart}(t)$ overlaid with frames $F_{cutoff}$ discarded in step-1 (blue circle), LOWESS fit $L(\phi)$ (black solid), upper and lower bounds or 95\% confidence interval $(L(\phi) \pm 2.0 * \hat{\sigma}_L)$ of valid non-respiratory frames (black dotted), and frames $F_{resp}$ (red circles) gated at the end of step-2.}
+\caption{Illustration of the respiratory gating method: (a) Frames $F_{cutoff}$ discarded (blue circles) in step-1 overlaid with the respiration, heart-beat, and respiratory phase signals (b) heart-beat signal $\hat{r}_{heart}(t)$ vs cardiac phase $\hat{\phi}_{heart}(t)$ overlaid with frames $F_{cutoff}$ discarded in step-1 (blue circle), LOWESS fit $L(\phi)$ (black solid), upper and lower bounds or 95\% confidence interval $(L(\phi) \pm 2.0 * \hat{\sigma}_L)$ of non-respiratory frames (black dotted), and frames $F_{resp}$ (red circles) gated after step-2.}
 \label{fig:respiratory_gating}
@@
 %
-Once the cardiac and respiratory phases of each frame have been estimated, they can be used to filter out or guide the extraction of quantitative measurements from a desired part/point of the periodic cycle, a process commonly referred to as gating.
-In this section, we present a robust two-step method that uses these phase estimates to gate out video frames with significant respiratory motion.
+Once the cardiac and respiratory phases of each frame have been estimated, they can be used to select or filter out frames from a desired part/point in the periodic cycle, a process commonly referred to as gating.
+In this section, we present a robust two-step method that uses these phase estimates to filter out video frames with significant respiratory motion.
 
@@
 Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert \hat{r}_{heart}(t) - L( \hat{\phi}_{heart}(t) ) \rvert   > k \times \hat{\sigma}_{L}  \right \}$
-whose heart-beat signal value deviates from the LOWESS fit is greater than $k = 2.0$ (95\% confidence interval) times the standard deviation $\hat{\sigma}_{L}$.	
+whose heart-beat signal value deviates from the LOWESS by more than $k = 2.0$ (95\% confidence interval) times the standard deviation $\hat{\sigma}_{L}$ (Figure~\ref{fig:respiratory_gating}(b)).	
 %
@@
 \vspace{-0.3cm}
-\section{Results}
+\section{Experiments and Results}
 \label{sec:results}
+\vspace{-0.3cm}
 %
@@
 Figure~\ref{fig:instaphase_vs_ecg}(b) presents a visual comparison of five video frames evenly spaced in time between two consecutive QRS peaks (top-row) and corresponding minima of the cardiac phase signal (bottom-row).
-Notice the high-level of similarity in the images of each column.
-Figure~\ref{fig:phase_ordering}(b,c) shows the frames of an m-mode line before and after ordering with the estimated cardiac phase.
+Figure~\ref{fig:phase_ordering}(b,c) shows frames of an m-mode line before and after ordering with the estimated cardiac phase.
 Figure~\ref{fig:phase_ordering}(d) shows m-mode frames ordered by cardiac phase after gating out respiratory frames.
@@
 %
-\vspace{-0.5cm}
+\vspace{-0.3cm}
 \section{Conclusion}
 \label{sec:conclusion}
+\vspace{-0.3cm}
 %
-In this paper, we have presented a novel method to estimate the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video.
+In this paper, we have presented a novel method to estimate the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video, thereby eliminating the need of additional hardware to monitor them in, for example, small animal studies.
 We have also presented a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
-Having demonstrated a proof of concept, we next plan to evaluate our methods on more datasets and address any remaining pitfalls.
+We next plan to evaluate our methods on more datasets and address any remaining pitfalls.
 One drawback of our phase estimation method is a strong assumption of periodicity which may breakdown in the case of subjects with cardiac arrhythmia.
@@
 Another drawback is that the use normalized correlation to measure inter-frame similarity relies on an inherent assumption that a large part of the image is pulsating.
-To relax this, we will try to devise more sophisticated local patch-based similarity measures.
+To relax this, we will try to devise local patch-based similarity measures.
 %
 \vspace{-0.5cm}
+%
 \bibliographystyle{splncs03}

### 2017-02-23 paper155.tex
First decent version under 8 pages

@@
 
-	Previous work on estimation of cardiac and/or respiratory phases directly from ultrasound videos is very limited.
+	Previous work on the estimation of cardiac and/or respiratory phases directly from ultrasound videos is very limited.
 In~\cite{Karadayi2006}, the authors compute a signal of the x- or y-coordinate of the center-of-mass of each frame, use bandpass filtering to remove components outside the cardiac frequency range, and apply matched sine/cosine filters to estimate the instantaneous phase.
 In~\cite{Sundar2009}, the authors compute a signal of phase correlation between consecutive frames and use an approach similar to~\cite{Karadayi2006} to recover the cardiac and respiratory signals.
-The center-of-mass signals are not reliable and phase correlation finds global translation that does not model beating heart and out-of-plane motion well.
+The center-of-mass signals are not reliable and phase correlation finds global translation that does not model beating and out-of-plane motion of the heart well.
 %
-\vspace{-0.3cm}
+\vspace{-0.9cm}
 \section{Method}
@@
 	We validate the cardiac phase estimates of our method by comparing them with the ECG signal which is the gold standard for cardiac gating.
-Figure~\ref{fig:instaphase_vs_ecg}(a) shows a head-to-head comparison between the ECG signal and the estimated cardiac phase signal for one of the 6 videos.
-Notice how well the locations of the QRS peaks (red circles) in the ECG signal and the corresponding minima (black circles) of the cardiac phase signal in each cycle match.
+Figure~\ref{fig:instaphase_vs_ecg}(a) shows a comparison between the ECG signal and the estimated cardiac phase signal for one of the 6 videos.
+Notice how well the locations of the QRS peaks in the ECG signal and the corresponding minima of the cardiac phase signal in each cycle match.
 Table 1 reports the $R^2$ correlation and the $mean \pm std$ error between the locations of the QRS peaks and the corresponding minima of the cardiac phase signals for all 6 videos.
 Figure~\ref{fig:instaphase_vs_ecg}(b) presents a visual comparison of five video frames evenly spaced in time between two consecutive QRS peaks (top-row) and corresponding minima of the cardiac phase signal (bottom-row).
-Figure~\ref{fig:phase_ordering}(b,c) shows frames of an m-mode line before and after ordering with the estimated cardiac phase.
-Figure~\ref{fig:phase_ordering}(d) shows m-mode frames ordered by cardiac phase after gating out respiratory frames.
-The supplementary material includes a video showing the estimated cardiac and respiratory phases of each frame for one of our datasets.
+Figure~\ref{fig:phase_ordering} shows frames of an m-mode line through the ventricle before and after ordering by cardiac phase.
+The supplementary material includes a video showing the cardiac and respiratory phases of each frame for one of our datasets.
 
@@
 \caption{Match between frame positions of QRS peaks in ECG signal and corresponding cardiac phase minima.}
+\vspace{-0.1cm}
 \begin{tabular}{|c|c|c|}
@@
 \caption{Evaluation of our kernel regression model Leave one out cross-validation (LOOCV) of 50 rounds.}
+\vspace{-0.1cm}
 \begin{tabular}{|c|c|c|}
@@
 \end{minipage}
-\vspace{-0.3cm}
+\vspace{-0.5cm}
 \end{table}

### 2017-02-23 paper155.tex
Minor corrections

@@
 To address this, we will look into univariate methods for phase estimation in quasi-periodic signals that are less susceptible noise.
-Another drawback is that the use normalized correlation to measure inter-frame similarity relies on an inherent assumption that a large part of the image is pulsating.
+Another drawback is the use of normalized correlation to measure inter-frame similarity relies on an inherent assumption that a large part of the image is pulsating.
 To relax this, we will try to devise local patch-based similarity measures.

### 2017-02-24 paper155.tex
Made a few corrections

@@
 	In the second step, we learn a robust mapping $L(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called Locally weighted regression (LOWESS)~\cite{Cleveland1988} to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and heart-beat signal $\hat{r}_{heart}(t)$  values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above.
-Next, we compute a robust estimate of the standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit based on the median absolute value of the difference between the LOWESS fit $L( \hat{\phi}_{heart}(t) )$ and heart-beat signal $\hat{r}_{heart}(t)$ for all frames.
+Next, we compute a robust estimate of the standard deviation $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit based on the median absolute deviation between the LOWESS fit $L( \hat{\phi}_{heart}(t) )$ and heart-beat signal $\hat{r}_{heart}(t)$ for all frames.
 Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert \hat{r}_{heart}(t) - L( \hat{\phi}_{heart}(t) ) \rvert   > k \times \hat{\sigma}_{L}  \right \}$
-whose heart-beat signal value deviates from the LOWESS by more than $k = 2.0$ (95\% confidence interval) times the standard deviation $\hat{\sigma}_{L}$ (Figure~\ref{fig:respiratory_gating}(b)).	
+whose heart-beat signal value deviates from the LOWESS fit by more than $k = 2.0$ (95\% confidence interval) times the standard deviation $\hat{\sigma}_{L}$ (Figure~\ref{fig:respiratory_gating}(b)).	
 %
@@
 The second RBF kernel weighs images inversely proportional to the deviation of their residual/heart-beat signal value from the LOWESS prediction $L(\phi)$.
-Its bandwidth $\sigma_{L}$ is set equal to a constant $k_L = 2.0$ times the robust estimate of standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames ($t \notin F_{resp}$) around the LOWESS fit.
+Its bandwidth $\sigma_{L}$ is set equal to a constant $k_L = 2.0$ times the robust estimate of standard deviation $\hat{\sigma}_{L}$ of non-respiratory frames ($t \notin F_{resp}$) around the LOWESS fit.
 The function $M(\phi)$ can now be used to reconstruct a single-cycle video representative of the subject's heart-beat at any desired resolution/sampling of phase in the range $[0, 1)$.

### 2017-02-24 paper155.tex
ENH: some minor changes (rkwitt)

@@
 \usepackage{wrapfig}
+\usepackage{hyperref}
+\usepackage{color}
+\usepackage[font=small,labelfont=bf]{caption}
+
+\newcommand{\rk}[1]{{\color{red}{#1}}}
+
 %
@@
 Cardiac ultrasound videos consist of two kinds of periodic motion, one corresponding to beating heart motion and the other corresponding to respiratory motion.
-Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating\cite{VonBirgelen1997}, quiescence detection~\cite{Wick2013}, 3D reconstruction~\cite{Wachinger2012}, and temporal super-resolution~\cite{Cherin2006}.
+Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating \cite{VonBirgelen1997}, quiescence detection~\cite{Wick2013}, 3D reconstruction~\cite{Wachinger2012}, and temporal super-resolution~\cite{Cherin2006}.
 These tasks are particularly challenging in pre-clinical studies involving small animals such as mice that have rapid heart (310-840 BPM) and respiration (80-230 BPM) rates.
-Typically, the position within cardiac cycle is tracked through ECG data and position within respiratory cycle is tracked by motion of markers placed on the subject's body\cite{Khamene2004}.
+Typically, the position within cardiac cycle is tracked through ECG data and position within \rk{the} respiratory cycle is tracked by motion of markers placed on the subject's body \cite{Khamene2004}.
 
-	In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video.
+\vskip0.5ex
+\noindent
+\rk{\textbf{Contribution.}} In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video.
 We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
 
-	Previous work on the estimation of cardiac and/or respiratory phases directly from ultrasound videos is very limited.
+\vskip0.5ex
+\noindent
+\rk{\textbf{Related work.}} Previous work estimation of cardiac and/or respiratory phases directly from ultrasound videos is very limited.
 In~\cite{Karadayi2006}, the authors compute a signal of the x- or y-coordinate of the center-of-mass of each frame, use bandpass filtering to remove components outside the cardiac frequency range, and apply matched sine/cosine filters to estimate the instantaneous phase.
 In~\cite{Sundar2009}, the authors compute a signal of phase correlation between consecutive frames and use an approach similar to~\cite{Karadayi2006} to recover the cardiac and respiratory signals.
-The center-of-mass signals are not reliable and phase correlation finds global translation that does not model beating and out-of-plane motion of the heart well.
-%
-\vspace{-0.9cm}
+The center-of-mass signals are not reliable and phase correlation finds global translation that does not model beating and out-of-plane \rk{heart motion} well.
+
 \section{Method}
@@
 In this section, we present the theory underlying the proposed methods.
-In Section~\ref{sec:method:phase_estimation}, we begin by describing our method for estimation of instantaneous cardiac and respiratory phases.
+In Section~\ref{sec:method:phase_estimation}, we \rk{describe} our method for estimation of instantaneous cardiac and respiratory phases.
 In Section~\ref{sec:method:gating}, we present a robust method to exclude video frames with significant respiratory motion.
@@
 
-	We first compute the similarity between all pairs of images/frames in the given periodic image sequence containing $N$ frames to create a symmetric matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames.
-Here in, we use normalized correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity; but in principle, other image similarity metrics can be used.
+	We first compute the similarity between all pairs of images/frames in the given periodic image sequence containing $N$ frames to create a symmetric matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ \rk{frame}.
+Here in, we use normalized correlation (also known as the Pearson correlation coefficient) to quantify inter-frame similarity; but in principle, other image similarity metrics can be used.
 Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series.
@@
 Given a cardiac ultrasound video $I(t) : \{1, ..., N\} \to R^m$ of $N$ images with $m$ pixels each, we extract the heart-beat signal $\hat{r}_{heart}(t)$, estimate the instantaneous cardiac phase $\hat{\phi}_{heart}(t)$ of each frame, and compute the robust LOWESS fit $L(\phi) : \left [  0, 1\right ) \to R$ that maps cardiac phase to the heart-beat signal as described in Sections~\ref{sec:method:phase_estimation} and ~\ref{sec:method:gating}.
-We then use Nadarya-Watson kernel regression\cite{Bishop2006} to learn a function $M(\phi): [0, 1) \to R^m $ that reconstructs the image for any cardiac phase $\phi$ using a kernel-weighted local average as follows:
+We then use Nadarya-Watson kernel regression \cite{Bishop2006} to learn a function $M(\phi): [0, 1) \to R^m $ that reconstructs the image for any cardiac phase $\phi$ using a kernel-weighted local average as follows:
 \begin{equation}
@@
 The second column of Table~\ref{table:image_reconstruction} shows the $mean \pm std$ of normalized correlation between the reconstructed and the original images over 50 rounds of LOCCV for all 6 videos.
-As a baseline, the third column of  Table~\ref{table:image_reconstruction} shows the $mean \pm std$ of the normalized correlation between frames at QRS peaks of the ECG signal.
+As a \emph{baseline}, the third column of  Table~\ref{table:image_reconstruction} shows the $mean \pm std$ of the normalized correlation between frames at QRS peaks of the ECG signal.
 The supplementary material includes single-cycle videos at 2x and 4x temporal resolution generated using the kernel regression model wherein the motion of heart chambers and valves is much clear than the original video.
@@
 \begin{table}[t]
+\begin{small}
 \begin{minipage}[t]{0.45\linewidth}
@@
 \caption{Match between frame positions of QRS peaks in ECG signal and corresponding cardiac phase minima.}
-\vspace{-0.1cm}
+%\vspace{-0.1cm}
 \begin{tabular}{|c|c|c|}
@@
 \centering
-\caption{Evaluation of our kernel regression model Leave one out cross-validation (LOOCV) of 50 rounds.}
-\vspace{-0.1cm}
+\caption{Evaluation of our kernel regression model. Leave one out cross-validation (LOOCV) of 50 rounds.}
+%\vspace{-0.1cm}
 \begin{tabular}{|c|c|c|}
@@
 \vspace{-0.5cm}
+\end{small}
 \end{table}
@@
 \label{sec:conclusion}
-\vspace{-0.3cm}
+\vspace{-0.2cm}
 %
@@
 We next plan to evaluate our methods on more datasets and address any remaining pitfalls.
-One drawback of our phase estimation method is a strong assumption of periodicity which may breakdown in the case of subjects with cardiac arrhythmia.
+One drawback of our phase estimation method is a strong assumption of periodicity which may \rk{fail} in the case of subjects with cardiac arrhythmia.
 To address this, we will look into univariate methods for phase estimation in quasi-periodic signals that are less susceptible noise.
@@
 %
-\vspace{-0.5cm}
+\vspace{-0.2cm}
 %

### 2017-02-24 paper155.tex
Incorporated roland's corrections

@@
 These tasks are particularly challenging in pre-clinical studies involving small animals such as mice that have rapid heart (310-840 BPM) and respiration (80-230 BPM) rates.
-Typically, the position within cardiac cycle is tracked through ECG data and position within \rk{the} respiratory cycle is tracked by motion of markers placed on the subject's body \cite{Khamene2004}.
+Typically, the position within cardiac cycle is tracked through ECG data and position within the respiratory cycle is tracked by motion of markers placed on the subject's body \cite{Khamene2004}.
 
@@
 \noindent
-\rk{\textbf{Contribution.}} In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video.
+\textbf{Contributions.} In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video.
 We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
@@
 \noindent
-\rk{\textbf{Related work.}} Previous work estimation of cardiac and/or respiratory phases directly from ultrasound videos is very limited.
+\textbf{Related work.} Previous work on estimation of instantaneous cardiac and/or respiratory phases directly from ultrasound videos is very limited.
 In~\cite{Karadayi2006}, the authors compute a signal of the x- or y-coordinate of the center-of-mass of each frame, use bandpass filtering to remove components outside the cardiac frequency range, and apply matched sine/cosine filters to estimate the instantaneous phase.
 In~\cite{Sundar2009}, the authors compute a signal of phase correlation between consecutive frames and use an approach similar to~\cite{Karadayi2006} to recover the cardiac and respiratory signals.
-The center-of-mass signals are not reliable and phase correlation finds global translation that does not model beating and out-of-plane \rk{heart motion} well.
+The center-of-mass signals are not reliable and phase correlation finds global translation that does not model beating and out-of-plane heart motion well.
 
@@
 In this section, we present the theory underlying the proposed methods.
-In Section~\ref{sec:method:phase_estimation}, we \rk{describe} our method for estimation of instantaneous cardiac and respiratory phases.
+In Section~\ref{sec:method:phase_estimation}, we describe our method for estimation of instantaneous cardiac and respiratory phases.
 In Section~\ref{sec:method:gating}, we present a robust method to exclude video frames with significant respiratory motion.
@@
 
-	We first compute the similarity between all pairs of images/frames in the given periodic image sequence containing $N$ frames to create a symmetric matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ \rk{frame}.
+	We first compute the similarity between all pairs of images/frames in the given periodic image sequence containing $N$ frames to create a symmetric matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frame.
 Here in, we use normalized correlation (also known as the Pearson correlation coefficient) to quantify inter-frame similarity; but in principle, other image similarity metrics can be used.
@@
 
-	In the first step, based on the observation that respiratory motion mainly occurs around the minima of the respiration signal $\hat{\tau}_{resp}(t)$ (pink curve in Fig.~\ref{fig:phase_estimation}(d)) with respiratory phase $\hat{\phi}_{resp}(t) = 0$, we perform a rough initial gating by discarding the frames $F_{cutoff} = \left \{ t \mid \hat{\phi}_{resp}(t) < c \parallel \hat{\phi}_{resp}(t) > (1 - c) \right \}$ whose phase distance from $\hat{\phi}_{resp}(t) = 0$ is below a specified cutoff value $c=0.2$.
+	In the first step, based on the observation that respiratory motion mainly occurs around the minima of the respiration signal $\hat{\tau}_{resp}(t)$ (pink curve in Fig.~\ref{fig:phase_estimation}(d)) with respiratory phase $\hat{\phi}_{resp}(t) = 0$, we perform a rough initial gating by discarding the frames $F_{cutoff} = \left \{ t \mid \hat{\phi}_{resp}(t) < c \vee \hat{\phi}_{resp}(t) > (1 - c) \right \}$ whose phase distance from $\hat{\phi}_{resp}(t) = 0$ is below a specified cutoff value $c=0.2$.
 Figure~\ref{fig:respiratory_gating}(a) shows the discarded frames $F_{cutoff}$ overlaid on the respiration $\hat{\tau}_{resp}(t)$, heart-beat  $\hat{r}_{heart}(t)$, and respiratory phase $\hat{\phi}_{resp}(t)$ signals.
 
-	In the second step, we learn a robust mapping $L(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called Locally weighted regression (LOWESS)~\cite{Cleveland1988} to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and heart-beat signal $\hat{r}_{heart}(t)$  values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above.
+	In the second step, we learn a robust mapping $L(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called Locally weighted regression (LOWESS)~\cite{Cleveland1988} to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid \forall t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and heart-beat signal $\hat{r}_{heart}(t)$  values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above.
 Next, we compute a robust estimate of the standard deviation $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit based on the median absolute deviation between the LOWESS fit $L( \hat{\phi}_{heart}(t) )$ and heart-beat signal $\hat{r}_{heart}(t)$ for all frames.
-Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert \hat{r}_{heart}(t) - L( \hat{\phi}_{heart}(t) ) \rvert   > k \times \hat{\sigma}_{L}  \right \}$
+Lastly, we gate out all frames $F_{resp} = \left \{ t : \lvert \hat{r}_{heart}(t) - L( \hat{\phi}_{heart}(t) ) \rvert   > k \times \hat{\sigma}_{L}  \right \}$
 whose heart-beat signal value deviates from the LOWESS fit by more than $k = 2.0$ (95\% confidence interval) times the standard deviation $\hat{\sigma}_{L}$ (Figure~\ref{fig:respiratory_gating}(b)).	
@@
 The second RBF kernel weighs images inversely proportional to the deviation of their residual/heart-beat signal value from the LOWESS prediction $L(\phi)$.
-Its bandwidth $\sigma_{L}$ is set equal to a constant $k_L = 2.0$ times the robust estimate of standard deviation $\hat{\sigma}_{L}$ of non-respiratory frames ($t \notin F_{resp}$) around the LOWESS fit.
+Its bandwidth $\sigma_{L}$ is set equal to a constant $k_L = 2.0$ times the robust estimate of standard deviation $\hat{\sigma}_{L}$ of non-respiratory frames ($\forall t \notin F_{resp}$) around the LOWESS fit.
 The function $M(\phi)$ can now be used to reconstruct a single-cycle video representative of the subject's heart-beat at any desired resolution/sampling of phase in the range $[0, 1)$.
@@
 We have also presented a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
-We next plan to evaluate our methods on more datasets and address any remaining pitfalls.
-One drawback of our phase estimation method is a strong assumption of periodicity which may \rk{fail} in the case of subjects with cardiac arrhythmia.
+We next plan to evaluate our methods on more datasets and address remaining pitfalls.
+Our phase estimation method makes a strong assumption of periodicity which may not hold in the case of subjects with cardiac arrhythmia.
 To address this, we will look into univariate methods for phase estimation in quasi-periodic signals that are less susceptible noise.
-Another drawback is the use of normalized correlation to measure inter-frame similarity relies on an inherent assumption that a large part of the image is pulsating.
+The use of normalized correlation to measure inter-frame similarity relies on an inherent assumption that a large part of the image is pulsating.
 To relax this, we will try to devise local patch-based similarity measures.

### 2017-02-24 paper155.tex
Added a comment about manifold kernel regression

@@
 Cardiac ultrasound videos consist of two kinds of periodic motion, one corresponding to beating heart motion and the other corresponding to respiratory motion.
-Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating \cite{VonBirgelen1997}, quiescence detection~\cite{Wick2013}, 3D reconstruction~\cite{Wachinger2012}, and temporal super-resolution~\cite{Cherin2006}.
+Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating \cite{Sundar2009}, quiescence detection~\cite{Wick2013}, 3D reconstruction~\cite{VonBirgelen1997}, and temporal super-resolution~\cite{Cherin2006}.
 These tasks are particularly challenging in pre-clinical studies involving small animals such as mice that have rapid heart (310-840 BPM) and respiration (80-230 BPM) rates.
@@
 \noindent
-\textbf{Contributions.} In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video.
+\textbf{Contributions.} In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video(Figure~\ref{fig:phase_estimation}).
 We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.
@@
 
+\vspace{-0.3cm}
 \section{Method}
@@
 Given a cardiac ultrasound video $I(t) : \{1, ..., N\} \to R^m$ of $N$ images with $m$ pixels each, we extract the heart-beat signal $\hat{r}_{heart}(t)$, estimate the instantaneous cardiac phase $\hat{\phi}_{heart}(t)$ of each frame, and compute the robust LOWESS fit $L(\phi) : \left [  0, 1\right ) \to R$ that maps cardiac phase to the heart-beat signal as described in Sections~\ref{sec:method:phase_estimation} and ~\ref{sec:method:gating}.
-We then use Nadarya-Watson kernel regression \cite{Bishop2006} to learn a function $M(\phi): [0, 1) \to R^m $ that reconstructs the image for any cardiac phase $\phi$ using a kernel-weighted local average as follows:
+We then use Nadarya-Watson (NW) kernel regression \cite{Bishop2006} to learn a function $M(\phi): [0, 1) \to R^m $ that reconstructs the image for any cardiac phase $\phi$ using a kernel-weighted local average as follows:
 \begin{equation}
@@
 To relax this, we will try to devise local patch-based similarity measures.
+Lastly, the local weighted average used by our NW kernel regression model may cause blurring at high temporal magnification.
+We will try to address this using manifold kernel regression~\cite{Davis2010} that computes the weighted average in a diffeomorphic registration sense.
 %

### 2017-02-24 paper155.tex
Minor corrections

@@
 The use of normalized correlation to measure inter-frame similarity relies on an inherent assumption that a large part of the image is pulsating.
-To relax this, we will try to devise local patch-based similarity measures.
+To relax this, we will devise local patch-based similarity measures.
 Lastly, the local weighted average used by our NW kernel regression model may cause blurring at high temporal magnification.
-We will try to address this using manifold kernel regression~\cite{Davis2010} that computes the weighted average in a diffeomorphic registration sense.
+We plan to alleviate this using manifold kernel regression~\cite{Davis2010} that computes the weighted average in a diffeomorphic registration sense.
 %

### 2017-02-24 paper155.tex
Minor correction
Updated LOOCV table with measures excluding corresponding frames from each cycle

@@
 \noindent
-\textbf{Related work.} Previous work on estimation of instantaneous cardiac and/or respiratory phases directly from ultrasound videos is very limited.
+\textbf{Related work.} Previous work on estimation of cardiac and/or respiratory phases directly from ultrasound videos is very limited.
 In~\cite{Karadayi2006}, the authors compute a signal of the x- or y-coordinate of the center-of-mass of each frame, use bandpass filtering to remove components outside the cardiac frequency range, and apply matched sine/cosine filters to estimate the instantaneous phase.
@@
 	We validate the accuracy of our kernel regression model for reconstructing images at any cardiac phase using leave-one-out-cross-validation (LOOCV).
-In each round of cross-validation, we randomly exclude one of the non-respiratory frames in the video, fit our kernel-regression model on the remaining frames, use the fitted model to reconstruct the image at the cardiac phase of the excluded frame, and compute the similarity between the reconstructed and original image using normalized correlation.
+In each round of cross-validation, we randomly pick one of the non-respiratory frames in the video, exclude the selected frame along with corresponding frames (closest in phase) in each cycle, fit our kernel-regression model on the remaining frames, use the fitted model to reconstruct the image at the cardiac phase of the selected frame, and compute the similarity between the reconstructed and original image using normalized correlation.
 The second column of Table~\ref{table:image_reconstruction} shows the $mean \pm std$ of normalized correlation between the reconstructed and the original images over 50 rounds of LOCCV for all 6 videos.
-As a \emph{baseline}, the third column of  Table~\ref{table:image_reconstruction} shows the $mean \pm std$ of the normalized correlation between frames at QRS peaks of the ECG signal.
+As a \emph{baseline}, the third column of  Table~\ref{table:image_reconstruction} reports the $mean \pm std$ of the normalized correlation between frames at QRS peaks of the ECG signal showing the effect of noise on similarity.
 The supplementary material includes single-cycle videos at 2x and 4x temporal resolution generated using the kernel regression model wherein the motion of heart chambers and valves is much clear than the original video.
@@
 \begin{tabular}[c]{@{}c@{}}VID \\ $\;$ \end{tabular} & \begin{tabular}[c]{@{}c@{}}LOOCV (50)\\ $mean \pm std \; ncorr$\end{tabular} & \begin{tabular}[c]{@{}c@{}}QRS Peak Frames\\ $mean \pm std \; ncorr$\end{tabular} \\ \hline
-1 & 0.8562 $\pm$ 0.0366 & 0.7981 $\pm$ 0.0532 \\ \hline
-2 & 0.8558 $\pm$ 0.0525 & 0.7877 $\pm$ 0.0503 \\ \hline
-3 & 0.8592 $\pm$ 0.0348 & 0.7849 $\pm$ 0.0457 \\ \hline
-4 & 0.8713 $\pm$ 0.0348 & 0.8395 $\pm$ 0.0486 \\ \hline
-5 & 0.8655 $\pm$ 0.0312 & 0.8196 $\pm$ 0.0476 \\ \hline
-6 & 0.8736 $\pm$ 0.0278 & 0.8270 $\pm$ 0.0778 \\ \hline
+1 & 0.8546 $\pm$ 0.0365 & 0.7981 $\pm$ 0.0532 \\ \hline
+2 & 0.8386 $\pm$ 0.0460 & 0.7877 $\pm$ 0.0503 \\ \hline
+3 & 0.8552 $\pm$ 0.0337 & 0.7849 $\pm$ 0.0457 \\ \hline
+4 & 0.8721 $\pm$ 0.0343 & 0.8395 $\pm$ 0.0486 \\ \hline
+5 & 0.8681 $\pm$ 0.0308 & 0.8196 $\pm$ 0.0476 \\ \hline
+6 & 0.8767 $\pm$ 0.0289 & 0.8270 $\pm$ 0.0778 \\ \hline
 \end{tabular}

### 2017-02-24 paper155.tex
Finished final edits ... strengthened motivation

@@
 Ultrasound is emerging as an increasingly effective tool for rapid non-invasive assessment of cardiac structure and function.
-Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in many applications and is particularly challenging in pre-clinical studies involving small animals with rapid heart and respiration rates.
+Knowledge of the phase of each video frame within the cardiac and/or respiratory cycles is essential in many applications and the ability to capture cardiac function at a sufficient temporal resolution is essential for accurate diagnosis.
+These tasks are particularly challenging in pre-clinical studies involving small animals with high heart and respiration rates.
 In this paper, we present a novel method for estimation of instantaneous cardiac and respiratory phases directly from cardiac ultrasound videos.
@@
 We also present a robust non-parametric regression technique for gating out respiratory frames and a kernel regression model that reconstructs images at any cardiac phase and facilitates temporal super-resolution.
-We validate our methods using cardiac ultrasound or echocardiography videos and ECG recordings of 6 mice.
+We validate our methods using 2D cardiac ultrasound or echocardiography videos and ECG recordings of 6 mice.
 %\keywords{Cardiac, Ultrasound, Echocardiography, Phase estimation, Gating, Temporal Super-resolution}
@@
 Cardiovascular disease is the leading cause of death worldwide and ultrasound is emerging as an increasingly effective tool for rapid non-invasive assessment of cardiac structure and function.
-Cardiac ultrasound videos consist of two kinds of periodic motion, one corresponding to beating heart motion and the other corresponding to respiratory motion.
-Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential in a variety of applications such as gating \cite{Sundar2009}, quiescence detection~\cite{Wick2013}, 3D reconstruction~\cite{VonBirgelen1997}, and temporal super-resolution~\cite{Cherin2006}.
-These tasks are particularly challenging in pre-clinical studies involving small animals such as mice that have rapid heart (310-840 BPM) and respiration (80-230 BPM) rates.
-Typically, the position within cardiac cycle is tracked through ECG data and position within the respiratory cycle is tracked by motion of markers placed on the subject's body \cite{Khamene2004}.
+Knowledge of the phase or location of each video frame within the cardiac and/or respiratory cycles is essential many applications (e.g.
+gating \cite{Sundar2009}, quiescence detection~\cite{Wick2013}, 3D reconstruction~\cite{VonBirgelen1997}) and the ability to capture cardiac function at a sufficient temporal resolution is vital for accurate diagnosis.
+Typically, the position within the cardiac cycle is tracked through ECG data and position within the respiratory cycle is tracked by motion of markers placed on the subject's body~\cite{Khamene2004}.
+Setting up such hardware is cumbersome particularly in pre-clinical studies involving small animals.
+Moreover, the frame rates of affordable commericial ultrasound transducers fall short~\cite{Cherin2006} in imaging small animals such as mice with high heart (310-840 BPM) and respiration (80-230 BPM) rates.
 

### 2017-02-24 paper155.tex
Correct error -- referenced wrong figure in introduction

@@
 \noindent
-\textbf{Contributions.} In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video(Figure~\ref{fig:phase_estimation}).
+\textbf{Contributions.} In this paper, we present a novel method that estimates the instantaneous cardiac and respiratory phases directly from the cardiac ultrasound video(Figure~\ref{fig:instaphase_vs_ecg}).
 We also present a robust non-parametric regression technique for gating out respiratory frames and a novel kernel regression model for reconstructing images at any cardiac phase to facilitate temporal super-resolution.

### 2017-02-24 paper155.tex
Minor corrections

@@
 Ultrasound is emerging as an increasingly effective tool for rapid non-invasive assessment of cardiac structure and function.
-Knowledge of the phase of each video frame within the cardiac and/or respiratory cycles is essential in many applications and the ability to capture cardiac function at a sufficient temporal resolution is essential for accurate diagnosis.
+Knowledge of the phase of each video frame within the cardiac and/or respiratory cycles is essential in many applications and the ability to capture cardiac function at a sufficient temporal resolution is necessary for accurate diagnosis.
 These tasks are particularly challenging in pre-clinical studies involving small animals with high heart and respiration rates.

### 2017-02-25 paper155.tex
Made minor corrections

@@
 These tasks are particularly challenging in pre-clinical studies involving small animals with high heart and respiration rates.
-In this paper, we present a novel method for estimation of instantaneous cardiac and respiratory phases directly from cardiac ultrasound videos.
+In this paper, we present a novel method for estimation of the instantaneous cardiac and respiratory phases directly from cardiac ultrasound videos.
 The method transforms the complex high-dimensional image sequence into a univariate time series with the same periodicity characteristics, decouples the periodic sources of beating heart and respiratory motion using a trend extraction technique, and estimates the cardiac and respiratory phases using the Hilbert transform.
@@
 If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted, this time series will preserve the periodicity characteristics of the original image sequence.
-Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos wherein the periodicity of low-frequency respiratory motion and high-frequency beating heart motion can observed.
+Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos wherein the periodicity of low-frequency respiratory motion and high-frequency beating heart motion can be observed.
 
@@
 \end{equation}
-where $\nabla^2\tau^i_{resp}(t) = \tau^i_{resp}(t+1) - 2 \tau^i_{resp}(t) + \tau^i_{resp}(t-1)$ is the second-order difference or derivative of the trend signal and $\lambda=6400$ is penalty parameter.
+where $\nabla^2\tau^i_{resp}(t) = \tau^i_{resp}(t+1) - 2 \tau^i_{resp}(t) + \tau^i_{resp}(t-1)$ is the second-order difference or derivative of the trend signal and $\lambda=6400$ is a penalty parameter.
 Let $A_{resp}$ and $A_{heart}$ be the matrices whose rows contain the trend/respiratory and residual/heart-beat components (Figures~\ref{fig:phase_estimation}(b,c)), respectively, of the frame similarity signal in the corresponding rows of matrix $A$.
@@
 
-	Next, considering the narrow-band nature of the respiration and heart-beat signals (see  Figure~\ref{fig:phase_estimation}(e)), we use the Hilbert transform~\cite{Boashash1992} to estimate the instantaneous phase of each frame.
+	Next, considering the narrow-band nature of the respiration and heart-beat signals (Figure~\ref{fig:phase_estimation}(e)), we use the Hilbert transform~\cite{Boashash1992} to estimate the instantaneous phase of each frame.
 Specifically, we compute the instantaneous phase $\phi(t) \in \left [  -\pi, \pi\right )$ of a periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows: $\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)$ and map $\phi(t)$ from $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$.
@@
 
-	In the second step, we learn a robust mapping $L(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called Locally weighted regression (LOWESS)~\cite{Cleveland1988} to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid \forall t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and heart-beat signal $\hat{r}_{heart}(t)$  values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above.
+	In the second step, we learn a regression function $L(\phi) : \left [  0, 1\right ) \to R$ to predict the heart-beat signal value for any cardiac phasel by fitting a robust non-parametric regression model called Locally weighted regression (LOWESS)~\cite{Cleveland1988} to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid \forall t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and heart-beat signal $\hat{r}_{heart}(t)$  values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above.
 Next, we compute a robust estimate of the standard deviation $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit based on the median absolute deviation between the LOWESS fit $L( \hat{\phi}_{heart}(t) )$ and heart-beat signal $\hat{r}_{heart}(t)$ for all frames.
@@
 Its bandwidth $\sigma_\phi$ is set equal to a constant $k_\phi = 0.4$ times the median difference in cardiac phase between consecutive frames of the given image sequence.
-The second RBF kernel weighs images inversely proportional to the deviation of their residual/heart-beat signal value from the LOWESS prediction $L(\phi)$.
+The second RBF kernel weighs images inversely proportional to the deviation of their heart-beat signal value from the LOWESS prediction $L(\phi)$.
 Its bandwidth $\sigma_{L}$ is set equal to a constant $k_L = 2.0$ times the robust estimate of standard deviation $\hat{\sigma}_{L}$ of non-respiratory frames ($\forall t \notin F_{resp}$) around the LOWESS fit.
@@
 Figure~\ref{fig:instaphase_vs_ecg}(a) shows a comparison between the ECG signal and the estimated cardiac phase signal for one of the 6 videos.
-Notice how well the locations of the QRS peaks in the ECG signal and the corresponding minima of the cardiac phase signal in each cycle match.
+Notice how well the locations of QRS complex peaks or R peaks in the ECG signal and the corresponding minima of the cardiac phase signal in each cycle match.
 Table 1 reports the $R^2$ correlation and the $mean \pm std$ error between the locations of the QRS peaks and the corresponding minima of the cardiac phase signals for all 6 videos.
